### 2017-09-17 article.tex
Submitted version

%  LaTeX support: latex@mdpi.com 
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

% You need to save the "mdpi.cls" and "mdpi.bst" files into the same folder as this template file.

%=================================================================
\documentclass[entropy,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi}

%
%--------------------
% Class Options:
%--------------------
% journal
%----------
% Choose between the following MDPI journals:
% actuators, admsci, aerospace, agriculture, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, atmosphere, atoms, axioms, batteries, behavsci, beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci, buildings, carbon, cancers, catalysts, cells, challenges, chemosensors, children, chromatography, climate, coatings, computation, computers, condensedmatter, cosmetics, cryptography, crystals, data, dentistry, designs, diagnostics, diseases, diversity, econometrics, economies, education, electronics, energies, entropy, environments, epigenomes, fermentation, fibers, fishes, fluids, foods, forests, futureinternet, galaxies, games, gels, genealogy, genes, geosciences, geriatrics, healthcare, horticulturae, humanities, hydrology, informatics, information, infrastructures, inorganics, insects, instruments, ijerph, ijfs, ijms, ijgi, ijtpp, inventions, jcdd, jcm, jdb, jfb, jfmk, jimaging, jof, jintelligence, jlpea, jmse, jpm, jrfm, jsan, land, languages, laws, life, literature, lubricants, machines, magnetochemistry, marinedrugs, materials, mathematics, mca, mti, medsci, medicines, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, molbank, molecules, mps, nanomaterials, ncrna, neonatalscreening, nutrients, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, plants, polymers, processes, proteomes, publications, recycling, religions, remotesensing, resources, risks, robotics, safety, sensors, separations, sexes, sinusitis, socsci, societies, soils, sports, standards, sustainability, symmetry, systems, technologies, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vetsci, viruses, water
%---------
% article
%---------
% The default type of manuscript is article, but can be replaced by: 
% addendum, article, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, correction, conferenceproceedings, conferencereport, expressionofconcern, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimage, letter, newbookreceived, opinion, obituary, projectreport, reply, retraction, review, preprints, shortnote, supfile, technicalnote
% supfile = supplementary materials
%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g. the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figure are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother 
\articlenumber{x}
\doinum{10.3390/------}
\pubvolume{xx}
\pubyear{2016}
\copyrightyear{2016}
\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{EEG characterization and classification based on Histogram of Gradients of Signal Plots}

% Authors, for the paper (add full first names)
\Author{Rodrigo Ramele $^{1,\dagger}$*, Ana Julia Villar $^{1}$ and Juan Miguel Santos  $^{1}$}
% Authors, for metadata in PDF
\AuthorNames{Rodrigo Ramele, Ana Julia Villar and Juan Miguel Santos}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
$^{1}$ \quad Computer Engineering Department, Instituto Tecnológico de Buenos Aires (ITBA); info@itba.edu.ar}

% Contact information of the corresponding author
\corres{Correspondence: rramele@itba.edu.ar; Tel.: +54-11-2150-4800(5834)}

% Current address and/or shared authorship
\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina}

% Simple summary
%\simplesumm{}

% Abstract (Do not use inserted blank lines, i.e. \\) 
\abstract{The analysis of Electroencephalographic (EEG) signals is of ulterior importance to elucidate patterns that could improve the implementation of Brain Computer Interfaces (BCI). These systems are meant to provide alternative pathways to transmit volitional information which could potentially enhance the quality of life of patients affected by neurodegenerative disorders or improve Human Computer Interaction systems.  Of particular interests are those which are based on the recognition of Event-Related Potentials (ERP) because they can be elicited by external stimuli and used to implement spellers, to control external devices or even avatars in virtual reality environments.  This work mimics what electroencephalographers have been doing clinically, visually inspecting and categorizing phenomena within the EEG by the extraction of features from the images of the plots of the signals.  It also aims to provide a framework to analyze, characterize and classify EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.  The validity of the method is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients.}

% Keywords
\keyword{electroencephalography (EEG); BCI; P300; ALS; classification; HOG; SIFT}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}
%\AMS{}

% If this is an expanded version of a conference paper, please cite it here: enter the full citation of your conference paper, and add $^\S$ in the end of the title of this article.
%\conference{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:

%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% For Conference Proceedings Papers: add the conference title here
%\conferencetitle{}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only for the journal Gels: Please place the Experimental Section after the Conclusions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.

\section{Introduction}

%\begin{enumerate}[leftmargin=*,labelsep=3mm]
%\item	Why
%\item	Purpose and significance
%\item	Current state of the research field and key publications cited
%\item   Controversial hypothesis
%\item   Main aim and principal conclusions.
%\end{enumerate}
%
%\begin{enumerate}[leftmargin=*,labelsep=3mm]
%\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
%\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
%\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
%\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
%\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
%\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
%\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
%\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
%\end{enumerate}

Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundred years old technology.  Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information \cite{DeVos2014}

A key contribution to this expansion has been the field of Brain Computer Interfaces (BCI) \citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neurodegenerative diseases.

One noteworthy aspect of this novel communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, in a Virtual Reality environment \citep{Lotte2013} or as aiding tool in a rehabilitation procedure \citep{Jure2016}.  The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.

EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process.  Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.  Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.

Moreover, EEG markers \citep{Clerc} that can be used to volitionally transmit information are limited, and each one of them has a particular combination of appropriate methods to decode them. Inevitably, it is necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.

%Patients suffering from ALS their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.

%Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.  It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.

BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon  \citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.  New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments \citep{Huggins2016}.  However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement \citep{Clerc}.

In this work, a new method to characterize and classify EEG signals is presented, expanded and detailed.  Its validity is verified by processing offline data for ALS patients.  This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by P300 Event Related Potential \citep{Knuth2006}.

The method is based on the morphological analysis of the shape of the EEG signal \citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots \citep{Hartman2005}.

This paper reports a method to, (1) characterize EEG signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from signal plots, and (2) how this characterization can be used to implement a BCI classification scheme to identify Event Related Potentials, particularly the well-known P300, on an offline and public dataset.

This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained: section \ref{Preprocessing} and \ref{Pipeline} describe the processing pipeline.  Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.  Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Near Neighbor  and the final Section shows results and discussion where we expose our remarks, conclusions and future work.

%\subsection{Morphological Signal Analysis}
%
%Several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
%CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016,Yamaguchi2009}.
%
%Morphological or 
%
%template matching
%
%spike
%sharp wave
%spike wave
%polyspike 
%slow wave complex
%
%shape domain
%
%\subsection{P300}
%
%Uhhhh I can talk a lot about P300 \citep{Knuth2006}.
%
%P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.

\section{Materials and Methods}

\subsection{Feature Extraction based on Histogram of Gradients of the Signal Plot} \label{Feature}

The P300 \citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs 300 ms after the onset of a rare and deviant stimulus that the subject is expected to attend.  It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and though it is quite consistent across different subjects, it has a lower amplitude ( $ \pm 5 \mu V $ ) compared to basal EEG activity, reaching a SNR of around -15 db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity \citep{Hu2010}.  This signal can be used to implement a speller application by means of a Speller Matrix \citep{Farwell1988} (Fig. \ref{fig:p300matrix}) where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.   Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{openvibep300matrix.png}
\caption{The P300 Matrix\citep{Farwell1988} used in the OpenVibe Open Source software \citep{Renard2010}. Rows are labeled from 1 to 6 and columns from 7 to 12.  Rows/columns flash intermittently in random permutations.}
\label{fig:p300matrix}
\end{figure}

%The proposed algorithm was designed to perform specifically this decoding.

% of the selected letter based on the detection of the P300 Event Related Potential response on the corresponding EEG signal.

\subsubsection{Preprocessing} \label{Preprocessing}

The first step consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. 
The processing pipeline starts by applying a notch filter to the raw signal, a 
4th degree 10 Hz lowpass Butterworth filter and finally a decimation with an FIR filter of order 30 from the original 256 Hz down to 16 Hz \citep{Krusienski2006}.

\subsubsection{Processing Pipeline} \label{Pipeline}

\begin{itemize}
\item \textbf{Artefact Removal}: The EEG signal matrix is processed on a channel by channel basis.   For every 12 flashing stimuli, i.e. one complete sequence of intensification of each of the 6 rows plus the 6 columns, a basic artefact elimination procedure is implemented by removing the entire segment when any signal deviates above/bellow $ \pm 70 \mu V $.
\item \textbf{Segmentation}: For each of the 12 stimuli,  a window of 1 second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.  Two of these segments are labeled as \textit{hit}, whereas the remaining 10 are labeled as \textit{no hit}.  A hit represents that the EEG segment should contain the ERP signature time-locked to the flashing stimulus. 
\item \textbf{Signal Averaging}: The P300 ERP is deeply buried under background EEG so the traditional approach to identify it is by point-to-point averaging the time-locked stacked signal segments.  Hence the values which are not related to, and not time-locked to the onset of the stimulus are canceled out \citep{Liang2008}. 
\end{itemize}

This last step determines the operation of any P300 Speller.  In order to obtain an improved signal in terms of its SNR, repetitions of the sequence of row/column intensification are necessary.  And, at the same time, as long as more repetitions are needed, the ability to transfer information faster is diminished, so there is a trade-off that must be acutely determined.

\subsubsection{Signal Plotting} \label{Plot}

The underlying idea of this method is to generate a template for the signal,  based on the image plot.  Hence, the first step is its transformation into a temporary binary image.

The signal is first scaled and standardized (i.e. z-score) by

\begin{equation}
\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma_x(c) } }\right \rfloor
\end{equation}

\noindent where $\gamma$ is the image scale, $t$ is the time and $ x(t,c) $ is the point-to-point averaged EEG matrix defined for each $t$ and for a channel $c$. Lastly, $\bar{x}(c)$ and $ \sigma_x $ are the mean and standard deviation of $x$.

Consequently, the image is constructed by placing the sample points according to

\begin{equation}
I(z_1,z_2) = \left\{ \begin{array}{rl}
255 & \text{if} \,  z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
0   & \mbox{otherwise}
\end{array}\right.
\label{eq:images}
\end{equation}

\noindent where $ z_1$ and $z_2$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image.  The function $z(c)$ is the \textit{zerolevel} which is the location on the image where the signal's zero value should be located in order to fit the entire signal within the image:

\begin{equation}
z(c) = \left \lfloor{ \frac{\max \tilde{x}(t,c)  - \min \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max \tilde{x}(t,c)  + \min \tilde{x}(t,c)}{ 2} }\right \rfloor
\label{eq:zerolevel}
\end{equation}

In order to complete the plot from the pixels, the Bresenham \citep{Ramele2016} algorithm is used to interpolate straight lines between each pair of these consecutive pixels.

%\noindent and is used to derive local representations that captures the visual shape of the signal.

\subsubsection{Feature Extraction: Histogram of Oriented Gradients} \label{SIFT}

%is calculated on local image patches centered on sample points \citep{Vedaldi2010}.

On the generated image, a keypoint is placed on a precise location along the signal trace over the image plot (Fig. \ref{fig:sampledescriptor}).  Around that keypoint, a local image patch is constructed as follows:  it is first divided in 16 blocks of size $ 3 \cdot s $ each, arranged in a 4x4 grid and centered on $T = (x,y)$.

A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the 16 blocks and creating a histogram of the oriented gradients.  This technique is based on Lowe's SIFT \citep{Lowe2004} descriptor, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations.   In order to calculate this histogram, the interval $ [0-360]$ of possible angles is divided in 8 bins, each one at 45 degrees.

The values of the image gradients for the 8 angle bins $ p = {0, 45, 90, 135, 180, 225, 270, 315} $ and for each grid block bin $ i,j = {1,2,3,4} $  are accumulated in the 3-dimensional histogram $ h $ through

\begin{equation}
 h(p,i,j) = 3 \cdot s \int w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - T}{3 s}\right)\, |J(\mathbf{x})|\, d\mathbf{x}\, d\theta_p
\label{eq:histogram}
\end{equation}

\noindent where $s$ is the size of the local patch, which can be converted into pixels by doing $ Px = 4 \cdot 3 \cdot s $ ,  $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each one of the 16 blocks of the patch, whereas $\angle J(\mathbf{x}) $ is the angle of the gradient vector and  $\theta_p$ is the angle bin.  On the other hand $ \textbf{x} = (x_i, x_j) $ corresponds to each one of the pixels of the local patch, and $ w_\mathrm{ang}  $  and $ w_{ij} $ are linear interpolation functions \citep{Lowe2004,Vedaldi2010}.  Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each grid block when the size of the patch is the unity.  As the patch is 4x4, it gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.

\begin{figure}[H]
\centering
\includegraphics[width=6cm]{sampledescriptor.png}
\includegraphics[width=9cm]{histogramsbines.png}
\caption{(Left) The local patch is located on a precise location within the generated plot. The 4x4 block grid can be seen on the patch. (Right) On each grid block, eight bins are used to describe the eight different orientations, and they are later arranged in a 128-dimension vector (4x4x8).}
\label{fig:sampledescriptor}
\end{figure}

\subsubsection{Classification} \label{Classification}

The classification is carried out by using a discriminative semi-supervised classification method, based on the Naive Bayes Nearest Neighbor (NBNN) \citep{Boiman2008}.

%The first step consists in extracting labeled descriptors from the training set for the  which are thus grouped together in two KD-tree \citep{Vedaldi2010} database structures. and group those marked as \textit{hit} in a template set $ T $

%BCI Simulation vs k-fold cross validation.  It is quite known and understood that because the same experience will not generate the same markers in the EEG, BCI Simulation is sometimes better suited to understand how the device will operate on a real online application (the graeat variability of BCI).

While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the image that were used to derive those descriptors.

The NBNN algorithm tackles this problem by comparing each image against a whole classification class which is characterized by the set of descriptors that are closest to each one of the descriptors obtained for the query image.  The algorithm is based on

\begin{equation}
\hat{C} = \arg \min_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
\label{eq:classification}
\end{equation}

\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.

This basic method was adapted to the P300 Speller and the classifier works according to the following modified equation:

\begin{equation}
\hat{Q}_{r\,c} = \arg \min_q \sum_{x_i \epsilon NN_T^k(q)}^{} \left\lVert x_i - q \right\rVert ^2
\label{eq:multiclassification}
\end{equation}

The first step consists in extracting labeled \textit{hit} descriptors from the training set which are thus grouped together in a KD-tree \citep{Vedaldi2010} template set $ T $.  On decoding stage, 12 new images from the signal segments are generated, and their descriptors extracted.  The 12 descriptors are divided in rows and columns.  For each one of the 6 query descriptors $ q $ for row/column, its $ k $ nearest neighbours in the template set $T$, $ NN_T^k(q) $ (acquired during the training phase)  are obtained and the distance between each one of them, $ x_i $, and the query descriptor $ q $ is summarized.  The query descriptor $ q $ that minimizes this summation is the one that is chosen, first between the 6 rows, and then between the remaining 6 columns.  By performing this procedure, the correct letter can be identified by matching the row $ r $ and column $ c $ from the P300 speller matrix.

Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent so that will unavoidably force the data to be unbalanced \citep{Tibon2015}.  The NBNN method suffers from biased classification on unbalanced classes \citep{Fornoni2014}.  By reversing the roles of the query and the class in Eq. \ref{eq:multiclassification}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.

%Additionally,

%El tema del balanceo lo mencionas y marcas lo cual está muy bien, pero no decí cómo lo solucionaste. Si llegaras a tener un poco de tiempo podrías mencionarlo.

\subsection{Experimental Protocol} \label{Protocol}

To verify the validity of the proposed framework and method, the public dataset 008-2014  \citep{Riccio2013} published on the BNCI-Horizon website \citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, was used to perform an offline BCI Simulation to decode the spelled words from the provided signals.  The algorithm was implemented using  VLFeat  \citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).

\subsubsection{P300 ALS Public Dataset}

The experimental protocol used to generate this dataset is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments. The P300 detection task designed for this experiment consisted of spelling 7 words (runs) of 5 letter each, using the traditional P300 Speller Matrix \citep{Farwell1988} where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.  The first 3 runs were used for training and the remaining 4 for testing with visual feedback.  A trial, as defined by the BCI2000 platform \citep{Schalk2004}, was every attempt to select a letter from the speller, and it was composed of signal segments corresponding to 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.  Flashing of row/columns was performed for 0.125 s, following by a resting period (i.e. Inter-stimulus interval) of the same length.  After 120 repetitions an inter-trial pause was included before resuming with the following letter.

The recorded dataset was sampled at 256 Hz and consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.

In order to asses and verify the identification of the P300 response, subjects are instructed to perform a copy-spelling task where they have to fix their attention to successive letters in order to copy a previously determined set of words (in contrast to an online free-running operation of the speller where each users decides on its own what letter to choose).

\subsubsection{Parameters}

Parameters were selected according to the experimental protocol. As the P300 event latency and amplitude vary greatly between subjects, it is necessary to provide a patch that will be able to capture an entire transient event.  Equations \ref{eq:mapping1} and \ref{eq:mapping2} can be used to map the original signal parameters to local image patch structure.

\begin{equation}
s = \frac{\Delta \mu V}{4 \cdot 3} \cdot \gamma 
\label{eq:mapping1}
\end{equation}

\begin{equation}
s = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
\label{eq:mapping2}
\end{equation}

\noindent where $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz), $ \lambda $ is the length in seconds covered by the patch, and $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch. By using $ s = 3 $ and a quadruple scale of the image $ \gamma = 4 $ this gives the local patch, and the descriptor, the ability to identify events of 9 $ \mu V $ of amplitude, a resolution of $ 1 $ Pixel $ = \frac{1}{4} \mu V $ and span   of $ \lambda = 0.56 s $ (because the patch is a geometric square, $ s $ must be the same to map both the height and the required span of the signal).  Finally, descriptor locations $ T $ were selected at $ x =  0.55 s \cdot Fs \cdot \gamma = 35 $ and $ y = z(c) $ (Eq. \ref{eq:zerolevel}).

%\begin{figure}[H]
%\centering
%\includegraphics[width=16cm]{localpatch.png}
%\caption{The local patch is mapped to the original image by their correspondence .}
%\label{fig:sampledescriptor}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Discussion} \label{Results}
\label{section:results}
In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects using the information from all the segments can be shown.  The P300 characteristic curve can be seen particularly in subjects 2, and 6 and in a lesser extend in the remaining subjects. In order to correctly decode the selected letter from each trial, particular care was observed to avoid unbalanced number of epochs  (i.e. an unequal number of epochs on each condition), because that may introduce a bias in the classification procedure (the variance of averaged signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).

\begin{figure}[H]
\centering
\includegraphics[width=16cm]{subjectaveraged.eps}
\caption{Point-to-point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz. The P300 characteristic curve can be well identified particularly on subjects 2 and 6.}
\label{fig:subjectaveraged}
\end{figure}

Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.  From the 7 trials for each subject, the first 3 were used as training, and the remaining 4 for testing.  Additionally the best performing channel is informed as well.  It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.

\begin{table}[H]
\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject. Chance level is 0.02 }
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Participant}	&  \textbf{BPC}	& \textbf{Performance}\\
\midrule
1     &     Cz   &   0.35  \\
2     &     Fz   &   0.85  \\
3     &     Cz   &   0.25  \\
4     &     PO8 &   0.55  \\
5     &     PO7 &   0.40 \\
6     &     PO7 &   0.60  \\
7     &     PO8 &   0.80  \\
8     &     PO7 &   0.95 \\

\bottomrule
\end{tabular}
\label{tab:results}
\end{table}

The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.  The Performance curves (Fig. \ref{fig:performance}) show how the percentage of correctly identified letters depends on the number of repetitions that were used to obtain the averaged signal.

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{performance.eps}
\caption{Performance curves for the eight subjects included in the published dataset.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
\label{fig:performance}
\end{figure}

We found that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that could allow the classification to reach a higher level.  We hypothesize that as subjects may have different latencies and amplitudes of the P300 signal \citep{Riccio2013}, it may also be the case that the shape of the generated ERP may vary greatly in an intra-subject manner, thus the pattern could not be well generalized and a higher performance could not be reached (Fig.  \ref{fig:p300templates}).

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{subject8.png}
\includegraphics[width=10cm]{subject3.png}
\caption{Ten P300 template patches found for subjects 8 (up) and 3(down).  In coincidence with the performance results, the P300 signature is more clear and consistent for subject 8 (higher performance) while for subject 3 (lower performance) the characteristic pattern is much more difficult to perceive.}
\label{fig:p300templates}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Discussion}
%
%This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect 
%
%
%is also based on how the image look like.
%
%SNR of p300 and how to detect it
%
%Check if you can use this to detect any kind of transient signal.
%
%Compare if it is possible with the descriptors from one subject, discriminate the others.
%
%Channal identification based on the metric distance between the bags
%
%Signal Averaging.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

A method to characterize and classify EEG signals where their characterization is transient in time-space, like the P300 ERP, has been presented.

%Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).

The adaptive behaviour of the algorithm make it well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.  This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e. \textit{how actually they looked like}) for the training and calibration step, and they do not require any prior knowledge about the signal.   In contrast, when the shape of the ERP response is not consistent across the same subject, this method would not be able to find the templates and speller performance could be penalized.

At the same time, by analyzing the generated descriptors, which map in a very detailed and synthetic structure the shape information contained within the patch, a metric about the consistency of the shape of the generated P300 response could also be derived.

%It may be worthy of further consideration to evaluate if is there any correlation between the lower performance obtained for some subjects and their characterization of the ALS progression and degree.

We believe that the expanding and the understanding of this tool in order to automatically classify those patterns in EEG that are specifically identified by their shapes (e.g. K-Complex, Vertex Waves, Positive Occipital Sharp Transient \citep{Hartman2005}) is a prospect future work to be considered.  It may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods, frequent in sleep research.

Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc}, i.e. to detect BCI illiteracy or to predict the achievable performance of a given method, or even as a tool for artefact removal (which is performed on many occasions by visually inspecting the signal).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Subsection}
%
%\subsubsection{Subsubsection}
%
%Bulleted lists look like this:
%\begin{itemize}[leftmargin=*,labelsep=4mm]
%\item	First bullet
%\item	Second bullet
%\item	Third bullet
%\end{itemize}
%
%Numbered lists can be added as follows:
%\begin{enumerate}[leftmargin=*,labelsep=3mm]
%\item	First item
%\item	Second item
%\item	Third item
%\end{enumerate}
%
%The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictofinterests{The authors declare that there is no conflict of interest regarding the publication of this article.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent EEG: Electroencephalography\\
BCI: Brain Computer Interfaces\\
SNR: Signal to Noise Ratio\\
CNS: Central Nervous System\\
ALS: Amyotrophic Lateral Sclerosis\\
ERP: Event-Related Potential\\
P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimulus\\
ITR: Information Transfer Rate\\
BTR: Bit Transfer Rate\\
SIFT: Scale Invariant Feature Transform\\
NBNN: Naive Bayes Nearest Neighbor\\
HOG: Histogram Of Gradients}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
%\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A?). If there is only one appendix section then change the argument to ?one? and no counter is printed (?Appendix?).
%\appendix
%\section{}
%The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathemtaical proofs of results not central to the paper can be added as an appendix.
%
%\section{}
%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here. 
\bibliographystyle{mdpi}

%=====================================
% References, variant A: internal bibliography
%=====================================
%\renewcommand\bibname{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem{ref-journal}
%Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
% Reference 2
%\bibitem{ref-book}
%Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
%\end{thebibliography}

%=====================================
% References, variant B: external bibliography
%=====================================
\bibliography{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\sampleavailability{Samples of the compounds ...... are available from the authors.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2017-09-17 articlebackup.tex
Submitted version

--- 2017-09-17 article.tex
@@
 % Full title of the paper (Capitalized)
-\Title{EEG characterization and classification based on Histogram of Gradients of Signal Plots}
+\Title{EEG characterization and classification based on histograms of visual gradients}
 
 % Authors, for the paper (add full first names)
-\Author{Rodrigo Ramele $^{1,\dagger}$*, Ana Julia Villar $^{1}$ and Juan Miguel Santos  $^{1}$}
+\Author{Rodrigo Ramele $^{1,\dagger}$, Ana Julia Villar $^{1}$ and Juan Miguel Santos  $^{1}$*}
 % Authors, for metadata in PDF
@@
 % Contact information of the corresponding author
-\corres{Correspondence: rramele@itba.edu.ar; Tel.: +54-11-2150-4800(5834)}
+\corres{Correspondence: rramele@itba.edu.ar; Tel.: +54-9-11-4193-9382}
 
@@
 % Abstract (Do not use inserted blank lines, i.e. \\) 
-\abstract{The analysis of Electroencephalographic (EEG) signals is of ulterior importance to elucidate patterns that could improve the implementation of Brain Computer Interfaces (BCI).
-These systems are meant to provide alternative pathways to transmit volitional information which could potentially enhance the quality of life of patients affected by neurodegenerative disorders or improve Human Computer Interaction systems.
-Of particular interests are those which are based on the recognition of Event-Related Potentials (ERP) because they can be elicited by external stimuli and used to implement spellers, to control external devices or even avatars in virtual reality environments.
-This work mimics what electroencephalographers have been doing clinically, visually inspecting and categorizing phenomena within the EEG by the extraction of features from the images of the plots of the signals.
-It also aims to provide a framework to analyze, characterize and classify EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.
-The validity of the method is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients.}
+\abstract{The analysis of Electroencephalograpy (EEG) signals for Amyotrophic Lateral Sclerosis (ALS) patients is of ulterior importance to elucidate different patterns or markers that could be potentially used to aid in diagnosis or treatment and to implement Brain Computer Interface (BCI) systems which may provide alternative pathways to transmit volitional information to enhance patient's quality of life.
+Of particular interests are those which are based on the recognition of Event-Related Potentials (ERP) which are shown to be more robust to the unavoidable physical impairment which comes with this motor neuron disease.
+This work mimics what electroencephalographers has been doing clinically, visually inspecting and categorizing phenomena within the EEG, and provides a framework to analyse and characterize EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.
+The validity of the method is shown by off-line processing a public dataset of ALS patients.}
 
@@
 
-Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundred years old technology.
-Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information \cite{DeVos2014}
+Although impressive advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} has diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundredth years old technology.
+Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, is pushing EEG to become the de-facto non invasive method to access and harness brain information.
 
-A key contribution to this expansion has been the field of Brain Computer Interfaces (BCI) \citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neurodegenerative diseases.
+Key contribution to this development has been the field of Brain Computer Interfaces (BCI) \citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neuro-degenerative diseases.
 
-One noteworthy aspect of this novel communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, in a Virtual Reality environment \citep{Lotte2013} or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
+One very noteworthy aspect of this communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS), to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
 The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.
 
-EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process.
-Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.
+EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process and even considered random enough as to be used as a source of pseudo random number generator \citep{Chen2014}.
+Additionally, they have a high variability between different subjects and even between different moments for the same subject, requiring an adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.
 Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.
 
-Moreover, EEG markers \citep{Clerc} that can be used to volitionally transmit information are limited, and each one of them has a particular combination of appropriate methods to decode them.
-Inevitably, it is necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
+EEG markers \citep{Clerc} that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of appropriate methods to decode them.
+Inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
 
-%Patients suffering from ALS their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
+In this work a new method to characterize EEG signals is presented, expanded and detailed.
+Its validity is verified by processing off-line data for ALS patients.
+This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by the P300\citep{Knuth2006}.
 
-%Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
-It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
+The method is based on the morphological analysis of the shape of the EEG signal\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signals.
 
-BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon  \citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
-New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments \citep{Huggins2016}.
-However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement \citep{Clerc}.
+%This paper reports on a study done to address
+%the hypothesis that communication task complexity
+%(copy-spelling vs. novel text generation of a picture
+%description) is correlated with BCI performance accuracy
+%and P300 characteristics.
 
-In this work, a new method to characterize and classify EEG signals is presented, expanded and detailed.
-Its validity is verified by processing offline data for ALS patients.
-This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by P300 Event Related Potential \citep{Knuth2006}.
+In brief, this paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials, particularly the well-known P300, on an off-line and public dataset.
 
-The method is based on the morphological analysis of the shape of the EEG signal \citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots \citep{Hartman2005}.
+\section{Materials and Methods}
 
-This paper reports a method to, (1) characterize EEG signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from signal plots, and (2) how this characterization can be used to implement a BCI classification scheme to identify Event Related Potentials, particularly the well-known P300, on an offline and public dataset.
+Besides new novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
+It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns \citep{Nijboer2009,Riener2014}.
 
-This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained: section \ref{Preprocessing} and \ref{Pipeline} describe the processing pipeline.
-Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.
-Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Near Neighbor  and the final Section shows results and discussion where we expose our remarks, conclusions and future work.
+BCI has gained mainstream public awareness and recently even held a peculiar Olympics \citep{Riener2014} and been broadcasted during the inauguration of the last World Cup.
+New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments\citep{Huggins2016}.
+However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of remaining muscular movement \citep{Clerc}.
+
+To verify the validity of the proposed framework and method, the public dataset 008-2014 published on the BNCI-Horizon website \citep{Riccio2013} was used to perform a binary classification task on the provided signals.
+The algorithm was implemented using the VLFeat \citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).
+
+\subsection{Experimental protocol}
+
+The protocol is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
+The P300 task objective was meant to spell 7 words of 5 letter each.
+The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
+A trial, for this experiment, was defined as every attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.
+Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.
+After 120 repetions and inter-trial pause was included before resuming with the following letter.
+
+The gathered dataset was sampled at 256 Hz and consisted of EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
+This dataset was published in the popular BNCI-Horizon 2020 website \citep{Brunner2014}.
 
@@
 
-\section{Materials and Methods}
+\subsection{Algorithm}
 
-\subsection{Feature Extraction based on Histogram of Gradients of the Signal Plot} \label{Feature}
+\subsubsection{Preprocessing}
 
-The P300 \citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs 300 ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
-It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and though it is quite consistent across different subjects, it has a lower amplitude ( $ \pm 5 \mu V $ ) compared to basal EEG activity, reaching a SNR of around -15 db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity \citep{Hu2010}.
-This signal can be used to implement a speller application by means of a Speller Matrix \citep{Farwell1988} (Fig.
-\ref{fig:p300matrix}) where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
-Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
+The mandatory first step is to help to enhance the SNR of the P300 pattern above the level of basal EEG.
+The processing pipeline starts with applying a notch filter to the raw signal, a decimation with a FIR filter of order 30 from the original 256 Hz to 32 Hz and finally applying a 4th degree 10 Hz lowpass Butterworth filter.
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=10cm]{openvibep300matrix.png}
-\caption{The P300 Matrix\citep{Farwell1988} used in the OpenVibe Open Source software \citep{Renard2010}.
-Rows are labeled from 1 to 6 and columns from 7 to 12.
-Rows/columns flash intermittently in random permutations.}
-\label{fig:p300matrix}
-\end{figure}
+\subsubsection{Segmentation, Artefact Removal and Signal Averaging}
 
-%The proposed algorithm was designed to perform specifically this decoding.
+The EEG matrix is processed on a channel by channel basis.
+For each word, each letter, a segment of EEG signal is extracted that encompass 10 repetitions of 12 flashings, 6 rows and 6 columns.
+For every 12 stimuli flashing, i.e.
+one complete sequence of repetitions of each row/column, a basic artefact removal procedure of removing the entire segment if any signal goes above 70 $\mu$V is applied.
 
-% of the selected letter based on the detection of the P300 Event Related Potential response on the corresponding EEG signal.
+From every segment of 12 flashings, two epoch corresponding to a hit and two matching a no-hit are constructed as segments of 1s length (256 sample points).
 
-\subsubsection{Preprocessing} \label{Preprocessing}
+These segments, 20 for each class, are later point-to-point averaged for the whole 120 repetitions group.
+This finally lead to 2 averaged signals for each of the 5 letters, for the 7 words, comprising a set of 70 averaged signals\citep{Liang2008}.
 
-The first step consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. 
-The processing pipeline starts by applying a notch filter to the raw signal, a 
-4th degree 10 Hz lowpass Butterworth filter and finally a decimation with an FIR filter of order 30 from the original 256 Hz down to 16 Hz \citep{Krusienski2006}.
+\subsubsection{Features}
 
-\subsubsection{Processing Pipeline} \label{Pipeline}
-
-\begin{itemize}
-\item \textbf{Artefact Removal}: The EEG signal matrix is processed on a channel by channel basis.
-For every 12 flashing stimuli, i.e.
-one complete sequence of intensification of each of the 6 rows plus the 6 columns, a basic artefact elimination procedure is implemented by removing the entire segment when any signal deviates above/bellow $ \pm 70 \mu V $.
-\item \textbf{Segmentation}: For each of the 12 stimuli,  a window of 1 second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
-Two of these segments are labeled as \textit{hit}, whereas the remaining 10 are labeled as \textit{no hit}.
-A hit represents that the EEG segment should contain the ERP signature time-locked to the flashing stimulus.
-\item \textbf{Signal Averaging}: The P300 ERP is deeply buried under background EEG so the traditional approach to identify it is by point-to-point averaging the time-locked stacked signal segments.
-Hence the values which are not related to, and not time-locked to the onset of the stimulus are canceled out \citep{Liang2008}.
-\end{itemize}
-
-This last step determines the operation of any P300 Speller.
-In order to obtain an improved signal in terms of its SNR, repetitions of the sequence of row/column intensification are necessary.
-And, at the same time, as long as more repetitions are needed, the ability to transfer information faster is diminished, so there is a trade-off that must be acutely determined.
-
-\subsubsection{Signal Plotting} \label{Plot}
-
-The underlying idea of this method is to generate a template for the signal,  based on the image plot.  Hence, the first step is its transformation into a temporary binary image.
-
-The signal is first scaled and standardized (i.e. z-score) by
+The central part of this method is the feature generation: the histogram of visual gradients.  To do so, the first step is the transformation of the signal into a temporary binary image.
 
 \begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma_x(c) } }\right \rfloor
+\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot ( x(t,c) - \bar{x}(t,c)  )}\right \rfloor
 \end{equation}
 
-\noindent where $\gamma$ is the image scale, $t$ is the time and $ x(t,c) $ is the point-to-point averaged EEG matrix defined for each $t$ and for a channel $c$.
-Lastly, $\bar{x}(c)$ and $ \sigma_x $ are the mean and standard deviation of $x$.
-
-Consequently, the image is constructed by placing the sample points according to
+where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the EEG matrix defined for each $t$ and for a particular channel $c$.
 
@@
 I(z_1,z_2) = \left\{ \begin{array}{rl}
-255 & \text{if} \,  z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
+255 & z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
 0   & \mbox{otherwise}
@@
 
-\noindent where $ z_1$ and $z_2$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image.
-The function $z(c)$ is the \textit{zerolevel} which is the location on the image where the signal's zero value should be located in order to fit the entire signal within the image:
+From the Equation \ref{eq:images} the image $ I(z_1,z_2) $ is constructed an used to derive local representations that will capture the visual shape of the signal.
+
+Finally, the histogram of gradients, which is based on a very successful technique from Computer Vision, the SIFT \citep{Lowe2004} descriptor, is calculated on local image patches centered on sample points \citep{Vedaldi2010}.
 
 \begin{equation}
-z(c) = \left \lfloor{ \frac{\max \tilde{x}(t,c)  - \min \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max \tilde{x}(t,c)  + \min \tilde{x}(t,c)}{ 2} }\right \rfloor
-\label{eq:zerolevel}
-\end{equation}
-
-In order to complete the plot from the pixels, the Bresenham \citep{Ramele2016} algorithm is used to interpolate straight lines between each pair of these consecutive pixels.
-
-%\noindent and is used to derive local representations that captures the visual shape of the signal.
-
-\subsubsection{Feature Extraction: Histogram of Oriented Gradients} \label{SIFT}
-
-%is calculated on local image patches centered on sample points \citep{Vedaldi2010}.
-
-On the generated image, a keypoint is placed on a precise location along the signal trace over the image plot (Fig.
-\ref{fig:sampledescriptor}).
-Around that keypoint, a local image patch is constructed as follows:  it is first divided in 16 blocks of size $ 3 \cdot s $ each, arranged in a 4x4 grid and centered on $T = (x,y)$.
-
-A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the 16 blocks and creating a histogram of the oriented gradients.
-This technique is based on Lowe's SIFT \citep{Lowe2004} descriptor, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations.
-In order to calculate this histogram, the interval $ [0-360]$ of possible angles is divided in 8 bins, each one at 45 degrees.
-
-The values of the image gradients for the 8 angle bins $ p = {0, 45, 90, 135, 180, 225, 270, 315} $ and for each grid block bin $ i,j = {1,2,3,4} $  are accumulated in the 3-dimensional histogram $ h $ through
-
-\begin{equation}
- h(p,i,j) = 3 \cdot s \int w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - T}{3 s}\right)\, |J(\mathbf{x})|\, d\mathbf{x}\, d\theta_p
+ h(p,i,j) = 3 \cdot s \int w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - T}{3 s}\right)\, |J(\mathbf{x})|\, d\mathbf{x}
 \label{eq:histogram}
@@
 
-\noindent where $s$ is the size of the local patch, which can be converted into pixels by doing $ Px = 4 \cdot 3 \cdot s $ ,  $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each one of the 16 blocks of the patch, whereas $\angle J(\mathbf{x}) $ is the angle of the gradient vector and  $\theta_p$ is the angle bin.
-On the other hand $ \textbf{x} = (x_i, x_j) $ corresponds to each one of the pixels of the local patch, and $ w_\mathrm{ang}  $  and $ w_{ij} $ are linear interpolation functions \citep{Lowe2004,Vedaldi2010}.
-Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each grid block when the size of the patch is the unity.
-As the patch is 4x4, it gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.
+where $s$ is the size of the local patch, $\angle J(\mathbf{x})$ is the gradient found at each of the 16 blocks of the patch aligned with the angle $\theta_p$.
+The patch is divided in 16 blocks, arranged in a 4x4 grid and centered on $T = (x,y)$, a location within the image.
+On each block, the bin consists of 45 degrees obtaining 8 aligning angles.
+The Equation \ref{eq:histogram} accumulates the values of the image gradients aligned with each angle bin.
+This gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.
 
@@
 \centering
-\includegraphics[width=6cm]{sampledescriptor.png}
-\includegraphics[width=9cm]{histogramsbines.png}
-\caption{(Left) The local patch is located on a precise location within the generated plot.
-The 4x4 block grid can be seen on the patch.
-(Right) On each grid block, eight bins are used to describe the eight different orientations, and they are later arranged in a 128-dimension vector (4x4x8).}
+\includegraphics[width=16cm]{sampledescriptor.eps}
+\caption{The local patch is located around a sample point plotted on the temporary signal.
+All the sample points are interpolated using the Bresenham algorithm.
+The 4x4 grid can be seen on the descriptor and also arrows showing the dominant gradient on each block.}
 \label{fig:sampledescriptor}
@@
 
-\subsubsection{Classification} \label{Classification}
+\subsubsection{Classification}
 
-The classification is carried out by using a discriminative semi-supervised classification method, based on the Naive Bayes Nearest Neighbor (NBNN) \citep{Boiman2008}.
+The classification is carried out by using the discriminative semi-supervised classification method, the Naive Bayes Nearest Neighbout (NBNN) \citep{Boiman2008}.
+% algorithm which can be described as follows.
 
-%The first step consists in extracting labeled descriptors from the training set for the  which are thus grouped together in two KD-tree \citep{Vedaldi2010} database structures.
-and group those marked as \textit{hit} in a template set $ T $
+\begin{equation}
+\hat{C} = argmin_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
+\end{equation}
 
@@
 
-While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the image that were used to derive those descriptors.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
 
-The NBNN algorithm tackles this problem by comparing each image against a whole classification class which is characterized by the set of descriptors that are closest to each one of the descriptors obtained for the query image.
-The algorithm is based on
-
-\begin{equation}
-\hat{C} = \arg \min_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
-\label{eq:classification}
-\end{equation}
-
-\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.
-
-This basic method was adapted to the P300 Speller and the classifier works according to the following modified equation:
-
-\begin{equation}
-\hat{Q}_{r\,c} = \arg \min_q \sum_{x_i \epsilon NN_T^k(q)}^{} \left\lVert x_i - q \right\rVert ^2
-\label{eq:multiclassification}
-\end{equation}
-
-The first step consists in extracting labeled \textit{hit} descriptors from the training set which are thus grouped together in a KD-tree \citep{Vedaldi2010} template set $ T $.
-On decoding stage, 12 new images from the signal segments are generated, and their descriptors extracted.
-The 12 descriptors are divided in rows and columns.
-For each one of the 6 query descriptors $ q $ for row/column, its $ k $ nearest neighbours in the template set $T$, $ NN_T^k(q) $ (acquired during the training phase)  are obtained and the distance between each one of them, $ x_i $, and the query descriptor $ q $ is summarized.
-The query descriptor $ q $ that minimizes this summation is the one that is chosen, first between the 6 rows, and then between the remaining 6 columns.
-By performing this procedure, the correct letter can be identified by matching the row $ r $ and column $ c $ from the P300 speller matrix.
-
-Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent so that will unavoidably force the data to be unbalanced \citep{Tibon2015}.
-The NBNN method suffers from biased classification on unbalanced classes \citep{Fornoni2014}.
-By reversing the roles of the query and the class in Eq.
-\ref{eq:multiclassification}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.
-
-%Additionally,
-
-%El tema del balanceo lo mencionas y marcas lo cual está muy bien, pero no decí cómo lo solucionaste. Si llegaras a tener un poco de tiempo podrías mencionarlo.
-
-\subsection{Experimental Protocol} \label{Protocol}
-
-To verify the validity of the proposed framework and method, the public dataset 008-2014  \citep{Riccio2013} published on the BNCI-Horizon website \citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, was used to perform an offline BCI Simulation to decode the spelled words from the provided signals.
-The algorithm was implemented using  VLFeat  \citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).
-
-\subsubsection{P300 ALS Public Dataset}
-
-The experimental protocol used to generate this dataset is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
-The P300 detection task designed for this experiment consisted of spelling 7 words (runs) of 5 letter each, using the traditional P300 Speller Matrix \citep{Farwell1988} where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
-The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
-A trial, as defined by the BCI2000 platform \citep{Schalk2004}, was every attempt to select a letter from the speller, and it was composed of signal segments corresponding to 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.
-Flashing of row/columns was performed for 0.125 s, following by a resting period (i.e.
-Inter-stimulus interval) of the same length.
-After 120 repetitions an inter-trial pause was included before resuming with the following letter.
-
-The recorded dataset was sampled at 256 Hz and consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
-
-In order to asses and verify the identification of the P300 response, subjects are instructed to perform a copy-spelling task where they have to fix their attention to successive letters in order to copy a previously determined set of words (in contrast to an online free-running operation of the speller where each users decides on its own what letter to choose).
-
-\subsubsection{Parameters}
-
-Parameters were selected according to the experimental protocol.
-As the P300 event latency and amplitude vary greatly between subjects, it is necessary to provide a patch that will be able to capture an entire transient event.
-Equations \ref{eq:mapping1} and \ref{eq:mapping2} can be used to map the original signal parameters to local image patch structure.
-
-\begin{equation}
-s = \frac{\Delta \mu V}{4 \cdot 3} \cdot \gamma 
-\label{eq:mapping1}
-\end{equation}
-
-\begin{equation}
-s = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
-\label{eq:mapping2}
-\end{equation}
-
-\noindent where $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz), $ \lambda $ is the length in seconds covered by the patch, and $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
-By using $ s = 3 $ and a quadruple scale of the image $ \gamma = 4 $ this gives the local patch, and the descriptor, the ability to identify events of 9 $ \mu V $ of amplitude, a resolution of $ 1 $ Pixel $ = \frac{1}{4} \mu V $ and span   of $ \lambda = 0.56 s $ (because the patch is a geometric square, $ s $ must be the same to map both the height and the required span of the signal).
-Finally, descriptor locations $ T $ were selected at $ x =  0.55 s \cdot Fs \cdot \gamma = 35 $ and $ y = z(c) $ (Eq.
-\ref{eq:zerolevel}).
-
-%\begin{figure}[H]
-%\centering
-%\includegraphics[width=16cm]{localpatch.png}
-%\caption{The local patch is mapped to the original image by their correspondence .}
-%\label{fig:sampledescriptor}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results and Discussion} \label{Results}
-\label{section:results}
-In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects using the information from all the segments can be shown.
+In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects is shown.
 The P300 characteristic curve can be seen particularly in subjects 2, and 6 and in a lesser extend in the remaining subjects.
-In order to correctly decode the selected letter from each trial, particular care was observed to avoid unbalanced number of epochs  (i.e.
-an unequal number of epochs on each condition), because that may introduce a bias in the classification procedure (the variance of averaged signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).
+In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (signal variance is proportional to the square root of the number of samples and the procedure would be discriminating signals with different variances).Particularly in the case of the P300 response, the oddball paradigm requires that one of the stimulus need to be infrequent so that will unavoidably force that the data will be unbalanced (that is, an unequal number of trials in each condition).
 
@@
 \includegraphics[width=16cm]{subjectaveraged.eps}
-\caption{Point-to-point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
-The P300 characteristic curve can be well identified particularly on subjects 2 and 6.}
+\caption{Point-To-Point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
+The P300 characteristic curve can be well identified on subjects 2 and 6.}
 \label{fig:subjectaveraged}
@@
 
-Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
-From the 7 trials for each subject, the first 3 were used as training, and the remaining 4 for testing.
-Additionally the best performing channel is informed as well.
-It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
+Results are shown in Table \ref{tab:results} where BCI accuracy
+for the Cz channel is calculated and compared with the values obtained by the published of the dataset.
+Additionally the best performing channel is informed and its accuracy value.
+It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed very good performance indeed \citep{Huggins2016,Jure2016}.
+
+% (CITE TAYLOR AND  FRANCOS PAPER).
 
 \begin{table}[H]
-\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject. Chance level is 0.02 }
+\caption{Results obtained for Accuracy levels by 3-fold cross validation.
+The values reported by the dataset publishers for Cz are reproduced here for comparison.
+Additionally, channel Cz BCI accuracies can be seen as well as the performance levels obtained for the BPC, the best performing channel for each subject.}
 \centering
 %% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccc}
+\begin{tabular}{ccccc}
 \toprule
-\textbf{Participant}	&  \textbf{BPC}	& \textbf{Performance}\\
+\textbf{Participants}	& \textbf{Original Cz}	& \textbf{ACC at Cz}	& \textbf{BPC}	& \textbf{Performance}\\
 \midrule
-1     &     Cz   &   0.35  \\
-2     &     Fz   &   0.85  \\
-3     &     Cz   &   0.25  \\
-4     &     PO8 &   0.55  \\
-5     &     PO7 &   0.40 \\
-6     &     PO7 &   0.60  \\
-7     &     PO8 &   0.80  \\
-8     &     PO7 &   0.95 \\
-
+1     &   0.84 &   0.79 & Cz  &   0.79 $\pm$ 0.01 \\
+2     &   0.86 &   0.81 & PO7 &   0.93 $\pm$ 0.01 \\
+3     &   0.87 &   0.78 & Cz  &   0.78 $\pm$ 0.03 \\
+4     &   0.86 &   0.68 & PO8 &   0.90 $\pm$ 0.01 \\
+5     &   0.86 &   0.80 & PO7 &   0.93 $\pm$ 0.01 \\
+6     &   0.89 &   0.96 & Cz  &   0.96 $\pm$ 0.01 \\
+7     &   0.89 &   0.78 & PO7 &   0.93 $\pm$ 0.01 \\
+8     &   0.92 &   0.91 & PO7 &   1.00 $\pm$ 0.00 \\
 \bottomrule
@@
 The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.
-The Performance curves (Fig.
-\ref{fig:performance}) show how the percentage of correctly identified letters depends on the number of repetitions that were used to obtain the averaged signal.
+There is a trade-off that needs to be balanced between the required number of repetitions for each trial to guarantee robust transmission and the the achieved speed of transmission affected by the repetions.
+As detailed in the table \ref{tab:singletrialreduction} this method is only reduced in a 20 percent in accuracy levels even by using only one sequence of repetitions.
 
-\begin{figure}[H]
+\begin{table}[H]
+\caption{Results obtained for Accuracy levels by performing a BCI simulation for the testing dataset.
+The BPC, best performing channel is shown and also performance levels for the 120, 60 and 12 repetitions.
+Finally the PR, performance reduction is described.
+Almost all the subjects performed around 70\% even with by averaging 2 signals for each class.}
 \centering
-\includegraphics[width=10cm]{performance.eps}
-\caption{Performance curves for the eight subjects included in the published dataset.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
-\label{fig:performance}
-\end{figure}
-
-We found that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that could allow the classification to reach a higher level.
-We hypothesize that as subjects may have different latencies and amplitudes of the P300 signal \citep{Riccio2013}, it may also be the case that the shape of the generated ERP may vary greatly in an intra-subject manner, thus the pattern could not be well generalized and a higher performance could not be reached (Fig.
-\ref{fig:p300templates}).
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=10cm]{subject8.png}
-\includegraphics[width=10cm]{subject3.png}
-\caption{Ten P300 template patches found for subjects 8 (up) and 3(down).
-In coincidence with the performance results, the P300 signature is more clear and consistent for subject 8 (higher performance) while for subject 3 (lower performance) the characteristic pattern is much more difficult to perceive.}
-\label{fig:p300templates}
-\end{figure}
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{cccccc}
+\toprule
+\textbf{Participants}	& \textbf{BPC}	& \textbf{ACC@120}	& \textbf{ACC@60}	&  \textbf{ACC@12} & \textbf{PR}\\
+\midrule
+1 & Cz &   0.80 &   0.82 &   0.72 & 9\% \\
+2 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
+3 & Cz &   0.80 &   0.82 &   0.72 & 9\%\\
+4 & PO8 &   0.93 &   0.78 &   0.65 & 29\%\\
+5 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
+6 & Cz &   0.80 &   0.82 &   0.72 & 9\%\\
+7 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
+8 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
+\bottomrule
+\end{tabular}
+\label{tab:singletrialreduction}
+\end{table}
 
@@
 %This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect 
-%
 %
@@
 %Channal identification based on the metric distance between the bags
-%
-%Signal Averaging.
 
@@
 
-A method to characterize and classify EEG signals where their characterization is transient in time-space, like the P300 ERP, has been presented.
+A method to characterize and classify EEG signals where the main characteristic can be both, rhythmic in nature as in motor imagery, and also transient in time space, like the P300 has been presented.
+
+Although single trial classification was not perfectly achieved, it has been shown that this method could be applied with a P300 Speller Matrix application with increased ITR.
 
@@
 
-The adaptive behaviour of the algorithm make it well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.
-This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e.
-\textit{how actually they looked like}) for the training and calibration step, and they do not require any prior knowledge about the signal.
-In contrast, when the shape of the ERP response is not consistent across the same subject, this method would not be able to find the templates and speller performance could be penalized.
+The adaptive behaviour of the algorithm make is well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.
+This is because, the descriptors are directly based on how the signals actually looked like for the training and calibration step, and they do not require any prior knowledge about the signal.
+This is of particular relevance for studies on outliers populations as may be the case for people who suffered some neuro-degenerative disease like Lou Gehrig's disease.
 
-At the same time, by analyzing the generated descriptors, which map in a very detailed and synthetic structure the shape information contained within the patch, a metric about the consistency of the shape of the generated P300 response could also be derived.
+The expanding of the understanding of this tool in order to automatically classify patterns in EEG, that are specifically identified by their shapes, is a prospect future work to be considered.
+It may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods \citep{Hartman2005}, frequent in sleep research.
 
-%It may be worthy of further consideration to evaluate if is there any correlation between the lower performance obtained for some subjects and their characterization of the ALS progression and degree.
-
-We believe that the expanding and the understanding of this tool in order to automatically classify those patterns in EEG that are specifically identified by their shapes (e.g.
-K-Complex, Vertex Waves, Positive Occipital Sharp Transient \citep{Hartman2005}) is a prospect future work to be considered.
-It may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods, frequent in sleep research.
-
-Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc}, i.e.
-to detect BCI illiteracy or to predict the achievable performance of a given method, or even as a tool for artefact removal (which is performed on many occasions by visually inspecting the signal).
+Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc} or as a tool for artefact removal (which is performed on many occasions by visually inspecting the signal).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{This projects is part of a the first author's PhD Thesis which is directed by Juan Miguel Santos and codirected by Ana Julia Villar.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictofinterests{The authors declare that there is no conflict of interest regarding the publication of this article.}
+\conflictofinterests{The authors declare no conflict of interest.}
 
@@
 
-\noindent EEG: Electroencephalography\\
+\noindent EEG: electroencephalography\\
 BCI: Brain Computer Interfaces\\
@@
 ERP: Event-Related Potential\\
-P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimulus\\
-ITR: Information Transfer Rate\\
-BTR: Bit Transfer Rate\\
-SIFT: Scale Invariant Feature Transform\\
-NBNN: Naive Bayes Nearest Neighbor\\
-HOG: Histogram Of Gradients}
+P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimuli\\
+ITR: Information Transfer Rate}
 

### 2017-09-17 coverletter.tex
Submitted version

Dear Editor,

This submitted article encompass the work that we have been doing and in the context of the setup of a Brain Computer Interfaces Laboratory at the Computer Engineering Department of the ITBA University in Buenos Aires Argentina.

EEG Processing and BCI are research areas in which we may be able to contribute as engineers and researchers in order to provide new methods and tools that could be used to improve the communication or treatment for persons affected by neuro-degenerative diseases.

We believe that this proposed method based on the histograms of gradients of signal plots, and most importantly, its successful application to identify ERP markers within EEG signals from ALS patients is relevant for this special issue and that it will contribute with a novel tool to analyse EEG signals.

Yours truly and deeply greatfull for your help
Rodrigo Ramele
PhD Candidate
ITBA University 
Argentina

### 2017-09-17 template.tex
Submitted version

--- 2017-09-17 articlebackup.tex
@@
 % Full title of the paper (Capitalized)
-\Title{EEG characterization and classification based on histograms of visual gradients}
+\Title{EEG characterization and classification based on the histograms of visual gradients of plots}
 
 % Authors, for the paper (add full first names)
-\Author{Rodrigo Ramele $^{1,\dagger}$, Ana Julia Villar $^{1}$ and Juan Miguel Santos  $^{1}$*}
+\Author{Rodrigo Ramele $^{1,\dagger,\ddagger}$, Ana Julia Villar $^{1,\ddagger}$ and Juan Miguel Santos  $^{1,\ddagger}$*}
 % Authors, for metadata in PDF
@@
 % Current address and/or shared authorship
-\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina}
+\firstnote{Current address: C1437FBH Lavarden 315, Postal Code C1437FBH, Ciudad Autónoma de Buenos Aires, Argentina} 
+\secondnote{These authors contributed equally to this work.}
 
@@
 % Abstract (Do not use inserted blank lines, i.e. \\) 
-\abstract{The analysis of Electroencephalograpy (EEG) signals for Amyotrophic Lateral Sclerosis (ALS) patients is of ulterior importance to elucidate different patterns or markers that could be potentially used to aid in diagnosis or treatment and to implement Brain Computer Interface (BCI) systems which may provide alternative pathways to transmit volitional information to enhance patient's quality of life.
-Of particular interests are those which are based on the recognition of Event-Related Potentials (ERP) which are shown to be more robust to the unavoidable physical impairment which comes with this motor neuron disease.
-This work mimics what electroencephalographers has been doing clinically, visually inspecting and categorizing phenomena within the EEG, and provides a framework to analyse and characterize EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.
-The validity of the method is shown by off-line processing a public dataset of ALS patients.}
+\abstract{A single paragraph of about 200 words maximum.
+For research articles, abstracts should give a pertinent overview of the work.
+We strongly encourage authors to use the following style of structured abstracts, but without headings: 1) Background: Place the question addressed in a broad context and highlight the purpose of the study; 2) Methods: Describe briefly the main methods or treatments applied; 3) Results: Summarize the article's main findings; and 4) Conclusion: Indicate the main conclusions or interpretations.
+The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.This is great.
+We had tremendous results.
+Our results are so great that nobody will believe that we actually managed to have them !}
 
 % Keywords
-\keyword{electroencephalography (EEG); BCI; P300; ALS; classification; HOG; SIFT}
+\keyword{EEG; BCI; P300; ALS; classification; hog; sift}
 
@@
 \setcounter{section}{-1} %% Remove this when starting to work on the template.
+\section{How to Use this Template}
+The template details the sections that can be used in a manuscript.
+Note that the order of article sections may differ from the requirements of the journal (e.g.
+for the positioning of the Materials and Methods section).
+Please check the instructions for authors page of the journal to verify the correct order.
+For any questions, please contact the editorial office of the journal or support@mdpi.com.
+For LaTeX related questions please contact Janine Daum at latex-support@mdpi.com.
+
+One of the most shoking reasons is that we can use the capacity of ML techniques to automatically explore the signal, by using these new featrues.
 
@@
 
-%\begin{enumerate}[leftmargin=*,labelsep=3mm]
-%\item	Why
-%\item	Purpose and significance
-%\item	Current state of the research field and key publications cited
-%\item   Controversial hypothesis
-%\item   Main aim and principal conclusions.
-%\end{enumerate}
-%
-%\begin{enumerate}[leftmargin=*,labelsep=3mm]
-%\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
-%\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
-%\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
-%\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
-%\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
-%\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
-%\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
-%\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
-%\end{enumerate}
+The introduction should briefly place the study in a broad context and highlight why it is important.
+It should define the purpose of the work and its significance.
+The current state of the research field should be reviewed carefully and key publications cited.
+Please highlight controversial and diverging hypotheses when necessary.
+Finally, briefly mention the main aim of the work and highlight the principal conclusions.
+As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
+Citing a journal paper \cite{ref-journal}.
+And now citing a book reference \cite{ref-book}.
 
-Although impressive advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} has diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundredth years old technology.
-Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, is pushing EEG to become the de-facto non invasive method to access and harness brain information.
+In this work EEG signals from ALS patients are offline processed with this radically different and horizontal approach.  This is the continuation of the work presented here and here.
 
-Key contribution to this development has been the field of Brain Computer Interfaces (BCI) \citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neuro-degenerative diseases.
+The same methodological algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300.
 
-One very noteworthy aspect of this communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS), to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
-The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.
+This method is based on a morphological analysis of the shape of the EEG signal.
 
-EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process and even considered random enough as to be used as a source of pseudo random number generator \citep{Chen2014}.
-Additionally, they have a high variability between different subjects and even between different moments for the same subject, requiring an adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.
-Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.
-
-EEG markers \citep{Clerc} that can be used to transmit volitionally information are limited, and each one of them has a particular set or combination of appropriate methods to decode them.
-Inevitably, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
-
-In this work a new method to characterize EEG signals is presented, expanded and detailed.
-Its validity is verified by processing off-line data for ALS patients.
-This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by the P300\citep{Knuth2006}.
-
-The method is based on the morphological analysis of the shape of the EEG signal\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signals.
-
-%This paper reports on a study done to address
-%the hypothesis that communication task complexity
-%(copy-spelling vs. novel text generation of a picture
-%description) is correlated with BCI performance accuracy
-%and P300 characteristics.
-
-In brief, this paper reports a straightforward method to, first characterize EEG signals based on the identification of their morphological shapes, and how this characterization can be used to implement a BCI scheme to identify Event Related Potentials, particularly the well-known P300, on an off-line and public dataset.
-
-\section{Materials and Methods}
-
-Besides new novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
-It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns \citep{Nijboer2009,Riener2014}.
-
-BCI has gained mainstream public awareness and recently even held a peculiar Olympics \citep{Riener2014} and been broadcasted during the inauguration of the last World Cup.
-New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments\citep{Huggins2016}.
-However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of remaining muscular movement \citep{Clerc}.
-
-To verify the validity of the proposed framework and method, the public dataset 008-2014 published on the BNCI-Horizon website \citep{Riccio2013} was used to perform a binary classification task on the provided signals.
-The algorithm was implemented using the VLFeat \citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).
-
-\subsection{Experimental protocol}
-
-The protocol is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
-The P300 task objective was meant to spell 7 words of 5 letter each.
-The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
-A trial, for this experiment, was defined as every attempt to select a letter from the speller, and it was composed of 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.
-Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.
-After 120 repetions and inter-trial pause was included before resuming with the following letter.
-
-The gathered dataset was sampled at 256 Hz and consisted of EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
-This dataset was published in the popular BNCI-Horizon 2020 website \citep{Brunner2014}.
-
-%\subsection{Morphological Signal Analysis}
-%
-%Several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
-%CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016,Yamaguchi2009}.
-%
-%Morphological or 
-%
-%template matching
-%
-%spike
-%sharp wave
-%spike wave
-%polyspike 
-%slow wave complex
-%
-%shape domain
-%
-%\subsection{P300}
-%
-%Uhhhh I can talk a lot about P300 \citep{Knuth2006}.
-%
-%P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.
-
-\subsection{Algorithm}
-
-\subsubsection{Preprocessing}
-
-The mandatory first step is to help to enhance the SNR of the P300 pattern above the level of basal EEG.
-The processing pipeline starts with applying a notch filter to the raw signal, a decimation with a FIR filter of order 30 from the original 256 Hz to 32 Hz and finally applying a 4th degree 10 Hz lowpass Butterworth filter.
-
-\subsubsection{Segmentation, Artefact Removal and Signal Averaging}
-
-The EEG matrix is processed on a channel by channel basis.
-For each word, each letter, a segment of EEG signal is extracted that encompass 10 repetitions of 12 flashings, 6 rows and 6 columns.
-For every 12 stimuli flashing, i.e.
-one complete sequence of repetitions of each row/column, a basic artefact removal procedure of removing the entire segment if any signal goes above 70 $\mu$V is applied.
-
-From every segment of 12 flashings, two epoch corresponding to a hit and two matching a no-hit are constructed as segments of 1s length (256 sample points).
-
-These segments, 20 for each class, are later point-to-point averaged for the whole 120 repetitions group.
-This finally lead to 2 averaged signals for each of the 5 letters, for the 7 words, comprising a set of 70 averaged signals\citep{Liang2008}.
-
-\subsubsection{Features}
-
-The central part of this method is the feature generation: the histogram of visual gradients.  To do so, the first step is the transformation of the signal into a temporary binary image.
-
-\begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot ( x(t,c) - \bar{x}(t,c)  )}\right \rfloor
-\end{equation}
-
-where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the EEG matrix defined for each $t$ and for a particular channel $c$.
-
-\begin{equation}
-I(z_1,z_2) = \left\{ \begin{array}{rl}
-255 & z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
-0   & \mbox{otherwise}
-\end{array}\right.
-\label{eq:images}
-\end{equation}
-
-From the Equation \ref{eq:images} the image $ I(z_1,z_2) $ is constructed an used to derive local representations that will capture the visual shape of the signal.
-
-Finally, the histogram of gradients, which is based on a very successful technique from Computer Vision, the SIFT \citep{Lowe2004} descriptor, is calculated on local image patches centered on sample points \citep{Vedaldi2010}.
-
-\begin{equation}
- h(p,i,j) = 3 \cdot s \int w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - T}{3 s}\right)\, |J(\mathbf{x})|\, d\mathbf{x}
-\label{eq:histogram}
-\end{equation}
-
-where $s$ is the size of the local patch, $\angle J(\mathbf{x})$ is the gradient found at each of the 16 blocks of the patch aligned with the angle $\theta_p$.
-The patch is divided in 16 blocks, arranged in a 4x4 grid and centered on $T = (x,y)$, a location within the image.
-On each block, the bin consists of 45 degrees obtaining 8 aligning angles.
-The Equation \ref{eq:histogram} accumulates the values of the image gradients aligned with each angle bin.
-This gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=16cm]{sampledescriptor.eps}
-\caption{The local patch is located around a sample point plotted on the temporary signal.
-All the sample points are interpolated using the Bresenham algorithm.
-The 4x4 grid can be seen on the descriptor and also arrows showing the dominant gradient on each block.}
-\label{fig:sampledescriptor}
-\end{figure}
-
-\subsubsection{Classification}
-
-The classification is carried out by using the discriminative semi-supervised classification method, the Naive Bayes Nearest Neighbout (NBNN) \citep{Boiman2008}.
-% algorithm which can be described as follows.
-
-\begin{equation}
-\hat{C} = argmin_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
-\end{equation}
-
-%BCI Simulation vs k-fold cross validation.
-It is quite known and understood that because the same experience will not generate the same markers in the EEG, BCI Simulation is sometimes better suited to understand how the device will operate on a real online application (the graeat variability of BCI).
+P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.
 
@@
 
-In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects is shown.
-The P300 characteristic curve can be seen particularly in subjects 2, and 6 and in a lesser extend in the remaining subjects.
-In order to obtain a valid binary classification on averaged signals, particular care was observed to avoid unbalanced samples, because that may introduce a bias in the classification procedure (signal variance is proportional to the square root of the number of samples and the procedure would be discriminating signals with different variances).Particularly in the case of the P300 response, the oddball paradigm requires that one of the stimulus need to be infrequent so that will unavoidably force that the data will be unbalanced (that is, an unequal number of trials in each condition).
+This section may be divided by subheadings.
+It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Subsection}
+
+\subsubsection{Subsubsection}
+
+Bulleted lists look like this:
+\begin{itemize}[leftmargin=*,labelsep=4mm]
+\item	First bullet
+\item	Second bullet
+\item	Third bullet
+\end{itemize}
+
+Numbered lists can be added as follows:
+\begin{enumerate}[leftmargin=*,labelsep=3mm]
+\item	First item
+\item	Second item
+\item	Third item
+\end{enumerate}
+
+The text continues here.
+
+\subsection{Figures, Tables and Schemes}
+
+All figures and tables should be cited in the main text as Figure 1, Table 1, etc.
 
@@
 \centering
-\includegraphics[width=16cm]{subjectaveraged.eps}
-\caption{Point-To-Point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
-The P300 characteristic curve can be well identified on subjects 2 and 6.}
-\label{fig:subjectaveraged}
+\includegraphics[width=9cm]{subjectaveraged.eps}
+\caption{This is a figure, Schemes follow the same formatting.
+If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
+(\textbf{b}) Description of what is contained in the second panel.
+Figures should be placed in the main text near to the first time they are cited.
+A caption on a single line should be centered.}
 \end{figure}
 
-Results are shown in Table \ref{tab:results} where BCI accuracy
-for the Cz channel is calculated and compared with the values obtained by the published of the dataset.
-Additionally the best performing channel is informed and its accuracy value.
-It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed very good performance indeed \citep{Huggins2016,Jure2016}.
-
-% (CITE TAYLOR AND  FRANCOS PAPER).
-
 \begin{table}[H]
-\caption{Results obtained for Accuracy levels by 3-fold cross validation.
-The values reported by the dataset publishers for Cz are reproduced here for comparison.
-Additionally, channel Cz BCI accuracies can be seen as well as the performance levels obtained for the BPC, the best performing channel for each subject.}
+\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
 \centering
 %% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccccc}
+\begin{tabular}{ccc}
 \toprule
-\textbf{Participants}	& \textbf{Original Cz}	& \textbf{ACC at Cz}	& \textbf{BPC}	& \textbf{Performance}\\
+\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
 \midrule
-1     &   0.84 &   0.79 & Cz  &   0.79 $\pm$ 0.01 \\
-2     &   0.86 &   0.81 & PO7 &   0.93 $\pm$ 0.01 \\
-3     &   0.87 &   0.78 & Cz  &   0.78 $\pm$ 0.03 \\
-4     &   0.86 &   0.68 & PO8 &   0.90 $\pm$ 0.01 \\
-5     &   0.86 &   0.80 & PO7 &   0.93 $\pm$ 0.01 \\
-6     &   0.89 &   0.96 & Cz  &   0.96 $\pm$ 0.01 \\
-7     &   0.89 &   0.78 & PO7 &   0.93 $\pm$ 0.01 \\
-8     &   0.92 &   0.91 & PO7 &   1.00 $\pm$ 0.00 \\
+entry 1		& data			& data\\
+entry 2		& data			& data\\
 \bottomrule
 \end{tabular}
-\label{tab:results}
 \end{table}
 
-The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.
-There is a trade-off that needs to be balanced between the required number of repetitions for each trial to guarantee robust transmission and the the achieved speed of transmission affected by the repetions.
-As detailed in the table \ref{tab:singletrialreduction} this method is only reduced in a 20 percent in accuracy levels even by using only one sequence of repetitions.
+\subsection{Formatting of Mathematical Components}
 
-\begin{table}[H]
-\caption{Results obtained for Accuracy levels by performing a BCI simulation for the testing dataset.
-The BPC, best performing channel is shown and also performance levels for the 120, 60 and 12 repetitions.
-Finally the PR, performance reduction is described.
-Almost all the subjects performed around 70\% even with by averaging 2 signals for each class.}
-\centering
-%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{cccccc}
-\toprule
-\textbf{Participants}	& \textbf{BPC}	& \textbf{ACC@120}	& \textbf{ACC@60}	&  \textbf{ACC@12} & \textbf{PR}\\
-\midrule
-1 & Cz &   0.80 &   0.82 &   0.72 & 9\% \\
-2 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
-3 & Cz &   0.80 &   0.82 &   0.72 & 9\%\\
-4 & PO8 &   0.93 &   0.78 &   0.65 & 29\%\\
-5 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
-6 & Cz &   0.80 &   0.82 &   0.72 & 9\%\\
-7 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
-8 & PO7 &   0.93 &   0.78 &   0.75 & 18\%\\
-\bottomrule
-\end{tabular}
-\label{tab:singletrialreduction}
-\end{table}
+This is an example of an equation:
+
+\begin{equation}
+\mathbb{S}
+\end{equation}
+
+%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
+This ensures correct linenumbering.
+The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
+Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
+%% Example of a theorem:
+\begin{Theorem}
+Example text of a theorem.
+\end{Theorem}
+
+The text continues here. Proofs must be formatted as follows:
+
+%% Example of a proof:
+\begin{proof}[Proof of Theorem 1]
+Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
+\end{proof}
+The text continues here.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\section{Discussion}
-%
-%This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect 
-%
-%is also based on how the image look like.
-%
-%SNR of p300 and how to detect it
-%
-%Check if you can use this to detect any kind of transient signal.
-%
-%Compare if it is possible with the descriptors from one subject, discriminate the others.
-%
-%Channal identification based on the metric distance between the bags
+\section{Discussion}
+
+Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses.
+The findings and their implications should be discussed in the broadest context possible.
+Future research directions may also be highlighted.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
+\section{Materials and Methods}
 
-A method to characterize and classify EEG signals where the main characteristic can be both, rhythmic in nature as in motor imagery, and also transient in time space, like the P300 has been presented.
+Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
+Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
+Please disclose at the submission stage any restrictions on the availability of materials or information.
+New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
 
-Although single trial classification was not perfectly achieved, it has been shown that this method could be applied with a P300 Speller Matrix application with increased ITR.
+Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers.
+If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
+They must be provided prior to publication.
 
-%Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).
-
-The adaptive behaviour of the algorithm make is well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.
-This is because, the descriptors are directly based on how the signals actually looked like for the training and calibration step, and they do not require any prior knowledge about the signal.
-This is of particular relevance for studies on outliers populations as may be the case for people who suffered some neuro-degenerative disease like Lou Gehrig's disease.
-
-The expanding of the understanding of this tool in order to automatically classify patterns in EEG, that are specifically identified by their shapes, is a prospect future work to be considered.
-It may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods \citep{Hartman2005}, frequent in sleep research.
-
-Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc} or as a tool for artefact removal (which is performed on many occasions by visually inspecting the signal).
+Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Subsection}
-%
-%\subsubsection{Subsubsection}
-%
-%Bulleted lists look like this:
-%\begin{itemize}[leftmargin=*,labelsep=4mm]
-%\item	First bullet
-%\item	Second bullet
-%\item	Third bullet
-%\end{itemize}
-%
-%Numbered lists can be added as follows:
-%\begin{enumerate}[leftmargin=*,labelsep=3mm]
-%\item	First item
-%\item	Second item
-%\item	Third item
-%\end{enumerate}
-%
-%The text continues here.
+\section{Conclusions}
+
+This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
 
@@
 %% optional
-%\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
+\supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.}
+\acknowledgments{All sources of funding of the study should be disclosed.
+Please clearly indicate grants that you have received in support of your research work.
+Clearly state if you received funds for covering the costs to publish in open access.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{This projects is part of a the first author's PhD Thesis which is directed by Juan Miguel Santos and codirected by Ana Julia Villar.}
+\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``X.X.
+and Y.Y.
+conceived and designed the experiments; X.X.
+performed the experiments; X.X.
+and Y.Y.
+analyzed the data; W.W.
+contributed reagents/materials/analysis tools; Y.Y.
+wrote the paper.'' Authorship must be limited to those who have contributed substantially to the work reported.}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictofinterests{The authors declare no conflict of interest.}
+\conflictofinterests{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
+Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
+If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.}
 
@@
 
-\noindent EEG: electroencephalography\\
-BCI: Brain Computer Interfaces\\
-SNR: Signal to Noise Ratio\\
-CNS: Central Nervous System\\
-ALS: Amyotrophic Lateral Sclerosis\\
-ERP: Event-Related Potential\\
-P300: Positive deflection of an Event-Related Potential which occurs 300 ms after onset of stimuli\\
-ITR: Information Transfer Rate}
+\noindent MDPI: Multidisciplinary Digital Publishing Institute\\
+DOAJ: Directory of open access journals\\
+TLA: Three letter acronym\\
+LD: linear dichroism}
 
@@
 %% optional
-%\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-%\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections.
+\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixsections{multiple} %Leave argument "multiple" if there are multiple sections.
 Then a counter is printed ("Appendix A?).
@@
 and no counter is printed (?Appendix?).
-%\appendix
-%\section{}
-%The appendix is an optional section that can contain details and data supplemental to the main text.
+\appendix
+\section{}
+The appendix is an optional section that can contain details and data supplemental to the main text.
 For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
 Mathemtaical proofs of results not central to the paper can be added as an appendix.
-%
-%\section{}
-%All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
+
+\section{}
+All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
 
@@
 %=====================================
-%\renewcommand\bibname{References}
-%\begin{thebibliography}{999}
+\renewcommand\bibname{References}
+\begin{thebibliography}{999}
 % Reference 1
-%\bibitem{ref-journal}
-%Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
+\bibitem{ref-journal}
+Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
 % Reference 2
-%\bibitem{ref-book}
-%Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
-%\end{thebibliography}
+\bibitem{ref-book}
+Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
+\end{thebibliography}
 
@@
 %=====================================
-\bibliography{article}
+%\bibliography{your_external_BibTeX_file}
 
@@
 %% optional
-%\sampleavailability{Samples of the compounds ...... are available from the authors.}
+\sampleavailability{Samples of the compounds ...... are available from the authors.}
 

### 2017-09-17 article.tex
Minor changes

@@
 
-This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained: section \ref{Preprocessing} and \ref{Pipeline} describe the processing pipeline.
+This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained.
+Section \ref{Preprocessing} and \ref{Pipeline} describe the processing pipeline.
 Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.

### 2017-09-19 coverletter.tex
Version sent by email

@@
 
-We believe that this proposed method based on the histograms of gradients of signal plots, and most importantly, its successful application to identify ERP markers within EEG signals from ALS patients is relevant for this special issue and that it will contribute with a novel tool to analyse EEG signals.
+We believe that this proposed method based on the histograms of gradients of signal plots, and most importantly, its  application to identify ERP markers within EEG signals relevant for this special issue and that it will contribute with a novel tool that can be used to enhance or improve HCI systems.
+
+We submitted previously on-time the
 

### 2017-11-24 article.tex
Version sent by email
Version submitted to Journal

@@
 % Keywords
-\keyword{electroencephalography (EEG); BCI; P300; ALS; classification; HOG; SIFT}
+\keyword{electroencephalography (EEG); BCI; P300; ALS; NBNN; HOG; SIFT}
 
@@
 
-This paper reports a method to, (1) characterize EEG signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from signal plots, and (2) how this characterization can be used to implement a BCI classification scheme to identify Event Related Potentials, particularly the well-known P300, on an offline and public dataset.
+This paper reports a method to, (1) characterize EEG signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) how this characterization can be used to implement a BCI classification scheme to identify Event Related Potentials, particularly the well-known P300, on an offline and public dataset.
 
@@
 The P300 \citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs 300 ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
-It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and though it is quite consistent across different subjects, it has a lower amplitude ( $ \pm 5 \mu V $ ) compared to basal EEG activity, reaching a SNR of around -15 db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity \citep{Hu2010}.
+It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and though it is quite consistent across different subjects, it has a lower amplitude ($\pm 5 \mu V $) compared to basal EEG activity, reaching a SNR of around -15 db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity \citep{Hu2010}.
 This signal can be used to implement a speller application by means of a Speller Matrix \citep{Farwell1988} (Fig.
@@
 \begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma_x(c) } }\right \rfloor
+\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor
 \end{equation}
@@
 \noindent where $\gamma$ is the image scale, $t$ is the time and $ x(t,c) $ is the point-to-point averaged EEG matrix defined for each $t$ and for a channel $c$.
-Lastly, $\bar{x}(c)$ and $ \sigma_x $ are the mean and standard deviation of $x$.
+Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
 
@@
 
-A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the 16 blocks and creating a histogram of the oriented gradients.
+A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the 16 blocks and creating a histogram of the such gradients.
 This technique is based on Lowe's SIFT \citep{Lowe2004} descriptor, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations.
@@
 For each one of the 6 query descriptors $ q $ for row/column, its $ k $ nearest neighbours in the template set $T$, $ NN_T^k(q) $ (acquired during the training phase)  are obtained and the distance between each one of them, $ x_i $, and the query descriptor $ q $ is summarized.
-The query descriptor $ q $ that minimizes this summation is the one that is chosen, first between the 6 rows, and then between the remaining 6 columns.
+The query descriptor $ q $ that minimizes this summation is the one that is chosen, first among the 6 rows, and then among the remaining 6 columns.
 By performing this procedure, the correct letter can be identified by matching the row $ r $ and column $ c $ from the P300 speller matrix.
@@
 In order to correctly decode the selected letter from each trial, particular care was observed to avoid unbalanced number of epochs  (i.e.
-an unequal number of epochs on each condition), because that may introduce a bias in the classification procedure (the variance of averaged signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).
+an unequal number of epochs on each condition), because that may introduce bias in the classification procedure (the variance of averaged signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).
 
@@
 From the 7 trials for each subject, the first 3 were used as training, and the remaining 4 for testing.
-Additionally the best performing channel is informed as well.
+The best performing channel is informed as well.
 It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
@@
 
-The adaptive behaviour of the algorithm make it well suited when the shape of the pattern elicited by the P300 response, does not conform to the predicted structure.
+The adaptive behaviour of the algorithm make it well suited when the shape of the pattern elicited by the P300 response does not conform to the predicted structure.
 This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e.
@@
 At the same time, by analyzing the generated descriptors, which map in a very detailed and synthetic structure the shape information contained within the patch, a metric about the consistency of the shape of the generated P300 response could also be derived.
-
-%It may be worthy of further consideration to evaluate if is there any correlation between the lower performance obtained for some subjects and their characterization of the ALS progression and degree.
+It may be worthy of further consideration to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
 

### 2017-11-24 cover.letter.tex
Version submitted to Journal

% cov
% Anpas

% \documentstyle[12pt,my-let-head-Nov-02]{letter}

\documentclass[11pt]{letter}

\usepackage{graphics}
%%%%%%%%%%%%%%%%%%%%% DEFINE USER-SPECIFIC MACROS BELOW %%%%%%%%%%%%%%%%%%%%%
\def\Who     {Rodrigo E. Ramele}
\def\What    {B.S. M.S.}
\def\Where   {Instituto Tecnológico de Buenos Aires}
\def\WhereWhere {Computer Science Department, CiC}
\def\Address {25 de Mayo 457, 6th floor}
\def\CityZip {Buenos Aires C.A.B.A.  C1002ABH}
\def\Email   {{\it E-mail:} rramele@itba.edu.ar}
\def\TEL     {{\it Phone:} (+54) (9-11)4193-9382}
\def\FAX     {}
\def\URL     {} % NOTE: use $\sim$
                % for tilde

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% MARGINS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textwidth       6.75in
\textheight      9.25in
\oddsidemargin   -.25in
\evensidemargin  -.25in
\topmargin       -.75in
\longindentation 0.50\textwidth
\parindent       6ex

%%%%%%%%%%%%%%%%%%%%%%%%%%% ADDRESS MACRO BELOW %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\address{ \hspace*{5.1in}
               \vskip -1.07in~\\
         \Large\hspace*{1.5in}  \hfill ~\vspace{.5ex}\\
               \hspace*{1.5in}  \hfill  \normalsize
               \makebox[0ex][r]{\bf  }\hspace{0.2cm} \vskip
               -6ex ~  \\
               \hspace*{0in}\hrulefill \\
               \hspace*{\fill}
               \setlength{\tabcolsep}{-3pt}
\hfill	\parbox[t]{2.95in}{ \footnotesize   \Who \\ \em
 				\WhereWhere \\
 				\Where  \\
				\Address \\
				\CityZip \\
				\TEL \\
				\Email
				} \hspace{-1.4in}
    \normalsize
	\vspace{-1.5in}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% OTHER MACROS BELOW %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\signature{\Who}

\def\opening#1{\ifx\@empty\fromaddress
 \thispagestyle{firstpage}
 \hspace*{\longindendation}\today\par
 \else \thispagestyle{empty}
 {\centering\fromaddress \vspace{1.9cm} \\*
 %\vfill   %%<---Comment out to put more on one page (shifts the date up
 \hspace*{0in}Summer School Application, March 2013\hspace*{\fill}\par}
 \fi
 %\vspace{1\parskip}
 {\raggedright \toname \\ \toaddress \par}\vspace{2\parskip}
 \noindent #1\par\raggedright\parindent 5ex\par
 %\def\thefootnote{}%
 %\def\footnoterule{\hrule}%
% \footnotetext{\hspace*{\fill}%
%              {\footnotesize\em An Equal Opportunity University}}
% \def\thefootnote{\arabic{footnote}}
 }

\long\def\closing#1{\par\nobreak\vspace{\parskip}
 \stopbreaks
 \noindent
 \ifx\@empty\fromaddress\else
 \hspace*{\longindentation}\fi
 \parbox{\indentedwidth}{\raggedright
 \ignorespaces #1\vskip .65in
 \ifx\@empty\fromsig
 \else \fromsig \fi\strut}
 \vspace*{\fill}
 \par}

% \def\footnoterule{\kern-1\p@\hrule width .4\columnwidth\kern .6\p@}

% \input{/u/disk9/maguest/math/.macros/psfig.tex}
% \input{/u/disk9/maguest/math/.macros/double.sty}

\begin{document}

\begin{letter}
{Berkeley University\\
CRCNS \\
Course in Mining and Modeling of Neuroscience Data
}

\opening{Dear Course Organizers:}

First of all, Thank you very much for your valuable time.  I would like to apply to the upcoming 2013 boreal summer course in Mining and Modeling of Neuroscience Data.

Allow me please to briefly introduce myself: I am currently a PhD student in Brain Computer Interfaces for  Neurorehabilitation Robotics in the "Instituto Tecnologico de Buenos Aires",  Buenos Aires, Argentina.  I have a B.S. in Computer Science and a Graduate Research Studentship in Bioengineering and Robotics in Japan. As Graduate Student, I took math courses in Fourier Analysis, Measure Theory and Machine Learning.  I use Matlab in daily basis, as well as other more statistical oriented tools like R or NumPy.  I also use frequently very popular BCI Matlab packages, like EEGLAB or BCILAB, and other like FASTICA (for Independent Component Analysis with Matlab).

As part of my PhD, I took courses in Biology and Bioengineering, but at the current stage of my theses, I need to advance in my understanding of new methods for processing neuroscience data.  As long as I am aware of, this course is precisely oriented in that direction, and I found it very exciting to be able to participate as a student.

Two years ago I had the chance to be part of another summer school in Spain.  I have to say that the experience was the cornerstone of the current work that we are doing at the lab where I belong.  We need to test our ideas against the work that others are doing in similar areas, and summer schools are one of the best ways to do that while we learn practical techniques for our daily work.

Let me tell you just a few words about our project:  we are working on developing Brain Computer Interface for assisting elderly to maintain or regain walking ability. The system is currently based on EPOC Emotiv as an EEG sensor that will allow us to drive a custom Walking Helper Device.  We are aiming to develop a Non Invasive, Passive and Online solution and we are currently improving our processing methods to better solve the complex problem of the real-time processing of the noisy EEG signal.

Attached to this application, I have included my curriculum vita.  You can find there that I also have experience working in Software Companies that I put into practice when it is required (specially, when time is short and my tutor needs something for yesterday, you know what I mean).  I will be pleased to provide you with any additional information and certificates that you may require.  Please, do not hesitate at all to contact me.  I will be tremendous grateful for any kind of feedback that you can provide me.

Finally, I thank you for your consideration and look forward to hearing from you.

\closing{Sincerely,}
\end{letter}
\end{document}

### 2017-11-27 response.letter.tex
First set of changes for the Resubmission

\documentclass[journal,onecolumn,12pt]{IEEEtran}

\usepackage{amsmath,amssymb,bm}
\usepackage{amsthm, amsfonts}	
\usepackage{bm,bbm}
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{fancybox}
\usepackage{url,booktabs}

\usepackage{xr}

\title{Reply to Reviewer's Comments on\\
``EEG characterization and classification based on image gradient histograms''\\
(now ``Oriented Gradient Histogram applied to P300 Detection'')}
\author{}

\begin{document}

\maketitle
\pagenumbering{roman}
\setcounter{page}{1}

We are grateful to the reviewers for pointing at relevant issues in our manuscript.

In the following, we discuss how we dealt with each raised issue.

\vskip+1ex
\noindent \dotfill

\section*{\fbox{Reviewer \#1:}}

I have read the manuscript "EEG characterization and classification based on Histogram of Gradients of Signal Plots" with great interest since its title promised a solution to an important problem. In spite of presenting an interesting methodology and some interesting issues related to EEG characterization, I consider the manuscript is not ready for publication.

First of all, the title is misleading since a first, and naive, read suggest being able to classify and characterize any EEG with their different components (i.e., P200, P300, P600, N100, N200, N400 to name a few). Clearly, the title does not match the content since the authors focus on the P300; of course, there is evidence of the benefits of working with these signals, but, is the proposed method extensible to other signals?, this is not clear from the manuscript.

Another issue with the manuscript is its mathematical vagueness. For example, the variable T is used as a vector or two-tuple but its notation is different from the notation the authors use for vectors. Moreover, there is something wrong in equation (4) and something inconsistent between equations (7) and (8). Also, the authors go back and forth the digital space and the continuous space and although there is an equivalence, the manuscript could benefit from a clearer explanation.

Another issue that I find worrisome is the design of experiments. First of all, it seems that the number of subjects is rather low, it would be interesting to know why the authors decided to use such a subset. Also, the authors speculate that their results are not that good because they did not take into account the latency and magnitude of the signals, there is previous evidence in the literature that the P300 suffers from these, and other, problems and, thus, one wonders why the authors did not anticipate such problems.

A couple of finale observations. First, it seems the authors assume that the readers have some previous deep knowledge of some topics such as that some channels should carry more or better information without a previous reference. Another issue has to do with the way the manuscript is written, sometimes is very difficult to follow the ideas the authors want to convey, an example can be found in the paragraph containing equation (5), the manuscript surely benefit from a careful review before submitting.

\subsection*{\ovalbox{General Comments}}

2.1) the title is misleading since a first, and naive, read suggest being able to classify and characterize any EEG with their different components (i.e., P200, P300, P600, N100, N200, N400 to name a few). Clearly, the title does not match the content since the authors focus on the P300

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation. You're very clever.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

2.2) these signals, but, is the proposed method extensible to other signals?

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
This comment is very important, thank you very much. 
An explanation  of which is a satisfactory result for Algorithm~\ref{Algoritmo} was added in sections~\ref{errorevaluation} and~\ref{sec:EdgeSimulatedData}.  
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex
\subsection*{\ovalbox{Third Paragraph}}

3.1) Another issue with the manuscript is its mathematical vagueness. For example, the variable T is used as a vector or two-tuple but its notation is different from the notation the authors use for vectors

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation.  You are so clever that it is very likely that you are God.}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

3.2) Moreover, there is something wrong in equation (4)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Oh my God, I can't believe how smart you are !!! You probably are rich by now.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

3.3) and something inconsistent between equations (7) and (8)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
We have revised the manuscript and made a clear distinction between ``methods'' and ``features''.
We thank you for noticing this, that led to a less ambiguous text.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

3.4) Also, the authors go back and forth the digital space and the continuous space and although there is an equivalence, the manuscript could benefit from a clearer explanation.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation.  Smart people like you should be president.
Following your suggestion, we added graphical evidence of the difficulty of separating targets characterized by $\mathcal{G}_I^0(-2,1,1)$, $\mathcal{G}_I^0(-3,1,1)$, $\mathcal{G}_I^0(-4,1,1)$, and $\mathcal{G}_I^0(-5,1,1)$ data; cf.\ Fig.~\ref{fig:DifficultProblem}.
With this, and the corresponding discussion, we believe we have addressed your point.
 }
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

\subsection*{\ovalbox{Fourth Paragraph}}

4.1) Another issue that I find worrisome is the design of experiments.\\

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation. 
We now specify details of the E-SAR image used in the applications to true data; cf. Sec.~\ref{sec:ActualData}.
Following your suggestion, we added  a comparison with the Triangular distance in actual data. 
We applied the method in edge detection with both distances and we discovered that both of them resulted in exactly the same edge point. 
Then, we concluded that these distances are very similar in terms of accuracy in finding the transition point, but the geodesic distance is better in terms of  computation time. 
We fully addressed your point.
 }
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

4.2) it seems that the number of subjects is rather low, it would be interesting to know why the authors decided to use such a subset.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for your suggestion. We increased the discussion and conclusions, and enhanced lines for future research.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

4.3) Also, the authors speculate that their results are not that good because they did not take into account the latency and magnitude of the signals, there is previous evidence in the literature that the P300 suffers from these, and other, problems and, thus, one wonders why the authors did not anticipate such problems.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Oh God. You are clearly an oracle. 
All these points were modified according to the suggestions, but instead of doing what you asked for we do what we want instead.
We hope this is acceptable.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

\subsection*{\ovalbox{Fifth Paragraph}}

5.1) it seems the authors assume that the readers have some previous deep knowledge of some topics such as that some channels should carry more or better information without a previous reference.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Oh God. You are clearly an oracle. 
All these points were modified according to the suggestions, but instead of doing what you asked for we do what we want instead.
We hope this is acceptable.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

5.2)  issue has to do with the way the manuscript is written

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Oh God. You are clearly an oracle. 
All these points were modified according to the suggestions, but instead of doing what you asked for we do what we want instead.
We hope this is acceptable.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

5.3) sometimes is very difficult to follow the ideas the authors want to convey, an example can be found in the paragraph containing equation (5)

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Oh God. You are clearly an oracle.

We hope this is acceptable.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

\end{document}

### 2017-11-27 article.tex
First set of changes for the Resubmission

@@
 % Full title of the paper (Capitalized)
-\Title{EEG characterization and classification based on Histogram of Gradients of Signal Plots}
+\Title{Oriented Gradient Histogram applied to P300 Detection}
 
@@
 Moreover, EEG markers \citep{Clerc} that can be used to volitionally transmit information are limited, and each one of them has a particular combination of appropriate methods to decode them.
-Inevitably, it is necessary to implement many distinct and specialized algorithmic methods, to filter the signal, enhance its SNR, and try to determine some meaning out of it.
+Inevitably, it is necessary to implement  distinct and specialized algorithmic methods, to filter the signal, enhance its Signal to Noise Ratio (SNR), and try to determine some meaning out of it.
 
@@
 
-The P300 \citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs 300 ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
+The P300 \citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around 300 ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
 It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and though it is quite consistent across different subjects, it has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a SNR of around -15 db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity \citep{Hu2010}.
-This signal can be used to implement a speller application by means of a Speller Matrix \citep{Farwell1988} (Fig.
-\ref{fig:p300matrix}) where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
+This signal can be used to implement a speller application by means of a Speller Matrix \citep{Farwell1988} as shown in Fig.
+\ref{fig:p300matrix}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
 Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
@@
 The processing pipeline starts by applying a notch filter to the raw signal, a 
-4th degree 10 Hz lowpass Butterworth filter and finally a decimation with an FIR filter of order 30 from the original 256 Hz down to 16 Hz \citep{Krusienski2006}.
+4th degree 10 Hz lowpass Butterworth filter and finally a decimation with an Finite Impulse Response (FIR) filter of order 30 from the original sampling frequency down to 16 Hz \citep{Krusienski2006}.
 
@@
 Two of these segments are labeled as \textit{hit}, whereas the remaining 10 are labeled as \textit{no hit}.
-A hit represents that the EEG segment should contain the ERP signature time-locked to the flashing stimulus.
+A hit represents that the EEG segment should contain the P300 ERP signature time-locked to the flashing stimulus.
 \item \textbf{Signal Averaging}:  The P300 ERP is deeply buried under background EEG so the traditional approach to identify it is by point-to-point averaging the time-locked stacked signal segments.
@@
 
-The underlying idea of this method is to generate a template for the signal,  based on the image plot.  Hence, the first step is its transformation into a temporary binary image.
+The underlying idea of this method is to generate a template for the signal \citep{Ramele2016},  based on the image plot.  Hence, the first step is its transformation into a temporary binary image.
 
@@
 
-In order to complete the plot from the pixels, the Bresenham \citep{Ramele2016} algorithm is used to interpolate straight lines between each pair of these consecutive pixels.
+In order to complete the plot from the pixels, the Bresenham \citep{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of these consecutive pixels.
 
@@
 
-On the generated image, a keypoint is placed on a precise location along the signal trace over the image plot (Fig.
-\ref{fig:sampledescriptor}).
-Around that keypoint, a local image patch is constructed as follows:  it is first divided in 16 blocks of size $ 3 \cdot s $ each, arranged in a 4x4 grid and centered on $T = (x,y)$.
+On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_kp, y_kp)$ over the image plot.
+In Figure  \ref{fig:sampledescriptor}.
+Around the keypoint, a local image patch is constructed by dividing it in 16 blocks of size $ 3 \cdot s $ where $ s is the size of the local patch $, arranged in a 4x4 grid and centered on $ \mathbf{kp} = (x_kp,y_kp)$, the patch center.
 
-A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the 16 blocks and creating a histogram of the such gradients.
+A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the 16 blocks and creating a histogram of such gradients.
 This technique is based on Lowe's SIFT \citep{Lowe2004} descriptor, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations.
@@
 \begin{equation}
- h(p,i,j) = 3 \cdot s \int w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - T}{3 s}\right)\, |J(\mathbf{x})|\, d\mathbf{x}\, d\theta_p
+ h(p,i,j) = 3 \cdot s \sum w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - \mathbf{kp_c}}{3 s}\right)\, |J(\mathbf{x})|
 \label{eq:histogram}
@@
 On the other hand $ \textbf{x} = (x_i, x_j) $ corresponds to each one of the pixels of the local patch, and $ w_\mathrm{ang}  $  and $ w_{ij} $ are linear interpolation functions \citep{Lowe2004,Vedaldi2010}.
-Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each grid block when the size of the patch is the unity.
+Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each grid block when the size of the patch is one (i.e.
+$s = 1$).
 As the patch is 4x4, it gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.
@@
 \centering
-\includegraphics[width=6cm]{sampledescriptor.png}
-\includegraphics[width=9cm]{histogramsbines.png}
-\caption{(Left) The local patch is located on a precise location within the generated plot.
+\includegraphics[width=5cm]{descriptorpatchsample.png}
+\includegraphics[width=6cm]{histogramsbines.png}
+%\includegraphics[width=9cm]{gradientszoom.png}
+\caption{(Left) AGREGAR UN GRAFICO QUE MUESTRE MEJOR EL HISTOGRAMA Y PONER UNO DONDE EL P300 CON SU DESCRIPTOR SE VEAN PERFECTO.The local patch is located on a precise location within the generated plot.
 The 4x4 block grid can be seen on the patch.
@@
 The experimental protocol used to generate this dataset is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
-The P300 detection task designed for this experiment consisted of spelling 7 words (runs) of 5 letter each, using the traditional P300 Speller Matrix \citep{Farwell1988} where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
+The P300 detection task designed for this experiment consisted of spelling 7 words (runs) of 5 letters each, using the traditional P300 Speller Matrix \citep{Farwell1988} where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
 The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
@@
 \begin{equation}
-s = \frac{\Delta \mu V}{4 \cdot 3} \cdot \gamma 
+s= \frac{\Delta \mu V}{4 \cdot 3} \cdot \gamma 
 \label{eq:mapping1}
@@
 
-%\begin{figure}[H]
-%\centering
-%\includegraphics[width=16cm]{localpatch.png}
-%\caption{The local patch is mapped to the original image by their correspondence .}
-%\label{fig:sampledescriptor}
-%\end{figure}
+\begin{figure}[H]
+\centering
+\includegraphics[width=10cm]{localpatch.png}
+\caption{The local patch captures the signal information based on both the vertical and horizontal scale.}
+\label{fig:sampledescriptor2}
+\end{figure}
 

### 2017-11-27 article.tex
cambios menores

@@
 
-On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_kp, y_kp)$ over the image plot.
-In Figure  \ref{fig:sampledescriptor}.
-Around the keypoint, a local image patch is constructed by dividing it in 16 blocks of size $ 3 \cdot s $ where $ s is the size of the local patch $, arranged in a 4x4 grid and centered on $ \mathbf{kp} = (x_kp,y_kp)$, the patch center.
+On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot.
+Figure~\ref{fig:sampledescriptor} shows a patch centered in the keypoint and.
+Around the keypoint, a local image patch is constructed by dividing it in 16 blocks of size $ 3 \cdot s $ where $ s$ is the size of the local patch , arranged in a $4 \times 4$ grid and centered on $ \mathbf{kp} = (x_{kp},y_{kp})$, the patch center.
 
@@
 \centering
-\includegraphics[width=5cm]{descriptorpatchsample.png}
-\includegraphics[width=6cm]{histogramsbines.png}
+\subfigure{[The local patch]\includegraphics[width=5cm]{descriptorpatchsample.png}\label{patch}}
+\subfigure[The local patch]{\includegraphics[width=6cm]{histogramsbines.png}\label{histogrambines}}
 %\includegraphics[width=9cm]{gradientszoom.png}

### 2017-11-28 article.tex
Ultimas Modificaciones

@@
 \documentclass[entropy,article,submit,moreauthors,pdftex,10pt,a4paper]{mdpi} 
+\usepackage{subfigure}
 
@@
 
-The P300 \citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around 300 ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
-It is produced under the oddball paradigm \cite{WolpawJonathanR2012} and though it is quite consistent across different subjects, it has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a SNR of around -15 db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity \citep{Hu2010}.
-This signal can be used to implement a speller application by means of a Speller Matrix \citep{Farwell1988} as shown in Fig.
-\ref{fig:p300matrix}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
+The P300~\citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around $300$ ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
+It is produced under the oddball paradigm~\cite{WolpawJonathanR2012} and though it is consistent across different subjects.
+It has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a SNR of around $-15$ db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity~\citep{Hu2010}.
+This signal can be used to implement a speller application by means of a Speller Matrix~\citep{Farwell1988} as shown in Fig.~\ref{fig:p300matrix}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
 Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
@@
 \includegraphics[width=10cm]{openvibep300matrix.png}
-\caption{The P300 Matrix\citep{Farwell1988} used in the OpenVibe Open Source software \citep{Renard2010}.
-Rows are labeled from 1 to 6 and columns from 7 to 12.
+\caption{The P300 Matrix used in the OpenVibe Open Source software~\citep{Renard2010}.
+Rows are labeled from $1$ to $6$ and columns from $7$ to $12$.
 Rows/columns flash intermittently in random permutations.}
@@
 The processing pipeline starts by applying a notch filter to the raw signal, a 
-4th degree 10 Hz lowpass Butterworth filter and finally a decimation with an Finite Impulse Response (FIR) filter of order 30 from the original sampling frequency down to 16 Hz \citep{Krusienski2006}.
+$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with an Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.
 
@@
 For every 12 flashing stimuli, i.e.
-one complete sequence of intensification of each of the 6 rows plus the 6 columns, a basic artefact elimination procedure is implemented by removing the entire segment when any signal deviates above/bellow $ \pm 70 \mu V $.
-\item \textbf{Segmentation}: For each of the 12 stimuli,  a window of 1 second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
-Two of these segments are labeled as \textit{hit}, whereas the remaining 10 are labeled as \textit{no hit}.
+one complete sequence of intensification of each of the $6$ rows plus the $6$ columns, a basic artefact elimination procedure is implemented by removing the entire segment when any signal deviates above/bellow $ \pm 70 \mu V $.
+\item \textbf{Segmentation}: For each of the $12$ stimuli,  a window of $1$ second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
+Two of these segments are labeled as \textit{hit}, whereas the remaining $10$ are labeled as \textit{no hit}.
 A hit represents that the EEG segment should contain the P300 ERP signature time-locked to the flashing stimulus.
@@
 On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot.
-Figure~\ref{fig:sampledescriptor} shows a patch centered in the keypoint and.
-Around the keypoint, a local image patch is constructed by dividing it in 16 blocks of size $ 3 \cdot s $ where $ s$ is the size of the local patch , arranged in a $4 \times 4$ grid and centered on $ \mathbf{kp} = (x_{kp},y_{kp})$, the patch center.
+Around the keypoint, a local image patch is constructed by dividing it in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+It is arranged in a $4 \times 4$ grid and $ \mathbf{kp} = (x_{kp},y_{kp})$ is the patch center.
 
-A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the 16 blocks and creating a histogram of such gradients.
-This technique is based on Lowe's SIFT \citep{Lowe2004} descriptor, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations.
-In order to calculate this histogram, the interval $ [0-360]$ of possible angles is divided in 8 bins, each one at 45 degrees.
-
-The values of the image gradients for the 8 angle bins $ p = {0, 45, 90, 135, 180, 225, 270, 315} $ and for each grid block bin $ i,j = {1,2,3,4} $  are accumulated in the 3-dimensional histogram $ h $ through
+A local representation of the shape of the signal within the patch can be described by obtaining the gradient orientations on each of the $16$ blocks and creating a histogram of gradients.
+This technique is based on Lowe's SIFT~\citep{Lowe2004} method, and it is biomimetically inspired in how the visual cortex detects shapes by analyzing orientations.
+In order to calculate the histogram, the interval $[0-360]$ of possible angles is divided in $8$ bins, each one at $45$ degrees.
+Then, the angles $\theta$  are $ \theta = \{0, 45, 90, 135, 180, 225, 270, 315\} $.
+For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation: %In each step, the histogram is computed by multiplying the length of the gradient vector by a Wang trilinear interpolation between the gradient angle and the bin, in order to give importance to the module of the gradient.
 
 \begin{equation}
- h(p,i,j) = 3 \cdot s \sum w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta_p)\, w_{ij}\left(\frac{\mathbf{x} - \mathbf{kp_c}}{3 s}\right)\, |J(\mathbf{x})|
+ h(\theta,i,j) = 3s \sum_{\mathbf{x}} w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta)\, w_{ij}\left(\frac{\mathbf{x} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{x})|
 \label{eq:histogram}
@@
 
-\noindent where $s$ is the size of the local patch, which can be converted into pixels by doing $ Px = 4 \cdot 3 \cdot s $ ,  $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each one of the 16 blocks of the patch, whereas $\angle J(\mathbf{x}) $ is the angle of the gradient vector and  $\theta_p$ is the angle bin.
-On the other hand $ \textbf{x} = (x_i, x_j) $ corresponds to each one of the pixels of the local patch, and $ w_\mathrm{ang}  $  and $ w_{ij} $ are linear interpolation functions \citep{Lowe2004,Vedaldi2010}.
-Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each grid block when the size of the patch is one (i.e.
-$s = 1$).
-As the patch is 4x4, it gives a descriptor of 128 dimension as shown on Figure \ref{fig:sampledescriptor}.
+\noindent  where $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
+in~\citep{Lowe2004,Vedaldi2010}.
+Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.
+As the patch is of size $4 \times 4$ and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained.
+In each step, the histogram is computed by multiplying by the length of the gradient vector, in order to take account of it.
+
+Fig.~\ref{fig:sampledescriptor} shows an example of a patch and a scheme of the histogram computation.
+Fig.~\ref{patch} is a plot of the signal and the patch centered in the keypoint.
+In Fig.~\ref{histogrambines} the orientations on each patch are illustrtaed which encompass the corresponding $\mathbf{kp}$-descriptor of $128$ coordinates.
+The first two bloks orientations are shown.
 
@@
 \centering
-\subfigure{[The local patch]\includegraphics[width=5cm]{descriptorpatchsample.png}\label{patch}}
-\subfigure[The local patch]{\includegraphics[width=6cm]{histogramsbines.png}\label{histogrambines}}
+\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.5\linewidth]{DiagramaConkeypoint.png}\label{patch}}
+\subfigure[Orientations on each patch.]{\includegraphics[width=.5\linewidth]{histogramsbines.png}\label{histogrambines}}
 %\includegraphics[width=9cm]{gradientszoom.png}
-\caption{(Left) AGREGAR UN GRAFICO QUE MUESTRE MEJOR EL HISTOGRAMA Y PONER UNO DONDE EL P300 CON SU DESCRIPTOR SE VEAN PERFECTO.The local patch is located on a precise location within the generated plot.
-The 4x4 block grid can be seen on the patch.
-(Right) On each grid block, eight bins are used to describe the eight different orientations, and they are later arranged in a 128-dimension vector (4x4x8).}
+\caption{ Example of a patch and a scheme of the orientations histogram computation.}
 \label{fig:sampledescriptor}

### 2017-11-29 article.tex
Mas cambios en las fórmulas

@@
 
-On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot.
-Around the keypoint, a local image patch is constructed by dividing it in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
-It is arranged in a $4 \times 4$ grid and $ \mathbf{kp} = (x_{kp},y_{kp})$ is the patch center.
+On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered.
+A local image patch is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center.
 
@@
 Then, the angles $\theta$  are $ \theta = \{0, 45, 90, 135, 180, 225, 270, 315\} $.
-For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation: %In each step, the histogram is computed by multiplying the length of the gradient vector by a Wang trilinear interpolation between the gradient angle and the bin, in order to give importance to the module of the gradient.
+For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block $B_{i,j}$,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation: %In each step, the histogram is computed by multiplying the length of the gradient vector by a Wang trilinear interpolation between the gradient angle and the bin, in order to give importance to the module of the gradient.
 
 \begin{equation}
- h(\theta,i,j) = 3s \sum_{\mathbf{x}} w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta)\, w_{ij}\left(\frac{\mathbf{x} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{x})|
+ h(\theta,i,j) = 3s \sum_{\mathbf{x} \in B_{i,j}} w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta)\, w_{ij}\left(\frac{\mathbf{x} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{x})|
 \label{eq:histogram}
@@
 
-\noindent  where $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
+\noindent  where $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
 in~\citep{Lowe2004,Vedaldi2010}.
@@
 As the patch is of size $4 \times 4$ and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained.
-In each step, the histogram is computed by multiplying by the length of the gradient vector, in order to take account of it.
+It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $ , so the method considers both, the magnitude and the orientation of the gradient vector.
 
@@
 Fig.~\ref{patch} is a plot of the signal and the patch centered in the keypoint.
-In Fig.~\ref{histogrambines} the orientations on each patch are illustrtaed which encompass the corresponding $\mathbf{kp}$-descriptor of $128$ coordinates.
-The first two bloks orientations are shown.
+In Fig.~\ref{histogrambines} the possible orientations on each patch are illustrated.
+They form the corresponding $\mathbf{kp}$-descriptor of $128$ coordinates.
+The first two blocks orientations are shown.
 
@@
 \centering
-\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.5\linewidth]{DiagramaConkeypoint.png}\label{patch}}
-\subfigure[Orientations on each patch.]{\includegraphics[width=.5\linewidth]{histogramsbines.png}\label{histogrambines}}
+\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.49\linewidth]{DiagramaConkeypoint.png}\label{patch}}
+\subfigure[Orientations on each patch.]{\includegraphics[width=.49\linewidth]{histogramsbines.png}\label{histogrambines}}
 %\includegraphics[width=9cm]{gradientszoom.png}
@@
 
-While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the image that were used to derive those descriptors.
+While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the whole image.
+%that were used to derive those descriptors.
 
-The NBNN algorithm tackles this problem by comparing each image against a whole classification class which is characterized by the set of descriptors that are closest to each one of the descriptors obtained for the query image.
-The algorithm is based on
+The NBNN algorithm tackles this problem by comparing each image against a whole classification class.
+It is characterized by the set of descriptors closest to each one of the descriptors obtained for the query image.
+The algorithm  find the predicted class $\hat{C}$ of a query image by means of the following minimization:
 
@@
 \end{equation}
-
-\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.
+where $C$ is the a class, $ d_i $ is a descriptor that belongs to the query image and $ NN_C(d_i) $ is the corresponding near neighbor descriptor for the class C.
+%\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.
 

### 2017-11-30 article.tex
Bulleteado de la Clasificacion

@@
 
+Steps
+
+\begin{itemize}
+\item For a Reactive Synchronous speller, like the P300 Speller, we need to identify the selected letter 1 out of 36 (like the matrix).
+\item Rows are 1-6 and columns are 7-12
+\item The matrix is randomly highlighted, rows and columns, making a permutation 1-12.
+\item So, each 12 stimulations there are one row and one column that matches the selected letter by the user and that it should generate the P300 waveform.
+\item Thus, we need to identify the segment that generated the p300 response.
+\item Descriptors are obtained from images of signal plots according to previously described algorithm.
+\item Descriptors thus belong to two classes: hits vs no-hits
+\item During training phase, descriptors from labeled segments are extracted.  These are the templates and they are grouped in a template set $ T $.
+\item On decoding phase, a new query image is generated from the signal plot of a segment of single-channel EEG (in the same way as for training phase).
+From this image a new \textit{query} descriptor is extracted.
+\item so we need to identify the row, and the column.  There are 6 rows and 6 columns for a given letter.
+\item the letter that the person was focusing on is i,j, corresponding to one out of the 6 rows, and one out of the 6 columns.
+\item how we can do this ?  Well we generate 12 images, 6 for rows, and 6 for columns
+\item we pick the first 6 rows, and get one descriptor for each.
+\item we have 6 descriptors and we build 6 classes.
+\item we compute \ref{eq:classification} and we pick the class that minimize it.
+\item we thus have now the row.
+\item repeat for column.
+\item we can now identify the selected letter. 
+\end{itemize}
+
 %The first step consists in extracting labeled descriptors from the training set for the  which are thus grouped together in two KD-tree \citep{Vedaldi2010} database structures.
@@
 \end{figure}
+
+\subsubsection{P300 for healthy subjects}
+
+We replicated the same experiment on healthy subjects using g.Tec g.Nautilus 8 channels, bla bla bla bla in order to validate
+the obtained results.
 

### 2017-11-30 article.tex
Nueva ecuacion para la clasificacion

--- 2017-11-29 article.tex
@@
 
-\noindent  where $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
+\noindent  where $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch using finite differences, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
 in~\citep{Lowe2004,Vedaldi2010}.
@@
 The first two blocks orientations are shown.
+Following this procedure for every assigned keypoint, we obtain $N_{kp}$ descriptors.
 
@@
 The classification is carried out by using a discriminative semi-supervised classification method, based on the Naive Bayes Nearest Neighbor (NBNN) \citep{Boiman2008}.
+
+The aim is to identify the selected letter from the matrix.
+Previously, during training phase, descriptors from labeled segments are extracted.
+The P300 templates are grouped in a template set called $ T $.
+Rows are labeled 1-6, whereas columns are labeled 7-12.
+The process has the following steps:
+
+\begin{enumerate}
+\item Highlight randomly the rows and columns from the matrix.  There are one row and one column that matches the selected letter by the user.
+\item For $ k= {1-6} $: Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{Feature}, $ d^r_1, \dots  d^r_6 $ for rows and  $ d^c_1, \dots  d^c_6 $
+\item For $ k= {7-12} $: Compute
+
+\begin{equation}
+\hat{r} = \arg \min_u \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2,\; u = 1,\dots,6
+\label{eq:multiclassification}
+\end{equation}
+
+\noindent This procedure is repeated for the 6 columns.
+
+\begin{equation}
+\hat{c} = \arg \min_u \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2,\; u = 7,\dots,12
+\label{eq:multiclassification}
+\end{equation}
+
+\end{enumerate}
 

### 2017-12-04 article.tex
Otras modificaciones

@@
 It has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a SNR of around $-15$ db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity~\citep{Hu2010}.
-This signal can be used to implement a speller application by means of a Speller Matrix~\citep{Farwell1988} as shown in Fig.~\ref{fig:p300matrix}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
+This signal can be used to implement a speller application by means of a Speller Matrix~\citep{Farwell1988}.
+Fig.~\ref{fig:p300matrix} shows an example of a Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
 Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
@@
 \centering
-\includegraphics[width=10cm]{openvibep300matrix.png}
-\caption{The P300 Matrix used in the OpenVibe Open Source software~\citep{Renard2010}.
-Rows are labeled from $1$ to $6$ and columns from $7$ to $12$.
-Rows/columns flash intermittently in random permutations.}
+\includegraphics[width=.49\linewidth]{openvibep300matrix.png}
+\caption{Example of a Speller Matrix.  Rows and columns flash intermittently in random permutations.}
 \label{fig:p300matrix}
@@
 
-In order to complete the plot from the pixels, the Bresenham \citep{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of these consecutive pixels.
+In order to complete the plot from the pixels, the Bresenham \citep{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of  consecutive pixels.
 
@@
 
-\subsubsection{Feature Extraction: Histogram of Oriented Gradients} \label{SIFT}
+\subsubsection{Feature Extraction: Histogram of Oriented Gradients}
+\label{SIFT}
 
@@
 In order to calculate the histogram, the interval $[0-360]$ of possible angles is divided in $8$ bins, each one at $45$ degrees.
-Then, the angles $\theta$  are $ \theta = \{0, 45, 90, 135, 180, 225, 270, 315\} $.
  For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block $B_{i,j}$,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation: %In each step, the histogram is computed by multiplying the length of the gradient vector by a Wang trilinear interpolation between the gradient angle and the bin, in order to give importance to the module of the gradient.
@@
 
-\noindent  where $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch using finite differences, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
+\noindent  where $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $, $ i,j \in \{1,2,3,4\} $, $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch using finite differences, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
 in~\citep{Lowe2004,Vedaldi2010}.
@@
 
-\subsubsection{Classification} \label{Classification}
+\subsubsection{Classification}
+\label{Classification}
 
-The classification is carried out by using a discriminative semi-supervised classification method, based on the Naive Bayes Nearest Neighbor (NBNN) \citep{Boiman2008}.
+The following step is classify the descriptors $\{d_{kp}, kp =1,\dots, 12\}$ in order to determine the correct row and column of the chosen letter en the Speller Matrix.
+The classification is carried out by using a discriminative semi-supervised classification method, based on the Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008}.
 
@@
 \item Highlight randomly the rows and columns from the matrix.  There are one row and one column that matches the selected letter by the user.
-\item For $ k= {1-6} $: Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{Feature}, $ d^r_1, \dots  d^r_6 $ for rows and  $ d^c_1, \dots  d^c_6 $
-\item For $ k= {7-12} $: Compute
+\item For $ k= {1-6} $: Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_1, \dots,  d^c_6 $.
+\item Compute
 
@@
 
-\noindent This procedure is repeated for the 6 columns.
+\noindent This procedure is repeated for the columns.
 
@@
 \end{equation}
+where $q$ is a posible descriptor, $NN_T(d)$ is the set of descriptors nearest to $d$ and $\hat{r}$ and$$
 

### 2017-12-04 article.tex
Pequeños ajustes

--- 2017-11-30 article.tex
@@
 
-The signal is first scaled and standardized (i.e. z-score) by
+Averaged signal segments obtained from the preprocessing stage are first scaled and standardized (i.e. z-score) by
 
@@
 
-\noindent where $\gamma$ is the image scale, $t$ is the time and $ x(t,c) $ is the point-to-point averaged EEG matrix defined for each $t$ and for a channel $c$.
+\noindent where $\gamma$ is the image scale, $t$ is the time and $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for  channel $c$.
 Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
@@
 In order to calculate the histogram, the interval $[0-360]$ of possible angles is divided in $8$ bins, each one at $45$ degrees.
-Then, the angles $\theta$  are $ \theta = \{0, 45, 90, 135, 180, 225, 270, 315\} $.
+Thus, angles $\theta$  are $ \theta = \{0, 45, 90, 135, 180, 225, 270, 315\} $.
 For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block $B_{i,j}$,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation: %In each step, the histogram is computed by multiplying the length of the gradient vector by a Wang trilinear interpolation between the gradient angle and the bin, in order to give importance to the module of the gradient.
@@
 As the patch is of size $4 \times 4$ and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained.
-It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $ , so the method considers both, the magnitude and the orientation of the gradient vector.
+It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $, so the method considers both, the magnitude and the orientation of the gradient vector.
 
@@
 They form the corresponding $\mathbf{kp}$-descriptor of $128$ coordinates.
-The first two blocks orientations are shown.
+The first two blocks are shown.
 Following this procedure for every assigned keypoint, we obtain $N_{kp}$ descriptors.
@@
 %\includegraphics[width=9cm]{gradientszoom.png}
-\caption{ Example of a patch and a scheme of the orientations histogram computation.}
+\caption{ Example of a patch and a scheme of the orientation's histogram computation.}
 \label{fig:sampledescriptor}
@@
 Previously, during training phase, descriptors from labeled segments are extracted.
-The P300 templates are grouped in a template set called $ T $.
+These descriptors are the P300 templates which are grouped in a template set called $ T $.
 Rows are labeled 1-6, whereas columns are labeled 7-12.
@@
 \begin{enumerate}
-\item Highlight randomly the rows and columns from the matrix.  There are one row and one column that matches the selected letter by the user.
-\item For $ k= {1-6} $: Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{Feature}, $ d^r_1, \dots  d^r_6 $ for rows and  $ d^c_1, \dots  d^c_6 $
-\item For $ k= {7-12} $: Compute
+\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the selected letter by the subject.
+\item Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{Feature}, $ d^r_1, \dots  d^r_6 $ for rows and  $ d^c_7, \dots  d^c_{12} $ for columns.
+\item Compute
 
@@
 \hat{r} = \arg \min_u \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2,\; u = 1,\dots,6
-\label{eq:multiclassification}
+\label{eq:multiclassificationrow}
 \end{equation}
 
-\noindent This procedure is repeated for the 6 columns.
+\noindent for rows, and  repeat for the 6 columns:
 
@@
 \hat{c} = \arg \min_u \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2,\; u = 7,\dots,12
-\label{eq:multiclassification}
+\label{eq:multiclassificationcol}
 \end{equation}
 
+Each $q$ that belongs to the $k$ nearest neighbors  $NN_T(d^c_u)$ is used to calculate the distances to the rows descriptors and the columns descriptors, and by minimizing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
+
+%The first step consists in extracting labeled \textit{hit} descriptors from the training set which are thus grouped together in a KD-tree \citep{Vedaldi2010} template set $ T $.
+On decoding stage, 12 new images from the signal segments are generated, and their descriptors extracted.
+The 12 descriptors are divided in rows and columns.
+For each one of the 6 query descriptors $ q $ for row/column, its $ k $ nearest neighbours in the template set $T$, $ NN_T^k(q) $ (acquired during the training phase)  are obtained and the distance between each one of them, $ x_i $, and the query descriptor $ q $ is summarized.
+The query descriptor $ q $ that minimizes this summation is the one that is chosen, first among the 6 rows, and then among the remaining 6 columns.
+By performing this procedure, the correct letter can be identified by matching the row $ r $ and column $ c $ from the P300 speller matrix.
+
 \end{enumerate}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=10cm]{classification.png}
+\caption{The classification algorithm identifies the corresponding row and column .}
+\label{fig:classification}
+\end{figure}
+
+Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent so that will unavoidably force the data to be unbalanced \citep{Tibon2015}.
+The NBNN method suffers from biased classification on unbalanced classes \citep{Fornoni2014}.
+By reversing the roles of the query and the class in equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.
 
@@
 
-While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the whole image.
+%While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the whole image.
 %that were used to derive those descriptors.
-
-The NBNN algorithm tackles this problem by comparing each image against a whole classification class.
+%
+%The NBNN algorithm tackles this problem by comparing each image against a whole classification class.
 It is characterized by the set of descriptors closest to each one of the descriptors obtained for the query image.
 The algorithm  find the predicted class $\hat{C}$ of a query image by means of the following minimization:
-
-\begin{equation}
-\hat{C} = \arg \min_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
-\label{eq:classification}
-\end{equation}
-where $C$ is the a class, $ d_i $ is a descriptor that belongs to the query image and $ NN_C(d_i) $ is the corresponding near neighbor descriptor for the class C.
-%\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.
-
-This basic method was adapted to the P300 Speller and the classifier works according to the following modified equation:
-
-\begin{equation}
-\hat{Q}_{r\,c} = \arg \min_q \sum_{x_i \epsilon NN_T^k(q)}^{} \left\lVert x_i - q \right\rVert ^2
-\label{eq:multiclassification}
-\end{equation}
-
-The first step consists in extracting labeled \textit{hit} descriptors from the training set which are thus grouped together in a KD-tree \citep{Vedaldi2010} template set $ T $.
+%
+%\begin{equation}
+%\hat{C} = \arg \min_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
+%\label{eq:classification}
+%\end{equation}
+%where $C$ is the a class, $ d_i $ is a descriptor that belongs to the query image and $ NN_C(d_i) $ is the corresponding near neighbor descriptor for the class C.
+%%\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.
+%
+%This basic method was adapted to the P300 Speller and the classifier works according to the following modified equation:
+%
+%\begin{equation}
+%\hat{Q}_{r\,c} = \arg \min_q \sum_{x_i \epsilon NN_T^k(q)}^{} \left\lVert x_i - q \right\rVert ^2
+%\label{eq:multiclassification}
+%\end{equation}
+%
+%The first step consists in extracting labeled \textit{hit} descriptors from the training set which are thus grouped together in a KD-tree \citep{Vedaldi2010} template set $ T $.
 On decoding stage, 12 new images from the signal segments are generated, and their descriptors extracted.
@@
 By performing this procedure, the correct letter can be identified by matching the row $ r $ and column $ c $ from the P300 speller matrix.
-
-Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent so that will unavoidably force the data to be unbalanced \citep{Tibon2015}.
+%
+%Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent so that will unavoidably force the data to be unbalanced \citep{Tibon2015}.
 The NBNN method suffers from biased classification on unbalanced classes \citep{Fornoni2014}.
@@
 
-We replicated the same experiment on healthy subjects using g.Tec g.Nautilus 8 channels, bla bla bla bla in order to validate
-the obtained results.
+We replicated the same experiment on healthy subjects using g.Tec g.Nautilus 8 channels.
+The experiment protocol is exactly the same and we 20 subjects, age $ 30$, 5 females and 15 males, 6 right handed and the remaining left handed.
+
+Results are shown in Figure 9191.
 

### 2017-12-04 article.tex
pequeño cambio

--- 2017-12-04 article.tex
@@
 In order to calculate the histogram, the interval $[0-360]$ of possible angles is divided in $8$ bins, each one at $45$ degrees.
- For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block $B_{i,j}$,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation: %In each step, the histogram is computed by multiplying the length of the gradient vector by a Wang trilinear interpolation between the gradient angle and the bin, in order to give importance to the module of the gradient.
+
+ For each spacial bin $ i,j = \{1,2,3,4\} $, corresponding to the indexes of each block $B_{i,j}$,  the orientations are accumulated in a  $3$-dimensional histogram $h$ through the following equation:
 
@@
 \begin{enumerate}
-\item Highlight randomly the rows and columns from the matrix.  There are one row and one column that matches the selected letter by the user.
-\item For $ k= {1-6} $: Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_1, \dots,  d^c_6 $.
+\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the selected letter by the subject.
+\item Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns. 
 \item Compute
@@
 \hat{r} = \arg \min_u \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2,\; u = 1,\dots,6
-\label{eq:multiclassification}
+\label{eq:multiclassificationrow}
 \end{equation}
@@
 \hat{c} = \arg \min_u \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2,\; u = 7,\dots,12
-\label{eq:multiclassification}
-\end{equation}
-where $q$ is a posible descriptor, $NN_T(d)$ is the set of descriptors nearest to $d$ and $\hat{r}$ and$$
-
-\end{enumerate}
-
-%The first step consists in extracting labeled descriptors from the training set for the  which are thus grouped together in two KD-tree \citep{Vedaldi2010} database structures.
-and group those marked as \textit{hit} in a template set $ T $
-
-%BCI Simulation vs k-fold cross validation.
-It is quite known and understood that because the same experience will not generate the same markers in the EEG, BCI Simulation is sometimes better suited to understand how the device will operate on a real online application (the graeat variability of BCI).
-
-While doing classification based on local features, which encode partial information, one problem that frequently arises is how to ensemble the individual classification of each local characteristic to the whole image.
-%that were used to derive those descriptors.
-
-The NBNN algorithm tackles this problem by comparing each image against a whole classification class.
-It is characterized by the set of descriptors closest to each one of the descriptors obtained for the query image.
-The algorithm  find the predicted class $\hat{C}$ of a query image by means of the following minimization:
-
-\begin{equation}
-\hat{C} = \arg \min_C \sum_{}^{} \left\lVert d_i - NN_C(d_i) \right\rVert ^2
-\label{eq:classification}
-\end{equation}
-where $C$ is the a class, $ d_i $ is a descriptor that belongs to the query image and $ NN_C(d_i) $ is the corresponding near neighbor descriptor for the class C.
-%\noindent where the predicted class $\hat{C}$ of a query image is calculated as the class $ C $ that minimize the summation of the $ L^2 $ distance between each descriptor $ d_i $ that belongs to the query image and its corresponding near neighbour $ NN_C(d_i) $ descriptor for class C.
-
-This basic method was adapted to the P300 Speller and the classifier works according to the following modified equation:
-
-\begin{equation}
-\hat{Q}_{r\,c} = \arg \min_q \sum_{x_i \epsilon NN_T^k(q)}^{} \left\lVert x_i - q \right\rVert ^2
-\label{eq:multiclassification}
+\label{eq:multiclassificationcol}
 \end{equation}
 
-The first step consists in extracting labeled \textit{hit} descriptors from the training set which are thus grouped together in a KD-tree \citep{Vedaldi2010} template set $ T $.
+where $q$ is a template descriptor that belongs to the $k$ nearest neighbors  $NN_T(d^c_u)$.
+By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
+
+%The first step consists in extracting labeled \textit{hit} descriptors from the training set which are thus grouped together in a KD-tree \citep{Vedaldi2010} template set $ T $.
 On decoding stage, 12 new images from the signal segments are generated, and their descriptors extracted.
@@
 
+\end{enumerate}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=10cm]{classification.png}
+\caption{For each letter, single trial segments are averaged for the 6 rows and 6 columns.}
+\label{fig:classification}
+\end{figure}
+
 Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent so that will unavoidably force the data to be unbalanced \citep{Tibon2015}.
-The NBNN method suffers from biased classification on unbalanced classes \citep{Fornoni2014}.
-By reversing the roles of the query and the class in Eq.
-\ref{eq:multiclassification}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.
-
-%Additionally,
-
-%El tema del balanceo lo mencionas y marcas lo cual está muy bien, pero no decí cómo lo solucionaste. Si llegaras a tener un poco de tiempo podrías mencionarlo.
+At the same time, the NBNN method suffers from biased classification on unbalanced classes \citep{Fornoni2014}.
+By reversing the roles of the query and the class in equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.
+Figure \ref{fig:classification} schematize this basic algorithm where the classification is only performed against the template set T, for each one of the twelve classes, thus avoiding the unbalanced situation (2 hit vs.
+10 no-hit).
 

### 2017-12-04 response.letter.tex
Carta de respuesta

@@
 \subsection*{\ovalbox{General Comments}}
+I have read the manuscript "EEG characterization and classification based on Histogram of Gradients of Signal Plots" with great interest since its title promised a solution to an important problem.
+In spite of presenting an interesting methodology and some interesting issues related to EEG characterization, I consider the manuscript is not ready for publication.
+\vskip+1ex
+\noindent \dotfill
+\vskip+1ex
+%
+\begin{quotation}
+{\color{blue}
+Thank you for your feedback. We have modified several points according to your recommendations, we hope the article is now better suited for publication. 
+}
+\end{quotation}
+%
+\vskip+1ex
+\noindent \dotfill
+\vskip+1ex
 

### 2017-12-04 response.letter.tex
Cambios en la Intro y la carta

@@
 {\color{blue}
-Thank you for the observation. You're very clever.
+Thank you for the observation. The title of the article was change accordingly to your suggestion.  
 }
@@
 
-2.2) these signals, but, is the proposed method extensible to other signals?
+2.2) there is evidence of the benefits of working with these signals P300, but, is the proposed method extensible to other signals?, this is not clear from the manuscript.
 

### 2017-12-04 article.tex
Cambios en la Intro y la carta

@@
 New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments \citep{Huggins2016}.
-However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement \citep{Clerc}.
+However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement~\citep{Clerc}.
 
-In this work, a new method to characterize and classify EEG signals is presented, expanded and detailed.
+In~\citep{Ramele2016}, the authors introduces a method for classification of rhythmic patterns using gradient histograms orientations.
+Inspired in that work, we proposed another application of the developed method for classify and describe transient events like those produced by P300 Event Related Potential.
  Its validity is verified by processing offline data for ALS patients.  
-This is the continuation of the work previously presented in \citep{Ramele2016}, where it was applied to rhythmic patterns, and that it can be extended to describe transient events like those produced by P300 Event Related Potential \citep{Knuth2006}.
+The proposed method is based on the morphological analysis of the shape of the EEG signal~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and it was built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots~\citep{Hartman2005}.
 
-The method is based on the morphological analysis of the shape of the EEG signal \citep{Alvarado-Gonzalez2016,Yamaguchi2009} and was inspired by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots \citep{Hartman2005}.
-
-This paper reports a method to, (1) characterize EEG signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) how this characterization can be used to implement a BCI classification scheme to identify Event Related Potentials, particularly the well-known P300, on an offline and public dataset.
+This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) describe the way in which  this classification can be used to implement a BCI Event Related Potentials identification, on an offline and public dataset and another own dataset.
 
@@
 \begin{equation}
-\hat{r} = \arg \min_u \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2,\; u = 1,\dots,6
+\hat{r} = \arg \min_{u \in \{1,\dots,6\}} \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2
 \label{eq:multiclassificationrow}
@@
 \begin{equation}
-\hat{c} = \arg \min_u \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2,\; u = 7,\dots,12
+\hat{c} = \arg \min_{u \in \{7,\dots,12\}} \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2
 \label{eq:multiclassificationcol}
@@
 
-where $q$ is a template descriptor that belongs to the $k$ nearest neighbors  $NN_T(d^c_u)$.
+where $NN_T(d^l_u), \;l=r,c$ is the set of the $k$ nearest neighbors to $d^l_u, \;l=r,c$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l=r,c$.
 By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.

### 2017-12-05 article.tex
Agregué la sección del nuevo dataset. Falta refinarla

@@
 
-In~\citep{Ramele2016}, the authors introduces a method for classification of rhythmic patterns using gradient histograms orientations.
-Inspired in that work, we proposed another application of the developed method for classify and describe transient events like those produced by P300 Event Related Potential.
+In~\citep{Ramele2016},  authors introduce a method for classification of rhythmic patterns using gradient histograms orientations.
+Inspired in that work, we proposed a novel application of the developed method to classify and describe transient events like those produced by P300 Event Related Potential.
  Its validity is verified by processing offline data for ALS patients.  
-The proposed method is based on the morphological analysis of the shape of the EEG signal~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and it was built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots~\citep{Hartman2005}.
+The proposed approach is based on the morphological analysis of the shape of the EEG signal~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and it was built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots~\citep{Hartman2005}.
 
-This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) describe the way in which  this classification can be used to implement a BCI Event Related Potentials identification, on an offline and public dataset and another own dataset.
+This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) describe the way in which this classification can be used to implement an offline P300-based BCI Speller application using two public datasets.
 
@@
 The P300~\citep{Farwell1988,Knuth2006} is a positive deflection of the EEG signal which occurs around $300$ ms after the onset of a rare and deviant stimulus that the subject is expected to attend.
-It is produced under the oddball paradigm~\cite{WolpawJonathanR2012} and though it is consistent across different subjects.
+It is produced under the oddball paradigm~\cite{WolpawJonathanR2012} and it is consistent across different subjects.
 It has a lower amplitude  ($\pm 5 \mu V $) compared to basal EEG activity, reaching a SNR of around $-15$ db estimated based on the amplitude of the P300 response signal divided by the standard deviation of the background EEG activity~\citep{Hu2010}.
@@
 The processing pipeline starts by applying a notch filter to the raw signal, a 
-$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with an Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.
+$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.
 
@@
 For every 12 flashing stimuli, i.e.
-one complete sequence of intensification of each of the $6$ rows plus the $6$ columns, a basic artefact elimination procedure is implemented by removing the entire segment when any signal deviates above/bellow $ \pm 70 \mu V $.
+one complete sequence of intensification of each of the $6$ rows plus the $6$ columns, a basic artefact elimination procedure is implemented by removing the entire sequence when any signal deviates above/bellow $ \pm 70 \mu V $.
 \item \textbf{Segmentation}: For each of the $12$ stimuli,  a window of $1$ second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
@@
 
-The underlying idea of this method is to generate a template for the signal \citep{Ramele2016},  based on the image plot.  Hence, the first step is its transformation into a temporary binary image.
+The underlying idea of this method is to generate a template for the signal \citep{Ramele2016},  based on the image plot.
+Hence, the first step is the signal transformation into a temporary binary image.
 
@@
 
-\noindent where $\gamma$ is the image scale, $t$ is the time and $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for  channel $c$.
+\noindent where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for  channel $c$.
 Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
@@
 
-The following step is classify the descriptors $\{d_{kp}, kp =1,\dots, 12\}$ in order to determine the correct row and column of the chosen letter en the Speller Matrix.
+The following step is classify the descriptors $\{d_{kp}, kp =1,\dots, 12\}$ in order to determine the correct row and column of the chosen letter on the Speller Matrix.
 The classification is carried out by using a discriminative semi-supervised classification method, based on the Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008}.
@@
 \begin{enumerate}
-\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the selected letter by the subject.
+\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
 \item Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns. 
@@
 \centering
-\includegraphics[width=10cm]{classification.png}
+\includegraphics[width=15cm]{classificationgraph.pdf}
 \caption{For each letter, single trial segments are averaged for the 6 rows and 6 columns.}
@@
 
-Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent so that will unavoidably force the data to be unbalanced \citep{Tibon2015}.
+Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent.
+Hence this will unavoidably force the data to be unbalanced \citep{Tibon2015}.
 At the same time, the NBNN method suffers from biased classification on unbalanced classes \citep{Fornoni2014}.
@@
 By using $ s = 3 $ and a quadruple scale of the image $ \gamma = 4 $ this gives the local patch, and the descriptor, the ability to identify events of 9 $ \mu V $ of amplitude, a resolution of $ 1 $ Pixel $ = \frac{1}{4} \mu V $ and span   of $ \lambda = 0.56 s $ (because the patch is a geometric square, $ s $ must be the same to map both the height and the required span of the signal).
-Finally, descriptor locations $ T $ were selected at $ x =  0.55 s \cdot Fs \cdot \gamma = 35 $ and $ y = z(c) $ (Eq.
+Finally, descriptor locations $\mathbf{kp}$  were selected at $ x =  0.55 s \cdot Fs \cdot \gamma = 35 $ and $ y = z(c) $ (Eq.
 \ref{eq:zerolevel}).
@@
 
-Results are shown in Figure 9191.
+Participants were recruited voluntarily and the experiment were conducted anonymyzing information in accordance with the declaration of Helsinki.
+No monetary compensation was delivered and all participants agreed and signed a written informed consent.
+All healthy participants (n=13, 7 males, 6 females, average age 22.5 years (SD = 1,8 years), range 21–26 years) had normal or corrected-to normal vision and no history of neuropsychiatric dis- orders.
+
+EEG data were recorded from the same 8 electrodes positions (C3, Cz, C4, CPz, P3, Pz, P4, POz) with all the systems.
+The reference positions were placed at the right ear.
+The ground positions was placed at the left ear, except with the g.Nautilus system (placed at AFz, because this could not be changed).
+Sampling rate was set at 250 Hz, the closest possible to predefined g.Mobilab sampling rate (256 Hz).
+Data were acquired notch filtered at 50 Hz and passband filtered between 2 Hz and 30 Hz.
+The filtering was performed via the proprietary Simulink HighSpeed Online Processing block modules for g.MOBIlab and g.Nautilus systems, and via MATLAB code for the Xpress system.
+Table 1 summarizes the acquisition settings for each system.
+
+8 gel-based active electrodes (g.LADYbird) + g.LADYbird (GND) + g.GAMMAearclip (REF) C3, Cz, C4, CPz, P3, Pz, P4, POz, GND: AFz, REF: right ear
 

### 2017-12-06 article.tex
Cambios y ajustes en el detalle del metodo
Algunos cambios, corté la figura de la columna para que quede igual que la otra y puse comentarios en el tex que hay que corregir.

@@
 \centering
-\includegraphics[width=.49\linewidth]{openvibep300matrix.png}
+\subfigure[Row flash.]{\includegraphics[width=.29\linewidth]{openvibep300matrix.png}}
+\subfigure[Column flash.]{\includegraphics[width=.29\linewidth]{p300matrixcoloriginal.png}}
 \caption{Example of a Speller Matrix.  Rows and columns flash intermittently in random permutations.}
@@
 \subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.49\linewidth]{DiagramaConkeypoint.png}\label{patch}}
-\subfigure[Orientations on each patch.]{\includegraphics[width=.49\linewidth]{histogramsbines.png}\label{histogrambines}}
+\subfigure[Orientations on two blocks of the patch.]{\includegraphics[width=.49\linewidth]{histogramsbines.png}\label{histogrambines}}
 %\includegraphics[width=9cm]{gradientszoom.png}
@@
 
-\noindent This procedure is repeated for the columns.
+\noindent and
 
@@
 By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
+The parameter $k$ is an input parameter of the algorithm.
 
@@
 
+%Este párrafo no se entiende nada
+%------------------
+Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli needs to be infrequent.
+Hence this will unavoidably force the data to be unbalanced~\citep{Tibon2015}.
+At the same time, the NBNN method suffers from biased classification on unbalanced classes~\citep{Fornoni2014}.
+By reversing the roles of the query and the class in equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.
+Figure~\ref{fig:classification} shows a schema of this idea,  where the classification is only performed against the template set T, for each one of the twelve classes, thus avoiding the unbalanced situation (2 hit vs.
+10 no-hit).
 \begin{figure}[H]
@@
 \end{figure}
-
-Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli need to be infrequent.
-Hence this will unavoidably force the data to be unbalanced \citep{Tibon2015}.
-At the same time, the NBNN method suffers from biased classification on unbalanced classes \citep{Fornoni2014}.
-By reversing the roles of the query and the class in equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.
-Figure \ref{fig:classification} schematize this basic algorithm where the classification is only performed against the template set T, for each one of the twelve classes, thus avoiding the unbalanced situation (2 hit vs.
-10 no-hit).
-
+%------------------
 \subsection{Experimental Protocol} \label{Protocol}
 
-To verify the validity of the proposed framework and method, the public dataset 008-2014  \citep{Riccio2013} published on the BNCI-Horizon website \citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, was used to perform an offline BCI Simulation to decode the spelled words from the provided signals.
-The algorithm was implemented using  VLFeat  \citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).
+To verify the validity of the proposed framework and method, the public dataset 008-2014~\citep{Riccio2013} published on the BNCI-Horizon website~\citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, was used to perform an offline BCI Simulation to decode the spelled words from the provided signals.
+The algorithm was implemented using  VLFeat~\citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).
 
@@
 
-The experimental protocol used to generate this dataset is explained in \citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
-The P300 detection task designed for this experiment consisted of spelling 7 words (runs) of 5 letters each, using the traditional P300 Speller Matrix \citep{Farwell1988} where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
-The first 3 runs were used for training and the remaining 4 for testing with visual feedback.
-A trial, as defined by the BCI2000 platform \citep{Schalk2004}, was every attempt to select a letter from the speller, and it was composed of signal segments corresponding to 10 repetitions of flashes of 6 rows and 6 columns of the traditional 6x6 P300 matrix , yielding 120 repetitions.
-Flashing of row/columns was performed for 0.125 s, following by a resting period (i.e.
+The experimental protocol used to generate this dataset is explained in~\citep{Riccio2013} but can be summarized as follows:  8 subjects with confirmed diagnoses but on different stages of ALS disease, were recruited and accepted to perform the experiments.
+The P300 detection task designed for this experiment consisted of spelling 7 words of 5 letters each, using the traditional P300 Speller Matrix~\citep{Farwell1988}.
+The flashing of rows and columns provide the deviant stimulus required to elicit this physiological response.
+The first 3 words are used for training and the remaining 4 words, for testing with visual feedback.
+A trial, as defined by the BCI2000 platform~\citep{Schalk2004}, is every attempt to select a letter from the speller.
+It is composed of signal segments corresponding to 10 repetitions of flashes of 6 rows and 10 repetitions of flashes of 6 columns of the matrix, yielding 120 repetitions.
+Flashing of a row or a column is performed for 0.125 s, following by a resting period (i.e.
 Inter-stimulus interval) of the same length.
-After 120 repetitions an inter-trial pause was included before resuming with the following letter.
+After 120 repetitions an inter-trial pause is included before resuming with the following letter.
 
@@
 
-In order to asses and verify the identification of the P300 response, subjects are instructed to perform a copy-spelling task where they have to fix their attention to successive letters in order to copy a previously determined set of words (in contrast to an online free-running operation of the speller where each users decides on its own what letter to choose).
+%Este párrafo no se entiende nada y repite in order
+In order to asses and verify the identification of the P300 response, subjects are instructed to perform a copy-spelling task.
+They have to fix their attention to successive letters in order to copy a previously determined set of words (in contrast to an online free-running operation of the speller where each users decides on its own what letter to choose).
 
@@
 
-Parameters were selected according to the experimental protocol.
+Parameters are selected according to the experimental protocol.
 As the P300 event latency and amplitude vary greatly between subjects, it is necessary to provide a patch that will be able to capture an entire transient event.
-Equations \ref{eq:mapping1} and \ref{eq:mapping2} can be used to map the original signal parameters to local image patch structure.
-
+Equations~\ref{eq:mapping1} and~\ref{eq:mapping2} can be used to map the original signal parameters to local image patch structure.
+%¿Qué es s? hay que hacer acordar. 
 \begin{equation}
@@
 \end{equation}
-
-\noindent where $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz), $ \lambda $ is the length in seconds covered by the patch, and $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
+%falta definir gamma
+\noindent where $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and $ \lambda $ is the length in seconds covered by the patch.
+% estos son los parámetros que vos usaste en tu experimento? 
 By using $ s = 3 $ and a quadruple scale of the image $ \gamma = 4 $ this gives the local patch, and the descriptor, the ability to identify events of 9 $ \mu V $ of amplitude, a resolution of $ 1 $ Pixel $ = \frac{1}{4} \mu V $ and span   of $ \lambda = 0.56 s $ (because the patch is a geometric square, $ s $ must be the same to map both the height and the required span of the signal).
@@
 \ref{eq:zerolevel}).
+%demasiado larga la frase, no se puede leer.
 
+% esta figura no está comentada en el texto.
 \begin{figure}[H]

### 2017-12-07 article.tex
cambié la figura del keypoint
últimas modificaciones

@@
 \centering
-\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.49\linewidth]{DiagramaConkeypoint.png}\label{patch}}
+\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.49\linewidth]{signalpatchkeypoint.png}\label{patch}}
 \subfigure[Orientations on two blocks of the patch.]{\includegraphics[width=.49\linewidth]{histogramsbines.png}\label{histogrambines}}
@@
 
-\subsubsection{Classification}
+\subsubsection{Speller Matrix letter Identification}
 \label{Classification}
 
-The following step is classify the descriptors $\{d_{kp}, kp =1,\dots, 12\}$ in order to determine the correct row and column of the chosen letter on the Speller Matrix.
+%The following step is classify the descriptors $\{d_{kp}, kp =1,\dots, 12\}$ in order to determine the correct row and column of the chosen letter on the Speller Matrix.
 The classification is carried out by using a discriminative semi-supervised classification method, based on the Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008}.
@@
 \item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
-\item Obtain the descriptor from the image of the signal plot according to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns. 
+\item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively.
+The parameter $k_a$ is an input of the algorithm.
 \item Compute  
+\begin{equation}
+x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c)
+\label{averaging}
+\end{equation}
+
+\item Plot the signal $x(t,c)$, according Section~\ref{Plot}.
+
+\item Obtain the descriptor from the image of the signal plot of $x(t,c)$ according to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns.
+
+\item Match with the Template $T$ by computing
 
@@
 By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
-The parameter $k$ is an input parameter of the algorithm.
+The  $k$ value is an input parameter of the algorithm.
 
+% By reversing the roles of the query and the class in equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.
 %The first step consists in extracting labeled \textit{hit} descriptors from the training set which are thus grouped together in a KD-tree \citep{Vedaldi2010} template set $ T $.
@@
 
+Figure~\ref{fig:classification} shows a schema of this process. 
+\begin{figure}
+\centering
+\includegraphics[width=15cm]{classificationgraph.pdf}
+\caption{Single trial segments are averaged for the 6 rows and 6 columns.}
+\label{fig:classification}
+\end{figure}
+
+%In the case of the P300 response, the oddball paradigm requires that one of the stimuli be infrequent.
+Hence this forces the data to be unbalanced~\citep{Tibon2015}.
+At the same time, the NBNN method suffers from biased classification on unbalanced classes~\citep{Fornoni2014}.
+%Para solucionar este problema,
+
 %Este párrafo no se entiende nada
 %------------------
-Especially in the case of the P300 response, the oddball paradigm requires that one of the stimuli needs to be infrequent.
-Hence this will unavoidably force the data to be unbalanced~\citep{Tibon2015}.
-At the same time, the NBNN method suffers from biased classification on unbalanced classes~\citep{Fornoni2014}.
-By reversing the roles of the query and the class in equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol}, it is only necessary to obtain the template set $ T $ with the learned descriptors representative of the P300 ERP, hence avoiding the problem of unbalanced classes.
-Figure~\ref{fig:classification} shows a schema of this idea,  where the classification is only performed against the template set T, for each one of the twelve classes, thus avoiding the unbalanced situation (2 hit vs.
-10 no-hit).
-\begin{figure}[H]
-\centering
-\includegraphics[width=15cm]{classificationgraph.pdf}
-\caption{For each letter, single trial segments are averaged for the 6 rows and 6 columns.}
-\label{fig:classification}
-\end{figure}
+
 %------------------
@@
 
-To verify the validity of the proposed framework and method, the public dataset 008-2014~\citep{Riccio2013} published on the BNCI-Horizon website~\citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, was used to perform an offline BCI Simulation to decode the spelled words from the provided signals.
-The algorithm was implemented using  VLFeat~\citep{Vedaldi2010} Computer Vision libraries on MATLAB 2014a (Mathworks Inc., Natick, MA, USA).
+To verify the validity of the proposed framework and method, the public dataset 008-2014~\citep{Riccio2013} published on the BNCI-Horizon website~\citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, is used.
+An own dataset with  the same experimental conditions is generated.
+Both of them are utilized to perform an offline BCI Simulation to decode the spelled words from the provided signals.
+
+The algorithm was implemented using  VLFeat~\citep{Vedaldi2010} Computer Vision libraries on MATLAB V2014a (Mathworks Inc., Natick, MA, USA).
+
+In the following sections the characteristics of the datasets and parameters of the identification algorithm are described.
 
@@
 A trial, as defined by the BCI2000 platform~\citep{Schalk2004}, is every attempt to select a letter from the speller.
-It is composed of signal segments corresponding to 10 repetitions of flashes of 6 rows and 10 repetitions of flashes of 6 columns of the matrix, yielding 120 repetitions.
+It is composed of signal segments corresponding to $k_a =10$ repetitions of flashes of 6 rows and $k_a =10$ repetitions of flashes of 6 columns of the matrix, yielding 120 repetitions.
 Flashing of a row or a column is performed for 0.125 s, following by a resting period (i.e.
@@
 
-The recorded dataset was sampled at 256 Hz and consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
+The recorded dataset was sampled at 256 Hz and it consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
 
@@
 In order to asses and verify the identification of the P300 response, subjects are instructed to perform a copy-spelling task.
-They have to fix their attention to successive letters in order to copy a previously determined set of words (in contrast to an online free-running operation of the speller where each users decides on its own what letter to choose).
+They have to fix their attention to successive letters for copying a previously determined set of words, in contrast to a free-running operation of the speller where each user decides on its own what letter to choose.
 
@@
 
-We replicated the same experiment on healthy subjects using g.Tec g.Nautilus 8 channels.
-The experiment protocol is exactly the same and we 20 subjects, age $ 30$, 5 females and 15 males, 6 right handed and the remaining left handed.
+We replicat the experiment on healthy subjects using g.Tec g.Nautilus 8 channels.
+The experiment protocol is  the same and we 20 subjects, age $ 30$, 5 females and 15 males, 6 right handed and the remaining left handed.
 
-Participants were recruited voluntarily and the experiment were conducted anonymyzing information in accordance with the declaration of Helsinki.
+Participants were recruited voluntarily and the experiment were conducted under anonymous information in accordance with the declaration of Helsinki.
 No monetary compensation was delivered and all participants agreed and signed a written informed consent.
-All healthy participants (n=13, 7 males, 6 females, average age 22.5 years (SD = 1,8 years), range 21–26 years) had normal or corrected-to normal vision and no history of neuropsychiatric dis- orders.
+All healthy participants had normal or corrected-to normal vision and no history of neuropsychiatric disorders.
+The experiment was performed with 13 subjects, 7 males, 6 females, average age 22.5 years, standard deviation  1,8 years, range 21–26 years.
 
@@
 The filtering was performed via the proprietary Simulink HighSpeed Online Processing block modules for g.MOBIlab and g.Nautilus systems, and via MATLAB code for the Xpress system.
-Table 1 summarizes the acquisition settings for each system.
+Table~\ref{tab:results} summarizes the acquisition settings for each system.
 
-8 gel-based active electrodes (g.LADYbird) + g.LADYbird (GND) + g.GAMMAearclip (REF) C3, Cz, C4, CPz, P3, Pz, P4, POz, GND: AFz, REF: right ear
+%8 gel-based active electrodes (g.LADYbird) + g.LADYbird (GND) + g.GAMMAearclip (REF) C3, Cz, C4, CPz, P3, Pz, P4, POz, GND: AFz, REF: right ear
 

### 2017-12-07 article.tex
Cambios en la sección de parámetros

@@
 
-%Este párrafo no se entiende nada y repite in order
 In order to asses and verify the identification of the P300 response, subjects are instructed to perform a copy-spelling task.
@@
 
-Parameters are selected according to the experimental protocol.
-As the P300 event latency and amplitude vary greatly between subjects, it is necessary to provide a patch that will be able to capture an entire transient event.
-Equations~\ref{eq:mapping1} and~\ref{eq:mapping2} can be used to map the original signal parameters to local image patch structure.
-%¿Qué es s? hay que hacer acordar. 
+Parameters are selected according to the experimental protocol. It is define in Section~\ref{Feature}, the patch size is $S_P = 3s.4 \times 3s.4$ .
+As the P300 event latency and amplitude vary greatly between subjects, it is necessary to utilize a size patch $S_P$ that be able to capture an entire transient event.
+With this purpose the $s$ value election is essential.
+%necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
+We proposed the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} for compute the scale value in horizontal and vertical directions, respectively. 
 \begin{equation}
-s= \frac{\Delta \mu V}{4 \cdot 3} \cdot \gamma 
+s_x = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
+\label{eq:mapping2}
+\end{equation}
+
+\begin{equation}
+s_y= \frac{\Delta \mu V}{4 \cdot 3} \cdot \gamma 
 \label{eq:mapping1}
@@
 
-\begin{equation}
-s = \frac{\lambda \cdot Fs}{4 \cdot 3} \cdot \gamma
-\label{eq:mapping2}
-\end{equation}
-%falta definir gamma
 \noindent where $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and $ \lambda $ is the length in seconds covered by the patch.
-% estos son los parámetros que vos usaste en tu experimento? 
-By using $ s = 3 $ and a quadruple scale of the image $ \gamma = 4 $ this gives the local patch, and the descriptor, the ability to identify events of 9 $ \mu V $ of amplitude, a resolution of $ 1 $ Pixel $ = \frac{1}{4} \mu V $ and span   of $ \lambda = 0.56 s $ (because the patch is a geometric square, $ s $ must be the same to map both the height and the required span of the signal).
-Finally, descriptor locations $\mathbf{kp}$  were selected at $ x =  0.55 s \cdot Fs \cdot \gamma = 35 $ and $ y = z(c) $ (Eq.
-\ref{eq:zerolevel}).
-%demasiado larga la frase, no se puede leer.
+In this way the whole event is capture.
+But we need a squared patch, then we discovered that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, span of $ \lambda = 0.56$ seconds.
+The resolution of $ 1 $ Pixel representing $  \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{Fs}\gamma$ in horizontal direction.
+Finally, descriptors  $\mathbf{kp}$  are located at $ x =  0.55 Fs.\gamma = 35 $ and $ y = z(c) $ (Eq.~\ref{eq:zerolevel}).
 

### 2017-12-10 response.letter.tex
Reescribi la sección del nuevo dataset, revisé la sección de los parámetros, cambie algunas imagenes, incluí algo de informacion de la selección del canal.  Complete parte de la carta de respuesta para el revisor. Pequeñas correcciones aqui y alla.

@@
 {\color{blue}
-Thank you for the observation. The title of the article was change accordingly to your suggestion.  
+Thank you for the observation. The title of the article was changed accordingly to your suggestion.  
 }
@@
 {\color{blue}
-This comment is very important, thank you very much. 
-An explanation  of which is a satisfactory result for Algorithm~\ref{Algoritmo} was added in sections~\ref{errorevaluation} and~\ref{sec:EdgeSimulatedData}.  
+You're absolutely right, thank you very much for your comment.
+We published in Ramele et al 2016 the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to decode Motor Imagery.
+We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and to decode SSVEP patterns.
+We modified the article to emphasize more clearly that the presented algorithm is  applicable to P300.
+
+% Como haces para que las referencias que estan aca en la respuesta apunten al articulo ???
+%An explanation  of which is a satisfactory result for Algorithm~\ref{Algoritmo} was added in sections~\ref{errorevaluation} and~\ref{sec:EdgeSimulatedData}.  
 }
@@
 {\color{blue}
-Thank you for the observation.  You are so clever that it is very likely that you are God.}
+Thank you for the observation.
+We worked hard to improve greatly all the mathematical description of the method.
+Section FEATURE was modified completely as well as section PARAMETERS.
+Section SIGNAL PLOT was revised and its mathematical terms corrected.}
 \end{quotation}
@@
 {\color{blue}
-Oh my God, I can't believe how smart you are !!! You probably are rich by now.
+We appreciate a lot your truthful comment.
+Equation (4) is very important for us because it is the essence of the method.
+We based that equation on the description of how the histogram of oriented gradients is calculated published in Vedaldi and Szeliski, Computer Vision and Algorithms.
+We reviewed the equation, clarified their terms and hope it is now more clear.
 }
@@
 {\color{blue}
-We have revised the manuscript and made a clear distinction between ``methods'' and ``features''.
-We thank you for noticing this, that led to a less ambiguous text.
+We have revised the manuscript and corrected equations (7) and (8), clarifying more precisely that they are part of the proposed method and that they help to understand how signal properties are related to the SIFT's geometric parameters.
+We thank you for noticing this, that led to a better clarification of mathematical terminology.
 }
@@
 {\color{blue}
-Thank you for the observation.  Smart people like you should be president.
-Following your suggestion, we added graphical evidence of the difficulty of separating targets characterized by $\mathcal{G}_I^0(-2,1,1)$, $\mathcal{G}_I^0(-3,1,1)$, $\mathcal{G}_I^0(-4,1,1)$, and $\mathcal{G}_I^0(-5,1,1)$ data; cf.\ Fig.~\ref{fig:DifficultProblem}.
-With this, and the corresponding discussion, we believe we have addressed your point.
+Thank you for the observation.
+We modified the equation to emphasize the point that is actually calculated in digital space.
+We also clarified that image gradients are also calculated based on finite differences.
+With this we hope we have addressed your point.
  }
@@
 Thank you for the observation. 
-We now specify details of the E-SAR image used in the applications to true data; cf. Sec.~\ref{sec:ActualData}.
-Following your suggestion, we added  a comparison with the Triangular distance in actual data. 
-We applied the method in edge detection with both distances and we discovered that both of them resulted in exactly the same edge point. 
-Then, we concluded that these distances are very similar in terms of accuracy in finding the transition point, but the geodesic distance is better in terms of  computation time. 
 We fully addressed your point.
@@
 {\color{blue}
-Thank you for your suggestion. We increased the discussion and conclusions, and enhanced lines for future research.
+Thank you for your suggestion.
+The original dataset was published at the BNCI-Horizon website, which is very popular in BCI research.
+Our approach emphasize the structure of the signal in terms of their morphological components in their visual description.
+THIS IS TOUGH TO JUSTIFY
+In order to address this issue, we included a new dataset that we generated ourselves, on healthy subjects that we hope now improves the number of different subjects and that can help to understand the method.
 }
@@
 {\color{blue}
-Oh God. You are clearly an oracle. 
-All these points were modified according to the suggestions, but instead of doing what you asked for we do what we want instead.
+THIS IS EVEN TOUGHER.
+
+ACA LE PUEDO PONER QUE HAY POCOS METODOS QUE PERMITAN ENGANCHAR BIEN QUE EL METODO
+
+All these points were modified according to the suggestions.
 We hope this is acceptable.
@@
 {\color{blue}
-Oh God. You are clearly an oracle. 
-All these points were modified according to the suggestions, but instead of doing what you asked for we do what we want instead.
+The Results section was modified to clarify more precisely the expectation about the EEG channels where a visual P300 detection task would be obtained.
+We added references and improved the section to clarify this point more thoughtfully.
 We hope this is acceptable.
@@
 {\color{blue}
-Oh God. You are clearly an oracle. 
-All these points were modified according to the suggestions, but instead of doing what you asked for we do what we want instead.
-We hope this is acceptable.
+Thank you a lot for this important comment.
+We re-read all the manuscript and asked to collegues to proofread the article.
+They gave us interesting suggestions that helped us to improve different sections.
+We would be very pleased if you find it now easier to read it.
 }
@@
 {\color{blue}
-Oh God. You are clearly an oracle.
-
-We hope this is acceptable.
+Thank you very much for this information.
+We reviewed completely that equation (it is now equation (6) and equation (7)) and we included a step-by-step procedure that describes the algorithm that we propose and we used.
 }

### 2017-12-10 article.tex
Reescribi la sección del nuevo dataset, revisé la sección de los parámetros, cambie algunas imagenes, incluí algo de informacion de la selección del canal.  Complete parte de la carta de respuesta para el revisor. Pequeñas correcciones aqui y alla.

@@
 Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.
-Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Near Neighbor  and the final Section shows results and discussion where we expose our remarks, conclusions and future work.
+Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008} and the final Section shows results and discussion where we expose our remarks, conclusions and future work.
 
@@
 This signal can be used to implement a speller application by means of a Speller Matrix~\citep{Farwell1988}.
-Fig.~\ref{fig:p300matrix} shows an example of a Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
+Fig.~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
 Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
@@
 
-The first step consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. 
+The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. 
 The processing pipeline starts by applying a notch filter to the raw signal, a 
@@
 The aim is to identify the selected letter from the matrix.
-Previously, during training phase, descriptors from labeled segments are extracted.
+Previously, during training phase, descriptors from \textit{hit} labeled segments are extracted.
 These descriptors are the P300 templates which are grouped in a template set called $ T $.
-Rows are labeled 1-6, whereas columns are labeled 7-12.
+Segments corresponding to rows are labeled 1-6, whereas those corresponding to columns are labeled 7-12.
 The process has the following steps:
@@
 \item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively.
-The parameter $k_a$ is an input of the algorithm.
+The parameter $k_a$ is an input of the algorithm, and corresponds to the number of repetitions.
 \item Compute  
@@
 
-\item Obtain the descriptor from the image of the signal plot of $x(t,c)$ according to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns.
+\item Obtain the descriptor from the image of the signal plot of $x(t,c)$ in accordance to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns.
 
-\item Match with the Template $T$ by computing
+\item Match to the Template $T$ by computing
 
@@
 By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
-The  $k$ value is an input parameter of the algorithm.
 
@@
 \centering
-\includegraphics[width=15cm]{classificationgraph.pdf}
+\includegraphics[width=15cm]{classificationgraph2.pdf}
 \caption{Single trial segments are averaged for the 6 rows and 6 columns.}
@@
 To verify the validity of the proposed framework and method, the public dataset 008-2014~\citep{Riccio2013} published on the BNCI-Horizon website~\citep{Brunner2014} by  IRCCS Fondazione Santa Lucia, is used.
-An own dataset with  the same experimental conditions is generated.
+Additionally, an own dataset with  the same experimental conditions is generated.
 Both of them are utilized to perform an offline BCI Simulation to decode the spelled words from the provided signals.
 
-The algorithm was implemented using  VLFeat~\citep{Vedaldi2010} Computer Vision libraries on MATLAB V2014a (Mathworks Inc., Natick, MA, USA).
+The algorithm is implemented using  VLFeat~\citep{Vedaldi2010} Computer Vision libraries on MATLAB V2014a (Mathworks Inc., Natick, MA, USA).
 
@@
 
-\subsubsection{P300 ALS Public Dataset}
+\subsubsection{P300 ALS Public Dataset} \label{ALSDataset}
 
@@
 
-The recorded dataset was sampled at 256 Hz and it consisted of an EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
+The recorded dataset was sampled at 256 Hz and it consisted of scalp EEG matrix for electrode channels Fz,Cz,Pz,Oz,P3,P4,PO7 and PO8, identified according to the 10-20 International System,  for each one of the 8 subjects.
+The recording device was a research-oriented digital EEG device (g.Mobilab, g.Tec, Austria) and the data acquisition and stimuli delivery were handled by the BCI2000 open source software~\citep{Schalk2004}.
 
@@
 
+\subsubsection{P300 for healthy subjects}
+
+We replicated the same experiment on healthy subjects using a wireless digital EEG device (g.Nautilus, g.Tec, Austria).
+The experimental conditions were the same as those used for the previous dataset, as detailed in section~\ref{ALSDataset}.
+
+Participants were recruited voluntarily and the experiment were conducted anonymously in accordance with the declaration of Helsinki published by the World Health Organization.
+No monetary compensation was handed out and all participants agreed and signed a written informed consent.
+All healthy subjects had normal or corrected-to normal vision and no history of neurological disorders.
+The experiment was performed with 20 subjects, 7 males, 4 females, average age 22.5 years, standard deviation  1,8 years, range 21–26 years.
+
+Gel-based active electrodes (g.LADYbird, g.Tec, Austria) were used on the same locations Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.
+Reference was set to the right ear lobe and ground was preset as the AFz position.
+Sampling frequency is slightly different, and was set to 250 Hz, which is the closest possible to the one used with the other dataset.
+
+%EEG data were recorded from the same 8 electrodes positions (C3, Cz, C4, CPz, P3, Pz, P4, POz) with all the systems.
+The reference positions were placed at the right ear.
+The ground positions was placed at the left ear, except with the g.Nautilus system (placed at AFz, because this could not be changed).
+Sampling rate was set at 250 Hz, the closest possible to predefined g.Mobilab sampling rate (256 Hz).
+Data were acquired notch filtered at 50 Hz and passband filtered between 2 Hz and 30 Hz.
+The filtering was performed via the proprietary Simulink HighSpeed Online Processing block modules for g.MOBIlab and g.Nautilus systems, and via MATLAB code for the Xpress system.
+Table~\ref{tab:results} summarizes the acquisition settings for each system.
+
+%8 gel-based active electrodes (g.LADYbird) + g.LADYbird (GND) + g.GAMMAearclip (REF) C3, Cz, C4, CPz, P3, Pz, P4, POz, GND: AFz, REF: right ear
+
 \subsubsection{Parameters}
 
-Parameters are selected according to the experimental protocol. It is define in Section~\ref{Feature}, the patch size is $S_P = 3s.4 \times 3s.4$ .
-As the P300 event latency and amplitude vary greatly between subjects, it is necessary to utilize a size patch $S_P$ that be able to capture an entire transient event.
-With this purpose the $s$ value election is essential.
+% Antes estaba esto aca
+
+% The P300 event latency and amplitude vary greatly between subjects,
+
+% lo saque porque confunde, ya que despues de todo, dejamos todo fijo, e ignoramos completamente el hecho de que varie.
+
+Parameters are selected according to the experimental protocol.
+First of all, the patch size in pixels is $S_P = 3s.4 \times 3s.4$.
+The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $.
+Hence it is necessary to utilize a size patch $S_P$ that could capture an entire transient event.
+With this purpose in consideration, the $s$ value election is essential.
 %necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
-We proposed the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} for compute the scale value in horizontal and vertical directions, respectively. 
+We propose the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
 \begin{equation}
@@
 \noindent where $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and $ \lambda $ is the length in seconds covered by the patch.
-In this way the whole event is capture.
-But we need a squared patch, then we discovered that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, span of $ \lambda = 0.56$ seconds.
-The resolution of $ 1 $ Pixel representing $  \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{Fs}\gamma$ in horizontal direction.
+The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.
+This also provides a resolution of $ 1 $ Pixel representing $  \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{Fs \cdot \gamma}$ seconds in the horizontal direction.
 Finally, descriptors  $\mathbf{kp}$  are located at $ x =  0.55 Fs.\gamma = 35 $ and $ y = z(c) $ (Eq.~\ref{eq:zerolevel}).
+In this way the whole transient event should be captured.
 
@@
 \centering
-\includegraphics[width=10cm]{localpatch.png}
+\includegraphics[width=10cm]{patchgeometry.pdf}
 \caption{The local patch captures the signal information based on both the vertical and horizontal scale.}
@@
 \end{figure}
-
-\subsubsection{P300 for healthy subjects}
-
-We replicat the experiment on healthy subjects using g.Tec g.Nautilus 8 channels.
-The experiment protocol is  the same and we 20 subjects, age $ 30$, 5 females and 15 males, 6 right handed and the remaining left handed.
-
-Participants were recruited voluntarily and the experiment were conducted under anonymous information in accordance with the declaration of Helsinki.
-No monetary compensation was delivered and all participants agreed and signed a written informed consent.
-All healthy participants had normal or corrected-to normal vision and no history of neuropsychiatric disorders.
-The experiment was performed with 13 subjects, 7 males, 6 females, average age 22.5 years, standard deviation  1,8 years, range 21–26 years.
-
-EEG data were recorded from the same 8 electrodes positions (C3, Cz, C4, CPz, P3, Pz, P4, POz) with all the systems.
-The reference positions were placed at the right ear.
-The ground positions was placed at the left ear, except with the g.Nautilus system (placed at AFz, because this could not be changed).
-Sampling rate was set at 250 Hz, the closest possible to predefined g.Mobilab sampling rate (256 Hz).
-Data were acquired notch filtered at 50 Hz and passband filtered between 2 Hz and 30 Hz.
-The filtering was performed via the proprietary Simulink HighSpeed Online Processing block modules for g.MOBIlab and g.Nautilus systems, and via MATLAB code for the Xpress system.
-Table~\ref{tab:results} summarizes the acquisition settings for each system.
-
-%8 gel-based active electrodes (g.LADYbird) + g.LADYbird (GND) + g.GAMMAearclip (REF) C3, Cz, C4, CPz, P3, Pz, P4, POz, GND: AFz, REF: right ear
 
@@
 Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
-From the 7 trials for each subject, the first 3 were used as training, and the remaining 4 for testing.
+From the 7 words for each subject, the first 3 were used as training, and the remaining 4 for testing.
 The best performing channel is informed as well.
+
+DRAFT: Traditionally P300 can be seen in this channel and this other channel.
+However, Wolpaw showed that the response may also arise in occipital channels as well.
+The authors of the original paper concluded that Cz was the best performing channel.
+However, in our approach
+
 It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.}
+\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.
+The authors would like to thank Dr.
+Juliana Gambini for their insights and help with the detailed description of the histogram of gradients.}
 

### 2017-12-11 response.letter.tex
cambios de hoy

@@
 {\color{blue}
-You're absolutely right, thank you very much for your comment.
+You're  right, thank you very much for your comment.
 We published in Ramele et al 2016 the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to decode Motor Imagery.
 We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and to decode SSVEP patterns.
-We modified the article to emphasize more clearly that the presented algorithm is  applicable to P300.
+Modificar%We modified the article to emphasize more clearly that the presented algorithm is  applicable to P300.
 
@@
 Thank you for the observation.
-We worked hard to improve greatly all the mathematical description of the method.
-Section FEATURE was modified completely as well as section PARAMETERS.
-Section SIGNAL PLOT was revised and its mathematical terms corrected.}
+We worked hard to improve all the mathematical description of the method.
+Section FEATURE was modified as well as section PARAMETERS.
+Section SIGNAL PLOT was revised and its mathematical terms were corrected.}
 \end{quotation}
@@
 Equation (4) is very important for us because it is the essence of the method.
-We based that equation on the description of how the histogram of oriented gradients is calculated published in Vedaldi and Szeliski, Computer Vision and Algorithms.
-We reviewed the equation, clarified their terms and hope it is now more clear.
+%We based that equation on the description of how the histogram of oriented gradients is calculated published in Vedaldi and Szeliski, Computer Vision and Algorithms.
+We reviewed the equation, clarified their terms and we hope it is now more clear.
 }
@@
 {\color{blue}
-We have revised the manuscript and corrected equations (7) and (8), clarifying more precisely that they are part of the proposed method and that they help to understand how signal properties are related to the SIFT's geometric parameters.
-We thank you for noticing this, that led to a better clarification of mathematical terminology.
+We have revised the manuscript and corrected equations (7) and (8) which specify the ration between the signal properties and the SIFT's geometric parameters. 
+We thank you for noticing this.
 }
@@
 Thank you for the observation.
-We modified the equation to emphasize the point that is actually calculated in digital space.
-We also clarified that image gradients are also calculated based on finite differences.
-With this we hope we have addressed your point.
+We modified all equations to emphasize the point that they are calculated in digital space.
+We also added that image gradients are also calculated using finite differences.
+With this, we hope we have addressed your point.
  }
@@
 The original dataset was published at the BNCI-Horizon website, which is very popular in BCI research.
-Our approach emphasize the structure of the signal in terms of their morphological components in their visual description.
-THIS IS TOUGH TO JUSTIFY
-In order to address this issue, we included a new dataset that we generated ourselves, on healthy subjects that we hope now improves the number of different subjects and that can help to understand the method.
+Our approach emphasize the structure of the signal in terms of their morphological components in their visual description.% THIS IS TOUGH TO JUSTIFY
+In order to address this issue, we included a new dataset that we generated ourselves, on healthy subjects that we hope now improves the number of different subjects and that can help to validate the method.
 }

### 2017-12-11 article.tex
cambios de hoy

@@
 On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered.
-A local image patch is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
-It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center.
+A local image patch of size $S_p \times S_p$ is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, then $S_p = 4.3s $ pixels.
 
@@
 Lastly, the fixed value of $ 3 $ is a magnification factor which corresponds to the number of pixels per each block when $s = 1$.
-As the patch is of size $4 \times 4$ and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained.
+As the patch has  $16$ blocks and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained.
 It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $ , so the method considers both, the magnitude and the orientation of the gradient vector.
@@
 \item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively.
-The parameter $k_a$ is an input of the algorithm, and corresponds to the number of repetitions.
+The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
 \item Compute  
@@
 
-Parameters are selected according to the experimental protocol.
-First of all, the patch size in pixels is $S_P = 3s.4 \times 3s.4$.
+%Parameters are selected according to the experimental protocol. 
+The patch size is $S_P = 3s.4 \times 3s.4$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
 The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $.
@@
 
-\noindent where $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and $ \lambda $ is the length in seconds covered by the patch.
+\noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
 The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.
-This also provides a resolution of $ 1 $ Pixel representing $  \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{Fs \cdot \gamma}$ seconds in the horizontal direction.
-Finally, descriptors  $\mathbf{kp}$  are located at $ x =  0.55 Fs.\gamma = 35 $ and $ y = z(c) $ (Eq.~\ref{eq:zerolevel}).
-In this way the whole transient event should be captured.
-
+This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds in the horizontal direction.
+Finally, descriptors  $\mathbf{kp}$  are located at $ x_{kp} =  0.55 Fs.\gamma = 35 $ and $ y_{kp} = z(c) $ (see Eq.~\ref{eq:zerolevel}).
+In this way the whole transient event is captured.
+Figure~\ref{fig:sampledescriptor2} shows a patch of a signal plot covering the complete amplitude (vertical direction) and the complete span of the signal event (horizontal direction). 
 % esta figura no está comentada en el texto.
@@
 \includegraphics[width=10cm]{patchgeometry.pdf}
-\caption{The local patch captures the signal information based on both the vertical and horizontal scale.}
+\caption{The scale of local patch is selected in order to capture the whole signal information}
 \label{fig:sampledescriptor2}
 \end{figure}
+
+%Hacemos algún comentario de k_a?
 

### 2017-12-12 article.tex
comentarios

@@
 \pubvolume{xx}
-\pubyear{2016}
-\copyrightyear{2016}
+\pubyear{2017}
+\copyrightyear{2017}
 \externaleditor{Academic Editor: name}
@@
 % Full title of the paper (Capitalized)
-\Title{Oriented Gradient Histogram applied to P300 Detection}
+\Title{Histogram of Gradient Orientations of Signal Plots applied to P300 Detection}
 
 % Authors, for the paper (add full first names)
-\Author{Rodrigo Ramele $^{1,\dagger}$*, Ana Julia Villar $^{1}$ and Juan Miguel Santos  $^{1}$}
+\Author{Rodrigo Ramele $^{1}$*, Ana Julia Villar $^{1}$ and Juan Miguel Santos  $^{1}$}
 % Authors, for metadata in PDF
@@
 
-% Current address and/or shared authorship
-\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina}
+% Current address and/or shared authorship  COMMENTED PER REVIEWER
+%\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina}
+
+%\firstnote{.}
 
@@
 It also aims to provide a framework to analyze, characterize and classify EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.
-The validity of the method is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients.}
+The validity of the method is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients and an own dataset for healthy subjects.}
 
 % Keywords
-\keyword{electroencephalography (EEG); BCI; P300; ALS; NBNN; HOG; SIFT}
+\keyword{EEG; BCI; P300; ALS; NBNN; HGOSP; SIFT}
 
@@
 
-%\begin{enumerate}[leftmargin=*,labelsep=3mm]
-%\item	Why
-%\item	Purpose and significance
-%\item	Current state of the research field and key publications cited
-%\item   Controversial hypothesis
-%\item   Main aim and principal conclusions.
-%\end{enumerate}
-%
-%\begin{enumerate}[leftmargin=*,labelsep=3mm]
-%\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
-%\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
-%\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
-%\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
-%\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
-%\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
-%\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
-%\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
-%\end{enumerate}
+Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitized devices has pressed for a revamping of this hundred years old technology.
+Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information~\cite{DeVos2014}.
 
-Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundred years old technology.
-Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information \cite{DeVos2014}
+A key contribution to this expansion has been the field of Brain Computer Interfaces (BCI)~\citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neurodegenerative diseases.
 
-A key contribution to this expansion has been the field of Brain Computer Interfaces (BCI) \citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neurodegenerative diseases.
-
-One noteworthy aspect of this novel communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, in a Virtual Reality environment \citep{Lotte2013} or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
-The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.
+One noteworthy aspect of this novel communication channel is the ability to transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair~\citep{Carlson2013}, as input to a speller application~\citep{Guger2009a}, in a Virtual Reality environment~\citep{Lotte2013} or as aiding tool in a rehabilitation procedure~\citep{Jure2016}.
+The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion~\citep{WolpawJonathanR2012}.
 
@@
 
-Moreover, EEG markers \citep{Clerc} that can be used to volitionally transmit information are limited, and each one of them has a particular combination of appropriate methods to decode them.
+Moreover, EEG markers~\citep{Clerc} that can be used to  transmit volitional information are limited, and each one of them has a particular combination of appropriate methods to decode them.
 Inevitably, it is necessary to implement  distinct and specialized algorithmic methods, to filter the signal, enhance its Signal to Noise Ratio (SNR), and try to determine some meaning out of it.
 
-%Patients suffering from ALS their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
-
-%Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
-It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
-
-BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon  \citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
-New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments \citep{Huggins2016}.
+BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon~\citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
+New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments~\citep{Guger2017,Huggins2016}.
 However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement~\citep{Clerc}.
 
-In~\citep{Ramele2016},  authors introduce a method for classification of rhythmic patterns using gradient histograms orientations.
-Inspired in that work, we proposed a novel application of the developed method to classify and describe transient events like those produced by P300 Event Related Potential.
- Its validity is verified by processing offline data for ALS patients.  
-The proposed approach is based on the morphological analysis of the shape of the EEG signal~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and it was built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting  raw signal plots~\citep{Hartman2005}.
+A few works have explored the idea of exploiting the signal waveform to analyze the EEG signal.
+In~\citep{Alvarado-Gonzalez2016} an approach based on Slope Horizontal Chain Code is presented, whereas in ~\citep{Yamaguchi2009} a similar procedure was implemented based on Mathematical Morphological Analysis.
+The seminal work of Bandt-Pompe Permutation Entropy~\citep{Berger2017} also explores succinctly this idea as a basis to establish the time series ordinal patterns.
+In the article~\citep{Ramele2016},  the authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves  and Motor Imagery rolandic central $\mu$ rhythms using the histogram of gradient orientations of signal plots.
+Inspired in that work, we propose a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.
+The proposed approach is based on the waveform analysis of the shape of the EEG signal, but using histogram of gradient orientations.
+The method is built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting raw signal plots, as it is described in~\citep{Hartman2005}.
 
-This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) describe the way in which this classification can be used to implement an offline P300-based BCI Speller application using two public datasets.
+This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of gradient orientations extracted from the image of signal plots, and (2) describe the way in which this classification can be used to implement an offline P300-based BCI Speller application using two public datasets.
+Its validity is verified by processing offline data for ALS patients and for healthy subjects.
 
-This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained.
-Section \ref{Preprocessing} and \ref{Pipeline} describe the processing pipeline.
-Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.
-Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008} and the final Section shows results and discussion where we expose our remarks, conclusions and future work.
-
-%\subsection{Morphological Signal Analysis}
-%
-%Several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
-%CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016,Yamaguchi2009}.
-%
-%Morphological or 
-%
-%template matching
-%
-%spike
-%sharp wave
-%spike wave
-%polyspike 
-%slow wave complex
-%
-%shape domain
-%
-%\subsection{P300}
-%
-%Uhhhh I can talk a lot about P300 \citep{Knuth2006}.
-%
-%P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.
+This article unfolds as follows: Section~\ref{Feature} is dedicated to explain the Feature Extraction method based on Histogram of Gradient Orientations of the Signal Plot: Section~\ref{Pipeline} shows the preprocessing pipeline,  Section~\ref{Plot}  describes the image generation of the signal plot, Section~\ref{SIFT}  presents the feature extraction procedure while  Section~\ref{Classification}  introduces the whole  Speller Matrix Letter Identification procedure including the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008}.
+In Section~\ref{Protocol}, the experimental protocol is expounded.
+Section~\ref{Results} shows the results of applying the proposal technique and a discussion is presented.
+In the final Section~\ref{conclusion}  we expose our remarks, conclusions and future work.
+%This kind of EEG events are described in~\citep{Hartman2005} and~\citep{Ben-Simon2013}.
 
 \section{Materials and Methods}
-
-\subsection{Feature Extraction based on Histogram of Gradients of the Signal Plot} \label{Feature}
 
@@
 This signal can be used to implement a speller application by means of a Speller Matrix~\citep{Farwell1988}.
-Fig.~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
-Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
+Fig.~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashes of rows and columns provide the deviant stimulus required to elicit this physiological response.
+Each time a row or a column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
 
@@
 \subfigure[Column flash.]{\includegraphics[width=.29\linewidth]{p300matrixcoloriginal.png}}
-\caption{Example of a Speller Matrix.  Rows and columns flash intermittently in random permutations.}
+\caption{Example of the $6 \times 6$ Speller Matrix used in the study.  Rows and columns flash intermittently in random permutations.}
 \label{fig:p300matrix}
@@
 
-%The proposed algorithm was designed to perform specifically this decoding.
+\subsection{Feature Extraction based on Histogram of Gradient Orientations of the Signal Plot} \label{Feature}
 
-% of the selected letter based on the detection of the P300 Event Related Potential response on the corresponding EEG signal.
+In this section, the signal preprocessing, the method for generating images from signal plots, the feature extraction procedure and the speller matrix identification are described.
 
-\subsubsection{Preprocessing} \label{Preprocessing}
-
-The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. 
-The processing pipeline starts by applying a notch filter to the raw signal, a 
-$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.
-
-\subsubsection{Processing Pipeline} \label{Pipeline}
+\subsubsection{Preprocessing Pipeline} \label{Pipeline}
 
 \begin{itemize}
-\item \textbf{Artefact Removal}: The EEG signal matrix is processed on a channel by channel basis.
+\item \textbf{Signal Enhancement}: The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG.
+The processing pipeline starts by applying a notch filter to the raw digital signal, a
+$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.
+\item \textbf{Artifact Removal}: The EEG signal matrix is processed on a channel by channel basis.
 For every 12 flashing stimuli, i.e.
 one complete sequence of intensification of each of the $6$ rows plus the $6$ columns, a basic artefact elimination procedure is implemented by removing the entire sequence when any signal deviates above/bellow $ \pm 70 \mu V $.
-\item \textbf{Segmentation}: For each of the $12$ stimuli,  a window of $1$ second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
-Two of these segments are labeled as \textit{hit}, whereas the remaining $10$ are labeled as \textit{no hit}.
-A hit represents that the EEG segment should contain the P300 ERP signature time-locked to the flashing stimulus.
+\item \textbf{Segmentation}: For each of the $12$ stimuli,  a window of $t_{max} = 1$ second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
+Two of these segments should contain the P300 ERP signature time-locked to the flashing stimulus, one for the row, and one for the column.
 \item \textbf{Signal Averaging}:  The P300 ERP is deeply buried under background EEG so the traditional approach to identify it is by point-to-point averaging the time-locked stacked signal segments.
-Hence the values which are not related to, and not time-locked to the onset of the stimulus are canceled out \citep{Liang2008}.
+Hence the values which are not related to, and not time-locked to the onset of the stimulus are canceled out~\citep{Liang2008}.
 \end{itemize}
@@
 
-The underlying idea of this method is to generate a template for the signal \citep{Ramele2016},  based on the image plot.
-Hence, the first step is the signal transformation into a temporary binary image.
-
-Averaged signal segments obtained from the preprocessing stage are first scaled and standardized (i.e. z-score) by
+Averaged signal segments are standardized and scaled by
 
 \begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor
+\tilde{x}(n,c) = \left \lfloor{ \gamma \cdot \frac{( x(n,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, n \in \{ 1, \dots, n_{max}\},c \in \{1,2,\dots,Ch\}
+\label{eq:standarizedaverages}
 \end{equation}
 
-\noindent where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for  channel $c$.
-Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
+\noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related with the image scale.
+In addition, $ x(n,c) $ is the point-to-point averaged EEG matrix for the sample point $n$ and for channel $c$, $n_{max}=F_s.t_{max}$ and $F_s$ is the sampling frequency.
+The parameter  $Ch$ is the number of available EEG channels.
+Lastly, $$\bar{x}(c) =\frac{1}{n_{max}}\sum_{n=1}^{n_{max}}x(n,c)$$ and $$ \sigma(c) = (\frac{1}{n_{max}-1}\sum_{n=1}^{n_{max}}(x(n,c)-\bar{x}(c))^2 )^{\frac{1}{2}}$$ are the mean and standard deviation of $x(n,c), n \in 1,\dots,n_{max}$, for each channel $c$.
 
@@
 I(z_1,z_2) = \left\{ \begin{array}{rl}
-255 & \text{if} \,  z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
+255 & \text{if} \,  z_1 = \gamma \cdot n; \; z_2 = \tilde{x}(n,c) + z(c) \\
 0   & \mbox{otherwise}
@@
 
-\noindent where $ z_1$ and $z_2$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image.
-The function $z(c)$ is the \textit{zerolevel} which is the location on the image where the signal's zero value should be located in order to fit the entire signal within the image:
+\noindent where $ (z_1,z_2) \in \mathbb{N} \times \mathbb{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$.
+The function $z(c)$, $c \in \{1,2,\dots,Ch\}$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
 
 \begin{equation}
-z(c) = \left \lfloor{ \frac{\max \tilde{x}(t,c)  - \min \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max \tilde{x}(t,c)  + \min \tilde{x}(t,c)}{ 2} }\right \rfloor
+z(c) = \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(n,c)  - \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(t,c)  + \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c)}{ 2} }\right \rfloor
 \label{eq:zerolevel}
@@
 
-%\noindent and is used to derive local representations that captures the visual shape of the signal.
-
-\subsubsection{Feature Extraction: Histogram of Oriented Gradients}
+\subsubsection{Feature Extraction: Histogram of Gradient Orientations}
 \label{SIFT}
@@
 
-On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered.
-A local image patch of size $S_p \times S_p$ is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
-It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, then $S_p = 4.3s $ pixels.
+On the generated image $I$, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered.
+A local image patch of size $S_p \times S_p$ pixels is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, thus $S_p = 4.3s $ pixels.
 
@@
 
-\noindent  where $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $, $ i,j \in \{1,2,3,4\} $, $ |J(\mathbf{x})| $ is the norm of the gradient vector found at each block of the patch using finite differences, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin, $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$ and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
+\noindent  where $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$, $ i,j \in \{1,2,3,4\} $, $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{x})| $ is the norm of the gradient vector in the pixel $\mathbf{x}$ and it is computed using finite differences, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin,   and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
 in~\citep{Lowe2004,Vedaldi2010}.
@@
 As the patch has  $16$ blocks and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained.
-It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $ , so the method considers both, the magnitude and the orientation of the gradient vector.
+It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $, so the method considers both, the magnitude and the orientation of the gradient vector.
 
@@
 The aim is to identify the selected letter from the matrix.
-Previously, during training phase, descriptors from \textit{hit} labeled segments are extracted.
+Previously, during training phase, two descriptors are extracted from averaged signal segments which correspond to the letter where the user was supposed to be focusing onto.
 These descriptors are the P300 templates which are grouped in a template set called $ T $.
@@
 \begin{enumerate}
-\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
-\item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively.
+\item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
+\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(n,c),\dots,S_{k_a}(n,c)$, where the variables $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to time and channel, respectively.
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
-\item Compute  
+\item \label{paso3} Compute the Ensemble Average by
 \begin{equation}
-x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c)
+x(n,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(n,c),n \in \{1, \dots, n_{max}\}, c \in \{1,\dots,Ch\}
 \label{averaging}
 \end{equation}  
+for each row and column.
 
-\item Plot the signal $x(t,c)$, according Section~\ref{Plot}.
+\item \label{paso4}Plot the signal $x(n,c)$,  $n \in \{1, \dots, n_{max}\}$, $c \in \{1,\dots,Ch\}$,  according Section~\ref{Plot}.
 
-\item Obtain the descriptor from the image of the signal plot of $x(t,c)$ in accordance to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns.
+\item Repeat steps~\ref{paso2}, \ref{paso3} and~\ref{paso4} in order to generate the images $I^{row}_1, \dots, I^{row}_6$ and $I^{col}_7,\dots,I^{col}_{12}$ for rows and columns, respectively.
+
+\item Obtain the descriptors $ d^{row}_1, \dots,  d^{row}_6 $  and  $ d^{col}_7, \dots,  d^{col}_{12} $ for rows and columns, respectively from $I^{row}_1, \dots, I^{row}_6$ and $I^{col}_7,\dots,I^{col}_{12}$  in accordance to the method described in Section~\ref{SIFT}.
 
@@
 \begin{equation}
-\hat{r} = \arg \min_{u \in \{1,\dots,6\}} \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2
+\hat{row} = \arg \min_{u \in \{1,\dots,6\}} \sum_{q \in NN_T(d^{row}_u)}^{} \left\lVert q -  d^{row}_u \right\rVert ^2
 \label{eq:multiclassificationrow}
@@
 \begin{equation}
-\hat{c} = \arg \min_{u \in \{7,\dots,12\}} \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2
+\hat{col} = \arg \min_{u \in \{7,\dots,12\}} \sum_{q \in NN_T(d^{col}_u)}^{} \left\lVert q -  d^{col}_u \right\rVert ^2
 \label{eq:multiclassificationcol}
@@
 
-where $NN_T(d^l_u), \;l=r,c$ is the set of the $k$ nearest neighbors to $d^l_u, \;l=r,c$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l=r,c$.
-By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
+where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u, \;l\in\{row,{col}\}$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l\in\{row,{col}\}$.
 
@@
 \end{enumerate}
-
-Figure~\ref{fig:classification} shows a schema of this process. 
-\begin{figure}
+By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{row} $ and column $ \hat{col} $. 
+Figure~\ref{fig:classification} shows a scheme of this process. 
+\begin{figure}[htp]
 \centering
-\includegraphics[width=15cm]{classificationgraph2.pdf}
-\caption{Single trial segments are averaged for the 6 rows and 6 columns.}
+\includegraphics[width=15cm]{classificationgraph.pdf}
+\caption{Single trial segments $S_i$ are averaged for the 6 rows and 6 columns.
+From the averaged signal, the image of the signal plot is generated and each descriptor is computed.
+By comparing each descriptor against the set of templates, the P300 ERP can be detected, and finally the desired letter from the matrix can be inferred.}
 \label{fig:classification}
@@
 
+%Este párrafo no se entiende nada
+%------------------
+
+%------------------
 %In the case of the P300 response, the oddball paradigm requires that one of the stimuli be infrequent.
@@
 
-%Este párrafo no se entiende nada
-%------------------
-
-%------------------
 \subsection{Experimental Protocol} \label{Protocol}
@@
 Flashing of a row or a column is performed for 0.125 s, following by a resting period (i.e.
-Inter-stimulus interval) of the same length.
+inter-stimulus interval) of the same length.
 After 120 repetitions an inter-trial pause is included before resuming with the following letter.
@@
 
-We replicated the same experiment on healthy subjects using a wireless digital EEG device (g.Nautilus, g.Tec, Austria).
-The experimental conditions were the same as those used for the previous dataset, as detailed in section~\ref{ALSDataset}.
+We replicate the same experiment on healthy subjects using a wireless digital EEG device (g.Nautilus, g.Tec, Austria).
+The experimental conditions are the same as those used for the previous dataset, as detailed in section~\ref{ALSDataset}.
 
-Participants were recruited voluntarily and the experiment were conducted anonymously in accordance with the declaration of Helsinki published by the World Health Organization.
-No monetary compensation was handed out and all participants agreed and signed a written informed consent.
-All healthy subjects had normal or corrected-to normal vision and no history of neurological disorders.
-The experiment was performed with 20 subjects, 7 males, 4 females, average age 22.5 years, standard deviation  1,8 years, range 21–26 years.
+Participants are recruited voluntarily and the experiment is conducted anonymously in accordance with the declaration of Helsinki published by the World Health Organization.
+No monetary compensation is handed out and all participants agree and sign a written informed consent.
+All healthy subjects have normal or corrected-to normal vision and no history of neurological disorders.
+The experiment is performed with 8 subjects, 6 males, 2 females, 6 right-handed, 2 left-handed, average age 28.25 years, standard deviation  9.58 years, range 20-50 years.
 
-Gel-based active electrodes (g.LADYbird, g.Tec, Austria) were used on the same locations Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.
-Reference was set to the right ear lobe and ground was preset as the AFz position.
-Sampling frequency is slightly different, and was set to 250 Hz, which is the closest possible to the one used with the other dataset.
+EEG data is collected in a single recording session.
+Participants are seated in a comfortable chair, with their vision aligned to a computer screen located one meter in front of them.
+The handling and processing of the data and stimulations is conducted by the OpenVibe platform~\citep{Renard2010}.
 
-%EEG data were recorded from the same 8 electrodes positions (C3, Cz, C4, CPz, P3, Pz, P4, POz) with all the systems.
-The reference positions were placed at the right ear.
-The ground positions was placed at the left ear, except with the g.Nautilus system (placed at AFz, because this could not be changed).
-Sampling rate was set at 250 Hz, the closest possible to predefined g.Mobilab sampling rate (256 Hz).
-Data were acquired notch filtered at 50 Hz and passband filtered between 2 Hz and 30 Hz.
-The filtering was performed via the proprietary Simulink HighSpeed Online Processing block modules for g.MOBIlab and g.Nautilus systems, and via MATLAB code for the Xpress system.
-Table~\ref{tab:results} summarizes the acquisition settings for each system.
+Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same locations Fz, Cz, Pz, P3,P4, Oz, PO7 and PO8.
+Reference is set to the right ear lobe and ground is preset as the AFz position.
+Sampling frequency is slightly different, and is set to 250 Hz, which is the closest possible to the one used with the other dataset.
+
+%Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.
 
@@
 
-% Antes estaba esto aca
-
-% The P300 event latency and amplitude vary greatly between subjects,
-
-% lo saque porque confunde, ya que despues de todo, dejamos todo fijo, e ignoramos completamente el hecho de que varie.
-
-%Parameters are selected according to the experimental protocol. 
-The patch size is $S_P = 3s.4 \times 3s.4$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
-The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $.
+The patch size is $S_P = 4.3s \times 4.3s$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $~\citep{Rao2013}.
 Hence it is necessary to utilize a size patch $S_P$ that could capture an entire transient event.
 With this purpose in consideration, the $s$ value election is essential.
+
 %necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
-We propose the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
+We propose the Equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
 \begin{equation}
@@
 
-\noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
+\noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sampling frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
 The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.
-This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds in the horizontal direction.
-Finally, descriptors  $\mathbf{kp}$  are located at $ x_{kp} =  0.55 Fs.\gamma = 35 $ and $ y_{kp} = z(c) $ (see Eq.~\ref{eq:zerolevel}).
+This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ on the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds on the horizontal direction.
+Descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ for the corresponding channel $c$ (see Eq.~\ref{eq:zerolevel}).
 In this way the whole transient event is captured.
 Figure~\ref{fig:sampledescriptor2} shows a patch of a signal plot covering the complete amplitude (vertical direction) and the complete span of the signal event (horizontal direction).
-% esta figura no está comentada en el texto.
+
 \begin{figure}[H]
@@
 \includegraphics[width=10cm]{patchgeometry.pdf}
-\caption{The scale of local patch is selected in order to capture the whole signal information}
+\caption{The scale of local patch is selected in order to capture the whole transient event.
+The size of the patch is $S_p \times S_p$ pixels.
+The vertical size consists of $4$ blocks of size $3 s_y$ pixels which is long enough as to contain the signal $\Delta  \mu V $, the peak-to-peak amplitude of the transient event.
+The horizontal size on the other hand, includes $4$ blocks  of $3 s_x$ and covers the entire duration in seconds of the transient signal event, $ \lambda $.
+}
+
 \label{fig:sampledescriptor2}
@@
 
-%Hacemos algún comentario de k_a?
+Lastly, the number of channels $Ch$ is equal to $8$ for both datasets, and the number of intensification sequences $k_a$ is statically assigned to $10$.
+The parameter $k$ used in the Near Neighbor $NN_T(d^l_u), \;l\in\{row,{col}\}$ is set to $k=7$, following the suggestion of the article~\citep{Boiman2008}.
+In addition, the norm used on  Equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol} was the cosine norm, and descriptors were normalized to $ \left[ -1, 1 \right] $.
 
@@
 \label{section:results}
-In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects using the information from all the segments can be shown.
+
+%In Figure \ref{fig:subjectaveraged} the grand average (point-to-point) for all the subjects using the information from all the segments can be shown.
 The P300 characteristic curve can be seen particularly in subjects 2, and 6 and in a lesser extend in the remaining subjects.
@@
 an unequal number of epochs on each condition), because that may introduce bias in the classification procedure (the variance of averaged signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).
+%
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=16cm]{subjectaveraged.eps}
+%\caption{Point-to-point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
+The P300 characteristic curve can be well identified particularly on subjects 2 and 6.}
+%\label{fig:subjectaveraged}
+%\end{figure}
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=16cm]{subjectaveraged.eps}
-\caption{Point-to-point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
-The P300 characteristic curve can be well identified particularly on subjects 2 and 6.}
-\label{fig:subjectaveraged}
-\end{figure}
+The P300 ERP  consists of two overlapping components: the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\citep{Polich2007} then, the classical approach finds the stronger response on the central channel Cz~\citep{Riccio2013}.
+However, Wolpaw et al~\citep{Krusienski2006} show that the response may also arise in occipital channels.
+In our approach, occipital channels PO8 and PO7 show higher performances for some subjects.
+%For example in the articles~\citep{Huggins2016,Jure2016} the authors show that channels PO8 and PO7 have higher performances.
 
-Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
-From the 7 words for each subject, the first 3 were used as training, and the remaining 4 for testing.
+%The authors~\citep{Riccio2013} that analyzed the public dataset concluded that Cz is the best performing channel, more representative of the difference in amplitudes.
+However, in our approach, occipital channels,  PO8 and PO7 show higher performances~\citep{Huggins2016,Jure2016}.
+
+Table~\ref{tab:results} shows the results of applying the algorithm to the subjects of the public dataset of ALS patients~\citep{Riccio2013}.
+The percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
+From the seven words for each subject, the first three are used as training, and the remaining four for testing.
 The best performing channel is informed as well.
-
-DRAFT: Traditionally P300 can be seen in this channel and this other channel.
-However, Wolpaw showed that the response may also arise in occipital channels as well.
-The authors of the original paper concluded that Cz was the best performing channel.
-However, in our approach
-
-It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
+The chance level is $2\%$.
+It can be observed that the best performance of the letter identification method is reached in various channels depending on the subject on study.
 
 \begin{table}[H]
-\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject. Chance level is 0.02 }
+\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject of the public dataset 008-2014.
+Twenty letters, four words of five letters each, were used for testing.
+}
 \centering
@@
 \midrule
-1     &     Cz   &   0.35  \\
-2     &     Fz   &   0.85  \\
-3     &     Cz   &   0.25  \\
-4     &     PO8 &   0.55  \\
-5     &     PO7 &   0.40 \\
-6     &     PO7 &   0.60  \\
-7     &     PO8 &   0.80  \\
-8     &     PO7 &   0.95 \\
+1     &     Cz   &   $35\%$  \\
+2     &     Fz   &   $85\%$  \\
+3     &     Cz   &   $25\%$  \\
+4     &     PO8 &   $55\%$  \\
+5     &     PO7 &   $40\%$ \\
+6     &     PO7 &   $60\%$  \\
+7     &     PO8 &   $80\%$  \\
+8     &     PO7 &   $95\%$ \\
 
@@
 
-The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.
-The Performance curves (Fig.
-\ref{fig:performance}) show how the percentage of correctly identified letters depends on the number of repetitions that were used to obtain the averaged signal.
+In Table~\ref{tab:resultsowndataset} results obtained for 8 healthy subjects are shown.   These results were obtained by the present method as described in previous sections.
+
+%We used an additional subject as control as he was instructed to count only the letters and to stare at the screen but  neglecting row/column flashing, which achieved zero performance.
+We also validated the procedure by visualizing clearly the SSVEP generated by the 4 Hz flashings.
+
+\begin{table}[H]
+\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each healthy subject.
+The method is applied on a channel by channel basis, and the best performing channel is shown.
+}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{cccc}
+\toprule
+\textbf{Participant}	&  \textbf{BPC}	& \textbf{Performance}\\
+\midrule
+1     &     Oz   &     $40\%$  \\
+2     &     PO7   &     $30\%$ \\
+3     &     P4   &     $40\%$ \\
+4     &     P4   &     $45\%$ \\
+5     &     P4   &      $60\%$ \\
+6     &     Pz   &      $50\%$ \\
+7     &     PO7   &      $70\%$ \\
+8     &     P4   &      $50\%$ \\
+
+\bottomrule
+\end{tabular}
+\label{tab:resultsowndataset}
+\end{table}
+
+The Information Transfer Rate (ITR), or Bit Transfer Rate (BTR), in the case of reactive BCIs~\citep{WolpawJonathanR2012}  depends on the amount of signal averaging required to transmit a valid and robust selection.
+Fig.~\ref{fig:performance} shows the performance curves for varying intensification sequences.
+It can be observed that the percentage of correctly identified letters depends on the number of intensification sequences $k_a$ that are used to obtain the averaged signal.
 
@@
 \centering
-\includegraphics[width=10cm]{performance.eps}
-\caption{Performance curves for the eight subjects included in the published dataset.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
+\includegraphics[width=10cm]{performance2.eps}
+\caption{Performance curves for the eight subjects included in the dataset of ALS patients.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
 \label{fig:performance}
@@
 
-We found that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that could allow the classification to reach a higher level.
-We hypothesize that as subjects may have different latencies and amplitudes of the P300 signal \citep{Riccio2013}, it may also be the case that the shape of the generated ERP may vary greatly in an intra-subject manner, thus the pattern could not be well generalized and a higher performance could not be reached (Fig.
-\ref{fig:p300templates}).
+As can be seen in the figure, as the number of intensification sequences tend to 1, which corresponds to single trial letter identification, the performance is reduced.
+As mentioned before, the SNR of the single trial P300 is very low.
+The ensemble average is carried out in order to improve the SNR but other problems arise, for example inter-trial variability due to latencies shifts that may produce an unstable P300 signature.
+To solve this problem, we tested an approach to assess if the morphological shape of the P300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied, but results are still in process.
+We also applied Dynamic Time Warping (DTW)~\citep{Casarotto2005} but we were unable to find a substantial improvement.
+On the other hand amplitude, width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals\citep{Ouyang2017}.
+This is another source of instability of the P300 signature component that may need to be addressed.
+
+As subjects may have different \textit{latencies}, \textit{amplitudes} and \textit{width} of their P300 components, they may also have distinct \textit{shapes} of the generated ERP.
+Figure~\ref{fig:p300templates} shows the P300 templates patches for patients 8 and 3 from the public dataset.
+It can be observed that in coincidence with the performance results, the P300 signature is more clear and consistent for subject 8 (Fig.~\ref{subject8})  while for subject 3 (Fig.~\ref{subject3}) the characteristic pattern is more difficult to perceive.
 
@@
 \centering
-\includegraphics[width=10cm]{subject8.png}
-\includegraphics[width=10cm]{subject3.png}
-\caption{Ten P300 template patches found for subjects 8 (up) and 3(down).
-In coincidence with the performance results, the P300 signature is more clear and consistent for subject 8 (higher performance) while for subject 3 (lower performance) the characteristic pattern is much more difficult to perceive.}
+\subfigure[Subject 8.]{\includegraphics[width=10cm]{subject8.png}\label{subject8}}
+\subfigure[Subject 3.]{\includegraphics[width=10cm]{subject3.png}\label{subject3}}
+\caption{P300 template patches for subjects 8 and 3. As traditional done in neuroscience research, downward is positive. }
 \label{fig:p300templates}
 \end{figure}
+
+%\begin{itemize}
+%\item Lo que pusimos antes de la sincronización de las señales apunta a la latencia.
+%\item Por el lado de la amplitud, lo que hicimos fue ir por un lado conservador de llevar todo a varianza uno.
+%\item Si bien el zscore nos ayuda también genera otro problema adicional que es el que ponemos en el texto (y por el cual con razón Juan me criticaba).
+
+%\end{itemize}
+
+%La estandarizacion produce dos efectos contrarios entre si, contrapues.
+
+%Por un lado reduce el ruido y outliers
+%por otro lado reduce aquellos p300 significativos positivos que ayudarian su identificacion.
+
+Another problem is the amplitude variation of the P300.
+We propose an approach by standardizing the signal, shown in Eq.~\ref{eq:standarizedaverages}.
+First, it has the effect of normalizing the peak-to-peak amplitude, moderating its variation.
+It has also the advantage of reducing noise that were not reduced by the averaging procedure.
+It is important to remark that the signal variance depends on the number of single-trials used to compute it \citep{van2006signal}.
+The standardizing process converts the signal to unit signal variance which makes it independent of the number $k_a$ of signals averaged.
+This is another advantage of this approach.
+On the downsize, the standardizing process reduces the amplitude of any significant P300 complex diminishing its automatic interpretation capability.
+
+For both datasets, the experimental protocol uses a very short inter-stimulus interval which has the potential to increase the ITR but at the same time it reduces the amplitude of the P300 response, hence it may be more difficult to detect it~\citep{Rao2013}.
+
+%As an unexpected consequence, $15$ out of $20$ subject reported high ocular tiredness by using this configuration.
 
@@
 %
-%Compare if it is possible with the descriptors from one subject, discriminate the others.
+%Compare if it is possible with the descriptors from one subject, discriminate the others.  Transfer learning.
 %
@@
 \section{Conclusion}
+\label{conclusion}
 
-A method to characterize and classify EEG signals where their characterization is transient in time-space, like the P300 ERP, has been presented.
+%In this paper, a new unsupervised method to enhance evoked response by target stimuli in an oddball paradigm was presented.
+Only given the time indexes of rows/columns intensifications, the proposed algorithm estimates the main components of the P300 subspace by providing the best SNR.
+It was shown to efficiently improve the quality of the evoked responses by taking into account the signal and the noise, as opposed to principal component analysis, which only considers the signal.
+Using this method to enhance P300 subspace before the BCI classification task speeds up the BCI since less words are required to train the spatial filters and the linear classifier, given a certain percentage of good symbol prediction.
+Moreover, using this spatial enhancement significantly reduces the dimension of the feature vector used to predict words.
 
-%Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).
+%RECORDAR CALIBRATION STEP
 
-The adaptive behaviour of the algorithm make it well suited when the shape of the pattern elicited by the P300 response does not conform to the predicted structure.
-This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e.
-\textit{how actually they looked like}) for the training and calibration step, and they do not require any prior knowledge about the signal.
-In contrast, when the shape of the ERP response is not consistent across the same subject, this method would not be able to find the templates and speller performance could be penalized.
+Among other applications of BCI analysis, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
 
-At the same time, by analyzing the generated descriptors, which map in a very detailed and synthetic structure the shape information contained within the patch, a metric about the consistency of the shape of the generated P300 response could also be derived.
-It may be worthy of further consideration to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
+In this work, a method to detect transient P300 components from EEG signals based on their waveform characterization in digital time-space, is presented.
+Additionally, its validity is evaluated using a public dataset of ALS patients and an own dataset of healthy subjects.
+The objective of this article is to explore new techniques of P300 automatic interpretation.
+We know that we have to keep working but the results are promising.
 
-We believe that the expanding and the understanding of this tool in order to automatically classify those patterns in EEG that are specifically identified by their shapes (e.g.
-K-Complex, Vertex Waves, Positive Occipital Sharp Transient \citep{Hartman2005}) is a prospect future work to be considered.
-It may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods, frequent in sleep research.
+The method works on a channel by channel basis; in this way the best performing channel can be identified and used it to reduce the number of required EEG electrodes, leading to the development of more ergonomic capturing device.
 
-Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc}, i.e.
-to detect BCI illiteracy or to predict the achievable performance of a given method, or even as a tool for artefact removal (which is performed on many occasions by visually inspecting the signal).
+We observed that the shape of the wave is more stable in occipital channels, where the performance for identifying letters is higher.
+% It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
+It is important to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
+Although,  we did not find any evidence in terms of the shape digital signal and the best response was obtained for an ALS subject.
+Moreover, this method can be used as an alternate BCI predictor.
+
+By analyzing the generated descriptors, a metric about the shape of the P300 response can also be derived.
+
+By means of the empirical experiments, we concluded that the stability of the P300 in terms of shape is crucial: synchronization averaging,  inter-stimulus interval,  montages, the signal to noise ratio and spatial filters can all of them affect the stability of the shape of the P300 ERP.
+
+In our opinion, the best application of this approach is that a closer collaboration with physicians can be fostered, because our method intent to imitate human visual observation.
+Automatically classify those patterns in EEG that are specifically identified by their shapes (e.g.
+K-Complex, Vertex Waves, Positive Occipital Sharp Transient~\citep{Hartman2005}) is a prospect future work to be considered.
+We are currently working in unpublished material analyzing KComplex that also provide  assistance to physician to locate these EEG patterns, specially in long recording periods, frequent in sleep research.
+Additionally, it can be used for artifact removal which is performed on many occasions by visually inspecting the signal.
+This is due to the fact that the descriptors are directly based on the signals behavior in shape domain.
+In line with these applications,  it can be used to build a database~\citep{Chavarriaga2017} of descriptors and improve atlases~\citep{Hartman2005}.
+
+We also want to solve the problem of signal stabilization by applying different shifts on the averaging procedure and another methods for obtaining a correct synchronization.
+
+%Finally, waveform analysis and the methods that can be used to analyze them is an area where active research can benefit greatly both, physicians who need to automate some of the tools and practices and Brain Computer Interfaces or automatic EEG processing by using the extensive knowledge body available for clinical practices~\citep{Cole2017}
+
+%To further improve the performance of the P300-speller BCI, additional work should be considered.
+For instance, to better estimate the response evoked by target stimuli, a multi- stimuli model should be deemed by assuming that all the non- target stimuli – as well as the target stimuli – evoked specific responses.
+Finally, sensor selection should also be considered, in order to drastically reduce the number of required EEG electrodes, leading thus to a more ergonomic BCI.
 
@@
 SIFT: Scale Invariant Feature Transform\\
-NBNN: Naive Bayes Nearest Neighbor\\
-HOG: Histogram Of Gradients}
+NBNN: Naive Bayes Nearest Neighbor}
 

### 2017-12-13 response.letter.tex
Saque finalmente hit y saque el grafico de epoching de los resultados.  Cambie el orden de los pasos, la descripcion del dataset own paso a estar en presente (suena rara), agregue referencias a la expansion de este metodo para otras casos, ajuste el epigrafe de lo que tendria que tener sentido mostrar en un grafico (el grafico por ahora lo deje lo voy a modificar en base al epigrafe) agregue porcentajes en los resultados, puse tambien los resultados hasta ahora con el dataset nuestro, aclare los sujetos de ese dataset, quienes son y que edades tienen, elimine la referencia al zscore, modifique

@@
 We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and to decode SSVEP patterns.
-Modificar%We modified the article to emphasize more clearly that the presented algorithm is  applicable to P300.
+We added this information on the article (section INTRODUCTION and section CONCLUSION).
+We hope this new information enrich the article and clarifies this point.
 
-% Como haces para que las referencias que estan aca en la respuesta apunten al articulo ???
-%An explanation  of which is a satisfactory result for Algorithm~\ref{Algoritmo} was added in sections~\ref{errorevaluation} and~\ref{sec:EdgeSimulatedData}.  
 }
@@
 Section FEATURE was modified as well as section PARAMETERS.
-Section SIGNAL PLOT was revised and its mathematical terms were corrected.}
+Section SIGNAL PLOT was revised and its mathematical terminology were corrected.}
 \end{quotation}

### 2017-12-13 article.tex
Saque finalmente hit y saque el grafico de epoching de los resultados.  Cambie el orden de los pasos, la descripcion del dataset own paso a estar en presente (suena rara), agregue referencias a la expansion de este metodo para otras casos, ajuste el epigrafe de lo que tendria que tener sentido mostrar en un grafico (el grafico por ahora lo deje lo voy a modificar en base al epigrafe) agregue porcentajes en los resultados, puse tambien los resultados hasta ahora con el dataset nuestro, aclare los sujetos de ese dataset, quienes son y que edades tienen, elimine la referencia al zscore, modifique

--- 2017-12-11 article.tex
@@
 
-% Current address and/or shared authorship
-\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina}
+% Current address and/or shared authorship  COMMENTED PER REVIEWER
+%\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina}
+
+\firstnote{.}
 
@@
 % Abstract (Do not use inserted blank lines, i.e. \\) 
-\abstract{The analysis of Electroencephalographic (EEG) signals is of ulterior importance to elucidate patterns that could improve the implementation of Brain Computer Interfaces (BCI).
+\abstract{EL OBJETIVO DE ESTE PAPER, EL HILO CONDUCTOR, ES DEMOSTRAR QUE ESTE METODO QUE ESTA BASADO EN LA ESTRUCTURA DE LA SENIAL SE PUEDE UTILIZAR TAMBIEN PARA IDENTIFICAR EVENTOS TRANSIENTES COMO EL P300, Y QUE CON ESO SE PUEDE IMPLEMENTAR UN SPELLER DE P300.
+The analysis of Electroencephalographic (EEG) signals is of ulterior importance to elucidate patterns that could improve the implementation of Brain Computer Interfaces (BCI).
 These systems are meant to provide alternative pathways to transmit volitional information which could potentially enhance the quality of life of patients affected by neurodegenerative disorders or improve Human Computer Interaction systems.
@@
 It also aims to provide a framework to analyze, characterize and classify EEG signals, with a focus on the P300, an ERP elicited by the oddball paradigm of rare events.
-The validity of the method is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients.}
+The validity of the method is shown by offline processing a public dataset of Amyotrophic Lateral Sclerosis (ALS) patients and an own dataset for healthy subjects.}
 
 % Keywords
-\keyword{electroencephalography (EEG); BCI; P300; ALS; NBNN; HOG; SIFT}
+\keyword{EEG; BCI; P300; ALS; NBNN; HOG; SIFT}
 
@@
 
-%\begin{enumerate}[leftmargin=*,labelsep=3mm]
-%\item	Why
-%\item	Purpose and significance
-%\item	Current state of the research field and key publications cited
-%\item   Controversial hypothesis
-%\item   Main aim and principal conclusions.
-%\end{enumerate}
-%
-%\begin{enumerate}[leftmargin=*,labelsep=3mm]
-%\item El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
-%\item La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
-%\item Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
-%\item La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
-%\item La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
-%\item La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
-%\item Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
-%\item ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).
-%\end{enumerate}
-
 Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundred years old technology.
@@
 
-In~\citep{Ramele2016},  authors introduce a method for classification of rhythmic patterns using gradient histograms orientations.
-Inspired in that work, we proposed a novel application of the developed method to classify and describe transient events like those produced by P300 Event Related Potential.
- Its validity is verified by processing offline data for ALS patients.  
+In~\citep{Ramele2016},  authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves~\citep{Ben-Simon2013}  and Motor Imagery rolandic central $\mu$ rhythms~\citep{Hartman2005} using gradient histograms orientations.
+Inspired in that work, we proposed a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.
+Its validity is verified by processing offline data for ALS patients and for healthy subjects.
 The proposed approach is based on the morphological analysis of the shape of the EEG signal~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and it was built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting raw signal plots~\citep{Hartman2005}.
@@
 
-%\subsection{Morphological Signal Analysis}
-%
-%Several approaches and methods have been applied to decode them.  Of particular interests are morphological signal analysis 
-%CITE (Paper del chabon del ITBA que detalla bien los diferentes metodos) \citep{Alvarado-Gonzalez2016,Yamaguchi2009}.
-%
-%Morphological or 
-%
-%template matching
-%
-%spike
-%sharp wave
-%spike wave
-%polyspike 
-%slow wave complex
-%
-%shape domain
-%
-%\subsection{P300}
-%
-%Uhhhh I can talk a lot about P300 \citep{Knuth2006}.
-%
-%P300 single trial is a golden grail of Brain computer interfaces.  It is studied and analyzed in this way, here in this other way, in this other way is here.
-
 \section{Materials and Methods}
-
-\subsection{Feature Extraction based on Histogram of Gradients of the Signal Plot} \label{Feature}
 
@@
 
-%The proposed algorithm was designed to perform specifically this decoding.
+\subsection{Feature Extraction based on Histogram of Gradients of the Signal Plot} \label{Feature}
 
-% of the selected letter based on the detection of the P300 Event Related Potential response on the corresponding EEG signal.
-
-\subsubsection{Preprocessing} \label{Preprocessing}
-
-The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG. 
-The processing pipeline starts by applying a notch filter to the raw signal, a 
-$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.
+The underlying idea of this method is to generate templates for the signal \citep{Ramele2016}, based on the image plot.
+Hence, the first step is the signal transformation into a temporary binary image.
 
@@
 \begin{itemize}
+\item \textbf{Preprocessing}: The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG.
+The processing pipeline starts by applying a notch filter to the raw signal, a
+$4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.
 \item \textbf{Artefact Removal}: The EEG signal matrix is processed on a channel by channel basis.
@@
 \item \textbf{Segmentation}: For each of the $12$ stimuli,  a window of $1$ second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
-Two of these segments are labeled as \textit{hit}, whereas the remaining $10$ are labeled as \textit{no hit}.
-A hit represents that the EEG segment should contain the P300 ERP signature time-locked to the flashing stimulus.
+Two of these segments should contain the P300 ERP signature time-locked to the flashing stimulus, one for the row, and one for the column.
 \item \textbf{Signal Averaging}:  The P300 ERP is deeply buried under background EEG so the traditional approach to identify it is by point-to-point averaging the time-locked stacked signal segments.
@@
 And, at the same time, as long as more repetitions are needed, the ability to transfer information faster is diminished, so there is a trade-off that must be acutely determined.
-
-\subsubsection{Signal Plotting} \label{Plot}
-
-The underlying idea of this method is to generate a template for the signal \citep{Ramele2016},  based on the image plot.
-Hence, the first step is the signal transformation into a temporary binary image.
-
-Averaged signal segments obtained from the preprocessing stage are first scaled and standardized (i.e. z-score) by
-
-\begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor
-\end{equation}
-
-\noindent where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for  channel $c$.
-Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
-
-Consequently, the image is constructed by placing the sample points according to
-
-\begin{equation}
-I(z_1,z_2) = \left\{ \begin{array}{rl}
-255 & \text{if} \,  z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
-0   & \mbox{otherwise}
-\end{array}\right.
-\label{eq:images}
-\end{equation}
-
-\noindent where $ z_1$ and $z_2$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image.
-The function $z(c)$ is the \textit{zerolevel} which is the location on the image where the signal's zero value should be located in order to fit the entire signal within the image:
-
-\begin{equation}
-z(c) = \left \lfloor{ \frac{\max \tilde{x}(t,c)  - \min \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max \tilde{x}(t,c)  + \min \tilde{x}(t,c)}{ 2} }\right \rfloor
-\label{eq:zerolevel}
-\end{equation}
-
-In order to complete the plot from the pixels, the Bresenham \citep{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of  consecutive pixels.
-
-%\noindent and is used to derive local representations that captures the visual shape of the signal.
 
@@
 
-%Este párrafo no se entiende nada
-%------------------
+\subsubsection{Signal Plotting} \label{Plot}
 
-%------------------
+Averaged signal segments are scaled and standardized by
+
+\begin{equation}
+\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor
+\end{equation}
+
+\noindent where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for  channel $c$.
+Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
+
+Consequently, the image is constructed by placing the sample points according to
+
+\begin{equation}
+I(z_1,z_2) = \left\{ \begin{array}{rl}
+255 & \text{if} \,  z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
+0   & \mbox{otherwise}
+\end{array}\right.
+\label{eq:images}
+\end{equation}
+
+\noindent where $ z_1$ and $z_2$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image.
+The function $z(c)$ is the \textit{zerolevel} which is the location on the image where the signal's zero value should be located in order to fit the entire signal within the image:
+
+\begin{equation}
+z(c) = \left \lfloor{ \frac{\max \tilde{x}(t,c)  - \min \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max \tilde{x}(t,c)  + \min \tilde{x}(t,c)}{ 2} }\right \rfloor
+\label{eq:zerolevel}
+\end{equation}
+
+In order to complete the plot from the pixels, the Bresenham \citep{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of  consecutive pixels.
+
 \subsection{Experimental Protocol} \label{Protocol}
@@
 
-We replicated the same experiment on healthy subjects using a wireless digital EEG device (g.Nautilus, g.Tec, Austria).
-The experimental conditions were the same as those used for the previous dataset, as detailed in section~\ref{ALSDataset}.
+We replicate the same experiment on healthy subjects using a wireless digital EEG device (g.Nautilus, g.Tec, Austria).
+The experimental conditions are the same as those used for the previous dataset, as detailed in section~\ref{ALSDataset}.
 
-Participants were recruited voluntarily and the experiment were conducted anonymously in accordance with the declaration of Helsinki published by the World Health Organization.
-No monetary compensation was handed out and all participants agreed and signed a written informed consent.
-All healthy subjects had normal or corrected-to normal vision and no history of neurological disorders.
-The experiment was performed with 20 subjects, 7 males, 4 females, average age 22.5 years, standard deviation  1,8 years, range 21–26 years.
+Participants are recruited voluntarily and the experiment is conducted anonymously in accordance with the declaration of Helsinki published by the World Health Organization.
+No monetary compensation is handed out and all participants agree and sign a written informed consent.
+All healthy subjects have normal or corrected-to normal vision and no history of neurological disorders.
+The experiment is performed with 20 subjects, 17 males, 3 females, average age 26.85 years, standard deviation  8.8 years, range 19-50 years.
 
-Gel-based active electrodes (g.LADYbird, g.Tec, Austria) were used on the same locations Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.
-Reference was set to the right ear lobe and ground was preset as the AFz position.
-Sampling frequency is slightly different, and was set to 250 Hz, which is the closest possible to the one used with the other dataset.
+Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same locations Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.
+Reference is set to the right ear lobe and ground is preset as the AFz position.
+Sampling frequency is slightly different, and is set to 250 Hz, which is the closest possible to the one used with the other dataset.
 
@@
 \includegraphics[width=10cm]{patchgeometry.pdf}
-\caption{The scale of local patch is selected in order to capture the whole signal information}
+\caption{The scale of local patch is selected in order to capture the whole signal information.
+The vertical size $4 \cdot 3 s_y$ of the patch is required to be long enough as to contain the signal $\Delta  \mu V $, which is the peak-to-peak amplitude of the transient event.
+The horizontal size $ 4 \cdot 3 s_x$ , on the other hand, needs to cover the entire duration in seconds of the transient signal event, $ \lambda $.
+}
 \label{fig:sampledescriptor2}
@@
 an unequal number of epochs on each condition), because that may introduce bias in the classification procedure (the variance of averaged signals is inversely proportional to the number of samples and the procedure would be discriminating signals with different variances).
+%
+%\begin{figure}[H]
+%\centering
+%\includegraphics[width=16cm]{subjectaveraged.eps}
+%\caption{Point-to-point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
+The P300 characteristic curve can be well identified particularly on subjects 2 and 6.}
+%\label{fig:subjectaveraged}
+%\end{figure}
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=16cm]{subjectaveraged.eps}
-\caption{Point-to-point grand averages of epochs obtained for hits (solid line) and no hits (dashed line) for each one of the 8 subjects for channel Cz.
-The P300 characteristic curve can be well identified particularly on subjects 2 and 6.}
-\label{fig:subjectaveraged}
-\end{figure}
-
-Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
+Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation for the public dataset of ALS patients.
 From the 7 words for each subject, the first 3 were used as training, and the remaining 4 for testing.
@@
 However, Wolpaw showed that the response may also arise in occipital channels as well.
-The authors of the original paper concluded that Cz was the best performing channel.
-However, in our approach
-
-It is of particular interest that using this method, the best performing channel was not always Cz, and instead occipital channels PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
+The authors of the original paper concluded that Cz was the best performing channel, more representative of the difference in amplitudes.
+However, in our approach occipital channels,  PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
 
 \begin{table}[H]
-\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject. Chance level is 0.02 }
+\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject. Chance level is $2\%$. }
 \centering
@@
 \ref{fig:performance}) show how the percentage of correctly identified letters depends on the number of repetitions that were used to obtain the averaged signal.
+For both datasets, the experimental protocol uses a very short ISI which has the potential to increase the ITR but at the same time is inversily proportional to the amplitude of the p300 response (i.e.
+harder to detect).
+As an unexpected consequence, 15 out of 20 subject reported high ocular tiredness by using this configuration.
 
@@
 We found that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that could allow the classification to reach a higher level.
+Moreover, as can be seen in the figure, as repetitions tend to 1, which will correspond to single trial letter identification, the performance suffers greatly as the signal plot is simple too complex to be used as template and the method is completely unable to find good matches  PODRIA ACA MOSTRAR ADEMAS PLOTS DE SINGLE TRIAL PARA QUE VEAN QUE SON UN QUILOMBO.
+
+UNA DE LAS RAZONES POR LA CUAL PODEMOS LLEGAR A ARGUMENTAR QUE NO FUIMOS MAS A FONDO CON EL TEMA ES QUE NO QUERIAMOS DISTORSIONAR DEMASIADO LA SENAL COMPLEJIZANDO EL ALGORITMO Y QUERIAMOS MANTENERNOS ENTRE LO MAS ESTANDAR POSIBLE (PUEDO CITAR MAS O MENOS QUE NUESTRO PROCESSING PIPELINE ES ESTANDARD).
+
 We hypothesize that as subjects may have different latencies and amplitudes of the P300 signal \citep{Riccio2013}, it may also be the case that the shape of the generated ERP may vary greatly in an intra-subject manner, thus the pattern could not be well generalized and a higher performance could not be reached (Fig.
 \ref{fig:p300templates}).
+It is known (REFERENCE) that the point-to-point averaging suffers from inter-trial variability (REFERENCIA A ESE PAPER Con ese nombre) due to latencies shifts.
+We tested a naive approach to verify if the morphological shape of the p300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied.
+We also applied DTW but couldn't found a sustantial improvement.
+
+Additionally we believe that our method suffers also from null signals.
+
+ACA EL PUNTO PARA MOSTRAR ES QUE DE ALGUNA MANERA ESTAMOS COMO MOSTRANDO UN "SI SE PUEDE" DE UN TRABAJO QUE PUEDE CONTINUAR Y EXPANDIRSE.
+POR ESO LA IDEA DE MOSTRAR PUNTAS DE COSAS QUE VIMOS E HICIMOS, PERO SIN MOSTRAR RESULTADOS CONCRETOS.
+
+Considering the amplitude variation of the p300 and its components, we proposed a conservative approach by standarizing the signal (which can have both effects, of reducing noise that were not reduced by the averaging procedure but at the same time it can reduce the amplitude of a significant p300 compoent).
+The standarization also provides another benefit which is sometimes neglected as unimportant.
+The variance of the averaged signal depends on the number of signals averaged.
+As this method analyze precisely the shape of the signal in the plot, which is very affected by the variance, the method will not be able to discriminate solely based on the underlying cognitive event, it will be confused by the number of signals that were used to calculate the averaged signal.
+
+The SIFT method is invariant to the scale, but in order to allow our method to be also invariant to the amplitude of the signal we had to modify the geometry of the descriptor to remove the squared constrain.
 
@@
 \end{figure}
+
+\begin{table}[H]
+\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each healthy subject. Chance level is $2\%$. }
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{cccc}
+\toprule
+\textbf{Participant}	&  \textbf{BPC}	& \textbf{Performance}\\
+\midrule
+1     &     Cz   &     0.7000  &  0.3500  \\
+1     &     Cz   &    0.1000  &  0.0500 \\
+1     &     Cz   &     0.1000  &  0.1000 \\
+1     &     Cz   &     0.1000  &  0.1500 \\
+1     &     Cz   &     0.1000  &  0.1000 \\
+1     &     Cz   &     0.5500  &  0.4000 \\
+1     &     Cz   &     0.1500   & 0.1000 \\
+1     &     Cz   &     0.4500   & 0.2000 \\
+1     &     Cz   &     0.3000  &  0.2000 \\
+1     &     Cz   &     0.2500  &  0.1500 \\
+1     &     Cz   &     0.5500  &  0.3000 \\
+1     &     Cz   &     0.3500  &  0.1000 \\
+1     &     Cz   &     0.3000  &  0.3500 \\
+1     &     Cz   &     0.5000  &  0.4000 \\
+1     &     Cz   &     0.3000  &  0.1500 \\
+1     &     Cz   &     0.3000  &  0.3000 \\
+1     &     Cz   &     0.3500  &  0.3500 \\
+1     &     Cz   &          0      & 0.0500 \\
+1     &     Cz   &     0.7000   & 0.5000 \\
+1     &     Cz   &     0.4000  &  0.3000 \\
+
+\bottomrule
+\end{tabular}
+\label{tab:resultsowndataset}
+\end{table}
+
+In Table \ref{tab:resultsowndataset} results obtained for the 20 healthy subjects is shown.
+The third column shows the results obtained by using the Krusienski feature and the SVM classifier.
+On the other hand, the fourth column shows the results obtained by the present method as described in previous sections.
+Subject 17 was used as control as he was instructed to count only the letters and to stare at the screen but not neglected row/column flashing.
 
@@
 K-Complex, Vertex Waves, Positive Occipital Sharp Transient \citep{Hartman2005}) is a prospect future work to be considered.
-It may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods, frequent in sleep research.
+We are currently working in unpublished material analyzing KComplex that may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods, frequent in sleep research.
 

### 2017-12-13 article.tex
Subo los cambios de Juliana.  Agregué el detalle de cómo fue la sesión de captura del dataset que tomamos.

@@
 This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained.
-Section \ref{Preprocessing} and \ref{Pipeline} describe the processing pipeline.
+Section~\ref{Pipeline} describe the processing pipeline.
 Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.
@@
 
+\subsubsection{Signal Plotting} \label{Plot}
+
+Averaged signal segments are scaled and standardized by
+
+\begin{equation}
+\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, t\geq 0,c \in \{1,2,\dots,8\}
+\end{equation}
+
+\noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related with the image scale.
+On the other hand, $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for channel $c$.
+Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
+
+Consequently, the image is constructed by placing the sample points according to
+
+\begin{equation}
+I(z_1,z_2) = \left\{ \begin{array}{rl}
+255 & \text{if} \,  z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
+0   & \mbox{otherwise}
+\end{array}\right.
+\label{eq:images}
+\end{equation}
+
+\noindent where $ (z_1,z_2) \in \mathcal{N} \times \mathcal{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image.
+The function $z(c)$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
+
+\begin{equation}
+z(c) = \left \lfloor{ \frac{\max \tilde{x}(t,c)  - \min \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max \tilde{x}(t,c)  + \min \tilde{x}(t,c)}{ 2} }\right \rfloor
+\label{eq:zerolevel}
+\end{equation}
+
+In order to complete the plot from the pixels, the Bresenham \citep{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of  consecutive pixels.
+
 \subsubsection{Feature Extraction: Histogram of Oriented Gradients}
@@
 
-On the generated image, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered.
-A local image patch of size $S_p \times S_p$ is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+On the generated image $I$, a keypoint $\mathbf{kp}$ is placed on a pixel $(x_{kp}, y_{kp})$ over the image plot and a window around the keypoint is considered.
+A local image patch of size $S_p \times S_p$ pixels is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
 It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, then $S_p = 4.3s $ pixels.
@@
 \begin{enumerate}
-\item Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
+\item Highlight randomly the rows and columns from the matrix.
+There is one row and one column that should match the letter selected by the subject.
+For each illuminated row or column, an image is obtained.
 \item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively.
@@
 \begin{equation}
-x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c)
+x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c), t\geq 0, c \in \{1,\dots,8\}
 \label{averaging}
 \end{equation}  
+for each row and column.
 
-\item Plot the signal $x(t,c)$, according Section~\ref{Plot}.
+\item Plot the signal $x(t,c)$, according Section~\ref{Plot}. The images $I^{row}_1, \dots, I^{row}_6$ and $I^{col}_7,\dots,I^{col}_{12}$ are obtained for rows and columns, respectively.
 
-\item Obtain the descriptor from the image of the signal plot of $x(t,c)$ in accordance to the method described in Section~\ref{SIFT}, $ d^r_1, \dots,  d^r_6 $ for rows and  $ d^c_7, \dots,  d^c_{12} $ for columns.
+\item Obtain the descriptors $ d^{row}_1, \dots,  d^{row}_6 $  and  $ d^{col}_7, \dots,  d^{col}_{12} $ for rows and columns, respectively from $I^{row}_1, \dots, I^{row}_6$ and $I^{col}_7,\dots,I^{col}_{12}$  in accordance to the method described in Section~\ref{SIFT}.
 
@@
 \begin{equation}
-\hat{r} = \arg \min_{u \in \{1,\dots,6\}} \sum_{q \in NN_T(d^r_u)}^{} \left\lVert q -  d^r_u \right\rVert ^2
+\hat{r} = \arg \min_{u \in \{1,\dots,6\}} \sum_{q \in NN_T(d^{row}_u)}^{} \left\lVert q -  d^{row}_u \right\rVert ^2
 \label{eq:multiclassificationrow}
@@
 \begin{equation}
-\hat{c} = \arg \min_{u \in \{7,\dots,12\}} \sum_{q \in NN_T(d^c_u)}^{} \left\lVert q -  d^c_u \right\rVert ^2
+\hat{c} = \arg \min_{u \in \{7,\dots,12\}} \sum_{q \in NN_T(d^{col}_u)}^{} \left\lVert q -  d^{col}_u \right\rVert ^2
 \label{eq:multiclassificationcol}
@@
 
-where $NN_T(d^l_u), \;l=r,c$ is the set of the $k$ nearest neighbors to $d^l_u, \;l=r,c$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l=r,c$.
+where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u, \;l\in\{row,{col}\}$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l\in\{row,{col}\}$.
 By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
@@
 Figure~\ref{fig:classification} shows a schema of this process. 
-\begin{figure}
+\begin{figure}[htp]
 \centering
@@
 %Para solucionar este problema,
-
-\subsubsection{Signal Plotting} \label{Plot}
-
-Averaged signal segments are scaled and standardized by
-
-\begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor
-\end{equation}
-
-\noindent where $\gamma$ is the image scale, $t$ is time and $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for  channel $c$.
-Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
-
-Consequently, the image is constructed by placing the sample points according to
-
-\begin{equation}
-I(z_1,z_2) = \left\{ \begin{array}{rl}
-255 & \text{if} \,  z_1 = \gamma \cdot t; z_2 = \tilde{x}(t,c) + z(c) \\
-0   & \mbox{otherwise}
-\end{array}\right.
-\label{eq:images}
-\end{equation}
-
-\noindent where $ z_1$ and $z_2$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image.
-The function $z(c)$ is the \textit{zerolevel} which is the location on the image where the signal's zero value should be located in order to fit the entire signal within the image:
-
-\begin{equation}
-z(c) = \left \lfloor{ \frac{\max \tilde{x}(t,c)  - \min \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max \tilde{x}(t,c)  + \min \tilde{x}(t,c)}{ 2} }\right \rfloor
-\label{eq:zerolevel}
-\end{equation}
-
-In order to complete the plot from the pixels, the Bresenham \citep{Bresenham1965,Ramele2016} algorithm is used to interpolate straight lines between each pair of  consecutive pixels.
 
@@
 The experiment is performed with 20 subjects, 17 males, 3 females, average age 26.85 years, standard deviation  8.8 years, range 19-50 years.
+
+EEG data is collected in a single recording session.
+Participants are seated in a comfortable chair, with their vision aligned to a computer screen located one meter in front of them.
+The handling and processing of the data and stimulations is handled by the OpenVibe platform.
 
@@
 This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds in the horizontal direction.
-Finally, descriptors  $\mathbf{kp}$  are located at $ x_{kp} =  0.55 Fs.\gamma = 35 $ and $ y_{kp} = z(c) $ (see Eq.~\ref{eq:zerolevel}).
+Finally, descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ (see Eq.~\ref{eq:zerolevel}).
 In this way the whole transient event is captured.
 Figure~\ref{fig:sampledescriptor2} shows a patch of a signal plot covering the complete amplitude (vertical direction) and the complete span of the signal event (horizontal direction).
-% esta figura no está comentada en el texto.
+
 \begin{figure}[H]
@@
 
-Results are shown in Table \ref{tab:results} where the percentage of correctly spelled letters is calculated while performing an offline BCI Simulation for the public dataset of ALS patients.
+Table~\ref{tab:results} shows the results of applying the algorithm to the subjects of the public dataset.
+The percentage of correctly spelled letters is calculated while performing an offline BCI Simulation for the public dataset of ALS patients.
 From the 7 words for each subject, the first 3 were used as training, and the remaining 4 for testing.
 The best performing channel is informed as well.
-
-DRAFT: Traditionally P300 can be seen in this channel and this other channel.
+The chance level is $2\%$.
+%JG esto después lo hablamos, ¿porque lo pondrías aca?
+%DRAFT: Traditionally P300 can be seen in this channel and this other channel.
 However, Wolpaw showed that the response may also arise in occipital channels as well.
@@
 \begin{table}[H]
-\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject. Chance level is $2\%$. }
+\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject.  }
 \centering
@@
 
-The ITR, or BTR, in the case of reactive BCIs \citep{WolpawJonathanR2012} strongly depends on the amount of signal averaging required to transmit a valid and robust selection.
-The Performance curves (Fig.
-\ref{fig:performance}) show how the percentage of correctly identified letters depends on the number of repetitions that were used to obtain the averaged signal.
-For both datasets, the experimental protocol uses a very short ISI which has the potential to increase the ITR but at the same time is inversily proportional to the amplitude of the p300 response (i.e.
+The ITR, or BTR, in the case of reactive BCIs~\citep{WolpawJonathanR2012}  depends on the amount of signal averaging required to transmit a valid and robust selection.
+Fig.~\ref{fig:performance} shows the performance curves.
+It can be observed that the percentage of correctly identified letters depends on the number of repetitions that were used to obtain the averaged signal.
+For both datasets, the experimental protocol uses a very short ISI which has the potential to increase the ITR but at the same time it is inversely proportional to the amplitude of the p300 response (i.e.
 harder to detect).

### 2017-12-15 article.tex
Cambios y ajustes sobre imagenes. Version base

@@
 
-\firstnote{.}
+%\firstnote{.}
 
@@
 % Keywords
-\keyword{EEG; BCI; P300; ALS; NBNN; HOG; SIFT}
+\keyword{EEG; BCI; P300; ALS; NBNN; HGOSP; SIFT}
 
@@
 
-In~\citep{Ramele2016},  authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves~\citep{Ben-Simon2013}  and Motor Imagery rolandic central $\mu$ rhythms~\citep{Hartman2005} using gradient histograms orientations.
+In~\citep{Ramele2016},  authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves~\citep{Ben-Simon2013}  and Motor Imagery rolandic central $\mu$ rhythms~\citep{Hartman2005} using the histogram of gradient orientations of signal plots.
 Inspired in that work, we proposed a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.
@@
 
-This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of oriented gradients extracted from the image of signal plots, and (2) describe the way in which this classification can be used to implement an offline P300-based BCI Speller application using two public datasets.
+This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of gradient orientations extracted from the image of signal plots, and (2) describe the way in which this classification can be used to implement an offline P300-based BCI Speller application using two public datasets.
 
-This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradients of the Signal Plot method is explained.
+This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradient Orientations of the Signal Plot method is explained.
 Section~\ref{Pipeline} describe the processing pipeline.
@@
 
-\subsection{Feature Extraction based on Histogram of Gradients of the Signal Plot} \label{Feature}
+\subsection{Feature Extraction based on Histogram of Gradient Orientations of the Signal Plot} \label{Feature}
 
-The underlying idea of this method is to generate templates for the signal \citep{Ramele2016}, based on the image plot.
-Hence, the first step is the signal transformation into a temporary binary image.
+The underlying idea of this method is to generate templates for the signal \citep{Ramele2016}, based on the image plot.  Hence, the first step is the processing of the EEG signal.
 
@@
 
-\subsubsection{Feature Extraction: Histogram of Oriented Gradients}
+\subsubsection{Feature Extraction: Histogram of Gradient Orientations}
 \label{SIFT}
@@
 
+%$(\hat{r},\hat{c})$
+
+%$d^{col}_{\hat{c}}$          $d^{row}_{\hat{r}}$
+
 \begin{equation}
@@
 \centering
-\includegraphics[width=15cm]{classificationgraph2.pdf}
-\caption{Single trial segments are averaged for the 6 rows and 6 columns.}
+\includegraphics[width=15cm]{classificationgraph.pdf}
+\caption{Single trial segments are averaged for the 6 rows and 6 columns.
+From the averaged signal, the image of the signal plot is generated and each descriptor is computed.
+By comparing each descriptor against the set of templates, the P300 ERP can be detected, and finally the desired letter from the matrix can be inferred.}
 \label{fig:classification}
@@
 Participants are seated in a comfortable chair, with their vision aligned to a computer screen located one meter in front of them.
-The handling and processing of the data and stimulations is handled by the OpenVibe platform.
+The handling and processing of the data and stimulations is handled by the OpenVibe platform~\citep{Renard2010}.
 
@@
 \includegraphics[width=10cm]{patchgeometry.pdf}
-\caption{The scale of local patch is selected in order to capture the whole signal information.
-The vertical size $4 \cdot 3 s_y$ of the patch is required to be long enough as to contain the signal $\Delta  \mu V $, which is the peak-to-peak amplitude of the transient event.
-The horizontal size $ 4 \cdot 3 s_x$ , on the other hand, needs to cover the entire duration in seconds of the transient signal event, $ \lambda $.
+\caption{The scale of local patch is selected in order to capture the whole transient event.
+The size of the patch is $S_p \times S_p$ pixels.
+The vertical size consists of $4$ blocks $B_{i,j}$  of size $3 s_y$ pixels which is long enough as to contain the signal $\Delta  \mu V $, the peak-to-peak amplitude of the transient event.
+The horizontal size on the other hand, includes $4$ blocks  of $3 s_x$ and cover the entire duration in seconds of the transient signal event, $ \lambda $.
 }
+
 \label{fig:sampledescriptor2}
@@
 
-Table~\ref{tab:results} shows the results of applying the algorithm to the subjects of the public dataset.
-The percentage of correctly spelled letters is calculated while performing an offline BCI Simulation for the public dataset of ALS patients.
+Table~\ref{tab:results} shows the results of applying the algorithm to the subjects of the public dataset of ALS patients.
+The percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
 From the 7 words for each subject, the first 3 were used as training, and the remaining 4 for testing.
@@
 %JG esto después lo hablamos, ¿porque lo pondrías aca?
+%RR: esto corresponde al punto 4.1 que dijo el revisor y tiene que ver con proveer informacion del canal en el que la clasificacion dio mejor.
 %DRAFT: Traditionally P300 can be seen in this channel and this other channel.
@@
 \includegraphics[width=10cm]{performance.eps}
-\caption{Performance curves for the eight subjects included in the published dataset.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
+\caption{Performance curves for the eight subjects included in the ALS dataset.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
 \label{fig:performance}
@@
 \midrule
-1     &     Cz   &     0.7000  &  0.3500  \\
-1     &     Cz   &    0.1000  &  0.0500 \\
-1     &     Cz   &     0.1000  &  0.1000 \\
-1     &     Cz   &     0.1000  &  0.1500 \\
-1     &     Cz   &     0.1000  &  0.1000 \\
-1     &     Cz   &     0.5500  &  0.4000 \\
-1     &     Cz   &     0.1500   & 0.1000 \\
-1     &     Cz   &     0.4500   & 0.2000 \\
-1     &     Cz   &     0.3000  &  0.2000 \\
-1     &     Cz   &     0.2500  &  0.1500 \\
-1     &     Cz   &     0.5500  &  0.3000 \\
-1     &     Cz   &     0.3500  &  0.1000 \\
-1     &     Cz   &     0.3000  &  0.3500 \\
-1     &     Cz   &     0.5000  &  0.4000 \\
-1     &     Cz   &     0.3000  &  0.1500 \\
-1     &     Cz   &     0.3000  &  0.3000 \\
-1     &     Cz   &     0.3500  &  0.3500 \\
-1     &     Cz   &          0      & 0.0500 \\
-1     &     Cz   &     0.7000   & 0.5000 \\
-1     &     Cz   &     0.4000  &  0.3000 \\
+1     &     Cz   &     0.7000  &  $45\%$  \\
+1     &     Cz   &    0.1000   &  $10\%$ \\
+1     &     Cz   &     0.1000  &  $15\%$ \\
+1     &     Cz   &     0.1000  &  $20\%$  \\
+1     &     Cz   &     0.1000  &  $10\%$ \\
+1     &     Cz   &     0.5500  &  $25\%$ \\
+1     &     Cz   &     0.1500  &  $15\%$ \\
+1     &     Cz   &     0.4500  &  $20\%$ \\
+1     &     Cz   &     0.3000  &  $20\%$ \\
+1     &     Cz   &     0.2500  &  $20\%$ \\
+1     &     Cz   &     0.5500  &  $15\%$ \\
+1     &     Cz   &     0.3500  &  $30\%$ \\
+1     &     Cz   &     0.3000  &  $10\%$ \\
+1     &     Cz   &     0.5000  &  $35\%$ \\
+1     &     Cz   &     0.3000  &  $45\%$ \\
+1     &     Cz   &     0.3000  &  $10\%$ \\
+1     &     Cz   &     0.3500  &  $30\%$ \\
+1     &     Cz   &          0      &  $45\%$ \\
+1     &     Cz   &          0      &  $15\%$ \\
+1     &     Cz   &     0.7000  &  $65\%$ \\
+1     &     Cz   &     0.4000  &  $40\%$ \\
 

### 2017-12-15 article.tex
Subo los cambios de Juliana

@@
 % Abstract (Do not use inserted blank lines, i.e. \\) 
-\abstract{EL OBJETIVO DE ESTE PAPER, EL HILO CONDUCTOR, ES DEMOSTRAR QUE ESTE METODO QUE ESTA BASADO EN LA ESTRUCTURA DE LA SENIAL SE PUEDE UTILIZAR TAMBIEN PARA IDENTIFICAR EVENTOS TRANSIENTES COMO EL P300, Y QUE CON ESO SE PUEDE IMPLEMENTAR UN SPELLER DE P300.
+\abstract{%EL OBJETIVO DE ESTE PAPER, EL HILO CONDUCTOR, ES DEMOSTRAR QUE ESTE METODO QUE ESTA BASADO EN LA ESTRUCTURA DE LA SENIAL SE PUEDE UTILIZAR TAMBIEN PARA IDENTIFICAR EVENTOS TRANSIENTES COMO EL P300, Y QUE CON ESO SE PUEDE IMPLEMENTAR UN SPELLER DE P300.
 The analysis of Electroencephalographic (EEG) signals is of ulterior importance to elucidate patterns that could improve the implementation of Brain Computer Interfaces (BCI).
@@
 
-In~\citep{Ramele2016},  authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves~\citep{Ben-Simon2013}  and Motor Imagery rolandic central $\mu$ rhythms~\citep{Hartman2005} using the histogram of gradient orientations of signal plots.
+In~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} an approach based on the morphological analysis of the shape of the EEG is presented. In this work the authors use slop chain code.
+
+In the article~\citep{Ramele2016},  the authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves  and Motor Imagery rolandic central $\mu$ rhythms using the histogram of gradient orientations of signal plots.
 Inspired in that work, we proposed a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.
-Its validity is verified by processing offline data for ALS patients and for healthy subjects.
-The proposed approach is based on the morphological analysis of the shape of the EEG signal~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} and it was built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting raw signal plots~\citep{Hartman2005}.
+The proposed approach is based on the morphological analysis of the shape of the EEG signal, but using histogram descriptors.
+The method is built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting raw signal plots, as it is described in~\citep{Hartman2005}.
 
 This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of gradient orientations extracted from the image of signal plots, and (2) describe the way in which this classification can be used to implement an offline P300-based BCI Speller application using two public datasets.
+Its validity is verified by processing offline data for ALS patients and for healthy subjects.
 
@@
 Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008} and the final Section shows results and discussion where we expose our remarks, conclusions and future work.
-
+%This kind of EEG events are described in~\citep{Hartman2005} and~\citep{Ben-Simon2013}.
 \section{Materials and Methods}
@@
 Fig.~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
-Each time a row or column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
+Each time a row or a column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
 
@@
 
-The underlying idea of this method is to generate templates for the signal \citep{Ramele2016}, based on the image plot.  Hence, the first step is the processing of the EEG signal.
+%The underlying idea of this method is to generate templates for the signal \citep{Ramele2016}, based on the image plot.  Hence, the first step is the processing of the EEG signal.
+In this section, the signal preprocessing, the method for generating images from signal plots, the feature extraction procedure and the speller matrix identification are described.
 
-\subsubsection{Processing Pipeline} \label{Pipeline}
+\subsubsection{Preprocessing Pipeline} \label{Pipeline}
 
 \begin{itemize}
-\item \textbf{Preprocessing}: The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG.
+\item \textbf{Signal Enhancement}: The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG.
 The processing pipeline starts by applying a notch filter to the raw signal, a
@@
 \begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, t\geq 0,c \in \{1,2,\dots,8\}
+\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, t \in \{0, 1, \dots, F_s\},c \in \{1,2,\dots,8\}
 \end{equation}
@@
 On the other hand, $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for channel $c$.
+The parameter $F_s$ is the sampling frequency.
 Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
@@
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
-\item Compute  
+\item \label{paso3} Compute  
 \begin{equation}
@@
 
-\item Plot the signal $x(t,c)$, according Section~\ref{Plot}. The images $I^{row}_1, \dots, I^{row}_6$ and $I^{col}_7,\dots,I^{col}_{12}$ are obtained for rows and columns, respectively.
+\item \label{paso4}Plot the signal $x(t,c)$, according Section~\ref{Plot}.
+
+\item Repeat steps~\ref{paso3} and~\ref{paso4} in order to generate the images $I^{row}_1, \dots, I^{row}_6$ and $I^{col}_7,\dots,I^{col}_{12}$ for rows and columns, respectively.
 
@@
 \begin{equation}
-\hat{r} = \arg \min_{u \in \{1,\dots,6\}} \sum_{q \in NN_T(d^{row}_u)}^{} \left\lVert q -  d^{row}_u \right\rVert ^2
+\hat{row} = \arg \min_{u \in \{1,\dots,6\}} \sum_{q \in NN_T(d^{row}_u)}^{} \left\lVert q -  d^{row}_u \right\rVert ^2
 \label{eq:multiclassificationrow}
@@
 \begin{equation}
-\hat{c} = \arg \min_{u \in \{7,\dots,12\}} \sum_{q \in NN_T(d^{col}_u)}^{} \left\lVert q -  d^{col}_u \right\rVert ^2
+\hat{col} = \arg \min_{u \in \{7,\dots,12\}} \sum_{q \in NN_T(d^{col}_u)}^{} \left\lVert q -  d^{col}_u \right\rVert ^2
 \label{eq:multiclassificationcol}
@@
 where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u, \;l\in\{row,{col}\}$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l\in\{row,{col}\}$.
-By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{r} $ and column $ \hat{c} $.
 
@@
 \end{enumerate}
-
+By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{row} $ and column $ \hat{col} $. 
 Figure~\ref{fig:classification} shows a schema of this process. 
@@
 \includegraphics[width=15cm]{classificationgraph.pdf}
-\caption{Single trial segments are averaged for the 6 rows and 6 columns.
+\caption{Single trial segments $S_i$ are averaged for the 6 rows and 6 columns.
 From the averaged signal, the image of the signal plot is generated and each descriptor is computed.

### 2017-12-18 article.tex
Mas cambios y ajustes.  Reforce con mas info la parte de resultados

@@
 
-In~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} an approach based on the morphological analysis of the shape of the EEG is presented. In this work the authors use slop chain code.
-
 In the article~\citep{Ramele2016},  the authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves  and Motor Imagery rolandic central $\mu$ rhythms using the histogram of gradient orientations of signal plots.
@@
 The method is built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting raw signal plots, as it is described in~\citep{Hartman2005}.
+
+Other works have explored the idea of describing the waveform of a signal.
+In~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} an approach based on Slope Chain Code is presented, and also the work of Bondt-Pompe Permutation Entropy~\citep{Berger2017}.
 
@@
 For each illuminated row or column, an image is obtained.
-\item Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and chanel, respectively.
+\item  \label{paso2} Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and channel, respectively.
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
@@
 
-\item Repeat steps~\ref{paso3} and~\ref{paso4} in order to generate the images $I^{row}_1, \dots, I^{row}_6$ and $I^{col}_7,\dots,I^{col}_{12}$ for rows and columns, respectively.
+\item Repeat steps~\ref{paso2}, \ref{paso3} and~\ref{paso4} in order to generate the images $I^{row}_1, \dots, I^{row}_6$ and $I^{col}_7,\dots,I^{col}_{12}$ for rows and columns, respectively.
 
@@
 The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.
-This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ in the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds in the horizontal direction.
+This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ on the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds on the horizontal direction.
 Finally, descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ (see Eq.~\ref{eq:zerolevel}).
@@
 The chance level is $2\%$.
-%JG esto después lo hablamos, ¿porque lo pondrías aca?
+
 %RR: esto corresponde al punto 4.1 que dijo el revisor y tiene que ver con proveer informacion del canal en el que la clasificacion dio mejor.
-%DRAFT: Traditionally P300 can be seen in this channel and this other channel.
-However, Wolpaw showed that the response may also arise in occipital channels as well.
-The authors of the original paper concluded that Cz was the best performing channel, more representative of the difference in amplitudes.
-However, in our approach occipital channels,  PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
+
+The P300 wave which was traditionally considered as one component, consists of two overlapping components the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\citep{Polich2007}.
+Although, the traditional approach is to expect to find a stronger response on the central channel Cz~\citep{Riccio2013} Wolpaw et al showed that the response may also arise in occipital channels as well.
+The authors~\citep{Riccio2013} that analyzed the public dataset concluded that Cz was the best performing channel, more representative of the difference in amplitudes.
+However, in our approach, occipital channels,  PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
 
@@
 
+\begin{itemize}
+\item PODRIA ACA MOSTRAR ADEMAS PLOTS DE SINGLE TRIAL PARA QUE VEAN QUE SON UN QUILOMBO.
+\item UNA DE LAS RAZONES POR LA CUAL PODEMOS LLEGAR A ARGUMENTAR QUE NO FUIMOS MAS A FONDO CON EL TEMA ES QUE NO QUERIAMOS DISTORSIONAR DEMASIADO LA SENAL COMPLEJIZANDO EL ALGORITMO Y QUERIAMOS MANTENERNOS ENTRE LO MAS ESTANDAR POSIBLE (PUEDO CITAR MAS O MENOS QUE NUESTRO PROCESSING PIPELINE ES ESTANDARD).
+\item ACA EL PUNTO PARA MOSTRAR ES QUE DE ALGUNA MANERA ESTAMOS COMO MOSTRANDO UN "SI SE PUEDE" DE UN TRABAJO QUE PUEDE CONTINUAR Y EXPANDIRSE.
+POR ESO LA IDEA DE MOSTRAR PUNTAS DE COSAS QUE VIMOS E HICIMOS, PERO SIN MOSTRAR RESULTADOS CONCRETOS.
+\item La pregunta:  ¿ Es posible utilizando solamente la información de los plots, aún utilizando preprocesamiento y una mejora en el SNR basado en el esquema tradicional de generación del patrón de p300, detectarlo con alto grado de éxito ?
+
+\begin{itemize}
+\item Es posible pero sólo en algunos pacientes.
+\item Se requiere de mecanismos para estabilizar el waveform de p300.
+
+\end{itemize}
+
+\item 
+\end{itemize}
+
 We found that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that could allow the classification to reach a higher level.
-Moreover, as can be seen in the figure, as repetitions tend to 1, which will correspond to single trial letter identification, the performance suffers greatly as the signal plot is simple too complex to be used as template and the method is completely unable to find good matches  PODRIA ACA MOSTRAR ADEMAS PLOTS DE SINGLE TRIAL PARA QUE VEAN QUE SON UN QUILOMBO.
+Moreover, as can be seen in the figure, as repetitions tend to 1, which will correspond to single trial letter identification, the performance suffers greatly as the signal plot is simple too complex to be used as template and the method is completely unable to find good matches
 
-UNA DE LAS RAZONES POR LA CUAL PODEMOS LLEGAR A ARGUMENTAR QUE NO FUIMOS MAS A FONDO CON EL TEMA ES QUE NO QUERIAMOS DISTORSIONAR DEMASIADO LA SENAL COMPLEJIZANDO EL ALGORITMO Y QUERIAMOS MANTENERNOS ENTRE LO MAS ESTANDAR POSIBLE (PUEDO CITAR MAS O MENOS QUE NUESTRO PROCESSING PIPELINE ES ESTANDARD).
-
-We hypothesize that as subjects may have different latencies and amplitudes of the P300 signal \citep{Riccio2013}, it may also be the case that the shape of the generated ERP may vary greatly in an intra-subject manner, thus the pattern could not be well generalized and a higher performance could not be reached (Fig.
+We hypothesize that as subjects may have different latencies, amplitudes and width of the P300 signal \citep{Riccio2013}, it may also be the case that the shape of the generated ERP may vary greatly in an intra-subject manner, thus the pattern could not be well generalized and a higher performance could not be reached (Fig.
 \ref{fig:p300templates}).
@@
 We also applied DTW but couldn't found a sustantial improvement.
-
-Additionally we believe that our method suffers also from null signals.
-
-ACA EL PUNTO PARA MOSTRAR ES QUE DE ALGUNA MANERA ESTAMOS COMO MOSTRANDO UN "SI SE PUEDE" DE UN TRABAJO QUE PUEDE CONTINUAR Y EXPANDIRSE.
-POR ESO LA IDEA DE MOSTRAR PUNTAS DE COSAS QUE VIMOS E HICIMOS, PERO SIN MOSTRAR RESULTADOS CONCRETOS.
+Amplitude width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals
 
@@
 \midrule
-1     &     Cz   &     0.7000  &  $45\%$  \\
-1     &     Cz   &    0.1000   &  $10\%$ \\
-1     &     Cz   &     0.1000  &  $15\%$ \\
-1     &     Cz   &     0.1000  &  $20\%$  \\
-1     &     Cz   &     0.1000  &  $10\%$ \\
-1     &     Cz   &     0.5500  &  $25\%$ \\
-1     &     Cz   &     0.1500  &  $15\%$ \\
-1     &     Cz   &     0.4500  &  $20\%$ \\
-1     &     Cz   &     0.3000  &  $20\%$ \\
-1     &     Cz   &     0.2500  &  $20\%$ \\
-1     &     Cz   &     0.5500  &  $15\%$ \\
-1     &     Cz   &     0.3500  &  $30\%$ \\
-1     &     Cz   &     0.3000  &  $10\%$ \\
-1     &     Cz   &     0.5000  &  $35\%$ \\
-1     &     Cz   &     0.3000  &  $45\%$ \\
-1     &     Cz   &     0.3000  &  $10\%$ \\
-1     &     Cz   &     0.3500  &  $30\%$ \\
-1     &     Cz   &          0      &  $45\%$ \\
-1     &     Cz   &          0      &  $15\%$ \\
-1     &     Cz   &     0.7000  &  $65\%$ \\
-1     &     Cz   &     0.4000  &  $40\%$ \\
+1     &     Fz   &     $40\%$  \\
+2     &     Oz   &     $30\%$ \\
+3     &     Cz   &     $40\%$ \\
+4     &     Cz   &     $45\%$ \\
+5     &     P4   &      $60\%$ \\
+6     &     P3   &      $50\%$ \\
+7     &     PO7   &      $70\%$ \\
+8     &     PO8   &      $50\%$ \\
 
@@
 
-In Table \ref{tab:resultsowndataset} results obtained for the 20 healthy subjects is shown.
-The third column shows the results obtained by using the Krusienski feature and the SVM classifier.
-On the other hand, the fourth column shows the results obtained by the present method as described in previous sections.
-Subject 17 was used as control as he was instructed to count only the letters and to stare at the screen but not neglected row/column flashing.
+In Table \ref{tab:resultsowndataset} results obtained for the 8 healthy subjects is shown.
+It shows the results obtained by the present method as described in previous sections.
+We used an additional subject as control as he was instructed to count only the letters and to stare at the screen but  neglecting row/column flashing, which achieved zero performance.
+We also validated the procedure by visualizing clearly the SSVEP generated by the 4 Hz flashing of lights.
+
+The method benefical aspect is that its single channel approach and also the most important part is that the method allows close collaboration because templates can be selected by visually inspecting them.
 

### 2017-12-18 article.tex
Version actualizada junto con lo de Juliana

@@
 
-Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitalized devices has pressed for a revamping of this hundred years old technology.
-Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information \cite{DeVos2014}
+Although recent advances in neuroimagining techniques (particularly radio-nuclear and radiological scanning methods) \citep{Schomer2010} have diminished the prospects of the traditional Electroencephalography (EEG), the advent and development of digitized devices has pressed for a revamping of this hundred years old technology.
+Their versatility, ease of use, temporal resolution, ease of development and fabrication, and its proliferation as consumer devices, are pushing EEG to become the de-facto non invasive portable or ambulatory method to access and harness brain information~\cite{DeVos2014}.
 
-A key contribution to this expansion has been the field of Brain Computer Interfaces (BCI) \citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neurodegenerative diseases.
+A key contribution to this expansion has been the field of Brain Computer Interfaces (BCI)~\citep{WolpawJonathanR2012} which is the pursuit of the development of a new channel of communication particularly aimed to persons affected by neurodegenerative diseases.
 
-One noteworthy aspect of this novel communication channel is the ability to volitionally transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair \citep{Carlson2013}, as input to a speller application \citep{Guger2009a}, in a Virtual Reality environment \citep{Lotte2013} or as aiding tool in a rehabilitation procedure \citep{Jure2016}.
-The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion \citep{WolpawJonathanR2012}.
+One noteworthy aspect of this novel communication channel is the ability to transmit information from the Central Nervous System (CNS) to a computer device and from there use that information to control a wheelchair~\citep{Carlson2013}, as input to a speller application~\citep{Guger2009a}, in a Virtual Reality environment~\citep{Lotte2013} or as aiding tool in a rehabilitation procedure~\citep{Jure2016}.
+The holly grail of BCI is to implement a new complete and alternative pathway to restore lost locomotion~\citep{WolpawJonathanR2012}.
 
@@
 
-Moreover, EEG markers \citep{Clerc} that can be used to volitionally transmit information are limited, and each one of them has a particular combination of appropriate methods to decode them.
+Moreover, EEG markers~\citep{Clerc} that can be used to volitionally transmit information are limited, and each one of them has a particular combination of appropriate methods to decode them.
 Inevitably, it is necessary to implement  distinct and specialized algorithmic methods, to filter the signal, enhance its Signal to Noise Ratio (SNR), and try to determine some meaning out of it.
@@
 
-BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon  \citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
-New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments \citep{Huggins2016}.
+BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon~\citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
+New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments~\citep{Huggins2016}.
 However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement~\citep{Clerc}.
+
+%JG Muy desordenada la intro, escribir un poco mejor
+Several works have explored the idea of exploiting the signal waveform to analyze the EEG signal.
+In~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} an approach based on Slope Chain Code is presented, as well as the seminal work of Bondt-Pompe Permutation Entropy~\citep{Berger2017}.%JG Y también qué?
+
+%In~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} an approach based on the morphological analysis of the shape of the EEG is presented. In this work the authors use slop chain code.
 
@@
 The method is built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting raw signal plots, as it is described in~\citep{Hartman2005}.
-
-Other works have explored the idea of describing the waveform of a signal.
-In~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} an approach based on Slope Chain Code is presented, and also the work of Bondt-Pompe Permutation Entropy~\citep{Berger2017}.
 
@@
 This signal can be used to implement a speller application by means of a Speller Matrix~\citep{Farwell1988}.
-Fig.~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashings of rows and columns provide the deviant stimulus required to elicit this physiological response.
+Fig.~\ref{fig:p300matrix} shows an example of the Speller Matrix used in the OpenVibe Open Source software~\citep{Renard2010}, where the flashes of rows and columns provide the deviant stimulus required to elicit this physiological response.
 Each time a row or a column that contains the desired letter flashes, the corresponding synchronized EEG signal should also contain the P300 signature and by detecting it, the selected letter can be identified.
@@
 
-\noindent where $ (z_1,z_2) \in \mathcal{N} \times \mathcal{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image.
-The function $z(c)$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
+\noindent where $ (z_1,z_2) \in \mathcal{N} \times \mathcal{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $t \in \{0, 1, \dots, F_s\}$ and $c \in \{1,2,\dots,8\}$.
+The function $z(c)$, $c \in \{1,2,\dots,8\}$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
 
@@
 \begin{enumerate}
-\item Highlight randomly the rows and columns from the matrix.
+\item \label{paso1}Highlight randomly the rows and columns from the matrix.
 There is one row and one column that should match the letter selected by the subject.
 For each illuminated row or column, an image is obtained.
-\item  \label{paso2} Repeat item 1 $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t$ and $c$ correspond to time and channel, respectively.
+\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t \in \{0, 1, \dots, F_s\}$ and $c \in \{1,2,\dots,8\}$ correspond to time and channel, respectively.
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
@@
 \begin{equation}
-x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c), t\geq 0, c \in \{1,\dots,8\}
+x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c), t \in \{0, 1, \dots, F_s\}, c \in \{1,\dots,8\}
 \label{averaging}
@@
 
-\item \label{paso4}Plot the signal $x(t,c)$, according Section~\ref{Plot}.
+\item \label{paso4}Plot the signal $x(t,c)$,  $t \in \{0, 1, \dots, F_s\}$, $c \in \{1,\dots,8\}$,  according Section~\ref{Plot}.
 
@@
 \item Match to the Template $T$ by computing
-
-%$(\hat{r},\hat{c})$
-
-%$d^{col}_{\hat{c}}$          $d^{row}_{\hat{r}}$
 
@@
 The percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
-From the 7 words for each subject, the first 3 were used as training, and the remaining 4 for testing.
+From the seven words for each subject, the first three were used as training, and the remaining four for testing.
 The best performing channel is informed as well.
 The chance level is $2\%$.
-
-%RR: esto corresponde al punto 4.1 que dijo el revisor y tiene que ver con proveer informacion del canal en el que la clasificacion dio mejor.
-
-The P300 wave which was traditionally considered as one component, consists of two overlapping components the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\citep{Polich2007}.
-Although, the traditional approach is to expect to find a stronger response on the central channel Cz~\citep{Riccio2013} Wolpaw et al showed that the response may also arise in occipital channels as well.
-The authors~\citep{Riccio2013} that analyzed the public dataset concluded that Cz was the best performing channel, more representative of the difference in amplitudes.
-However, in our approach, occipital channels,  PO8 and PO7 showed higher performances \citep{Huggins2016,Jure2016}.
 
@@
 
-The ITR, or BTR, in the case of reactive BCIs~\citep{WolpawJonathanR2012}  depends on the amount of signal averaging required to transmit a valid and robust selection.
+%RR: esto corresponde al punto 4.1 que dijo el revisor y tiene que ver con proveer informacion del canal en el que la clasificacion dio mejor.
+
+The P300 wave which was traditionally considered as one component, consists of two overlapping components the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\citep{Polich2007}.
+Although the traditional approach is to expect to find a stronger response on the central channel Cz~\citep{Riccio2013}, Wolpaw et al.
+shows that the response may also arise in occipital channels as well.
+The authors~\citep{Riccio2013} that analyzed the public dataset concluded that Cz was the best performing channel, more representative of the difference in amplitudes.
+However, in our approach, occipital channels,  PO8 and PO7 showed higher performances~\citep{Huggins2016,Jure2016}.
+
+%JG ITR y BTR nunca fueron definidos
+The Information Transfer Rate (ITR), or Bit Transfer Rate (BTR), in the case of reactive BCIs~\citep{WolpawJonathanR2012}  depends on the amount of signal averaging required to transmit a valid and robust selection.
 Fig.~\ref{fig:performance} shows the performance curves.
-It can be observed that the percentage of correctly identified letters depends on the number of repetitions that were used to obtain the averaged signal.
-For both datasets, the experimental protocol uses a very short ISI which has the potential to increase the ITR but at the same time it is inversely proportional to the amplitude of the p300 response (i.e.
+It can be observed that the percentage of correctly identified letters depends on the number of repetitions that are used to obtain the averaged signal.
+For both datasets, the experimental protocol uses a very short ISI which has the potential to increase the ITR but at the same time it is inversely proportional to the amplitude of the P300 response (i.e.
 harder to detect).
-As an unexpected consequence, 15 out of 20 subject reported high ocular tiredness by using this configuration.
+As an unexpected consequence, $15$ out of $20$ subject reported high ocular tiredness by using this configuration.
 
@@
 
-\begin{itemize}
-\item PODRIA ACA MOSTRAR ADEMAS PLOTS DE SINGLE TRIAL PARA QUE VEAN QUE SON UN QUILOMBO.
-\item UNA DE LAS RAZONES POR LA CUAL PODEMOS LLEGAR A ARGUMENTAR QUE NO FUIMOS MAS A FONDO CON EL TEMA ES QUE NO QUERIAMOS DISTORSIONAR DEMASIADO LA SENAL COMPLEJIZANDO EL ALGORITMO Y QUERIAMOS MANTENERNOS ENTRE LO MAS ESTANDAR POSIBLE (PUEDO CITAR MAS O MENOS QUE NUESTRO PROCESSING PIPELINE ES ESTANDARD).
-\item ACA EL PUNTO PARA MOSTRAR ES QUE DE ALGUNA MANERA ESTAMOS COMO MOSTRANDO UN "SI SE PUEDE" DE UN TRABAJO QUE PUEDE CONTINUAR Y EXPANDIRSE.
+%\begin{itemize}
+%\item PODRIA ACA MOSTRAR ADEMAS PLOTS DE SINGLE TRIAL PARA QUE VEAN QUE SON UN QUILOMBO.%JG no sé, no lo pidió
+%\item UNA DE LAS RAZONES POR LA CUAL PODEMOS LLEGAR A ARGUMENTAR QUE NO FUIMOS MAS A FONDO CON EL TEMA ES QUE NO QUERIAMOS DISTORSIONAR DEMASIADO LA SENAL COMPLEJIZANDO EL ALGORITMO Y QUERIAMOS MANTENERNOS ENTRE LO MAS ESTANDAR POSIBLE (PUEDO CITAR MAS O MENOS QUE NUESTRO PROCESSING PIPELINE ES ESTANDARD).%%JG No, a menos que lo pregunte.
+Tenemos que transmitir que fuimos a fondo en el tema
+%\item ACA EL PUNTO PARA MOSTRAR ES QUE DE ALGUNA MANERA ESTAMOS COMO MOSTRANDO UN "SI SE PUEDE" DE UN TRABAJO QUE PUEDE CONTINUAR Y EXPANDIRSE.
 POR ESO LA IDEA DE MOSTRAR PUNTAS DE COSAS QUE VIMOS E HICIMOS, PERO SIN MOSTRAR RESULTADOS CONCRETOS.
-\item La pregunta:  ¿ Es posible utilizando solamente la información de los plots, aún utilizando preprocesamiento y una mejora en el SNR basado en el esquema tradicional de generación del patrón de p300, detectarlo con alto grado de éxito ?
+%\item La pregunta:  ¿ Es posible utilizando solamente la información de los plots, aún utilizando preprocesamiento y una mejora en el SNR basado en el esquema tradicional de generación del patrón de p300, detectarlo con alto grado de éxito ?
+%%JG esto puede ir a las conclusiones.
 
-\begin{itemize}
-\item Es posible pero sólo en algunos pacientes.
-\item Se requiere de mecanismos para estabilizar el waveform de p300.
+%\begin{itemize}
+%\item Es posible pero sólo en algunos pacientes.
+%\item Se requiere de mecanismos para estabilizar el waveform de p300.
 
-\end{itemize}
+%\end{itemize}
 
-\item 
-\end{itemize}
-
-We found that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that could allow the classification to reach a higher level.
-Moreover, as can be seen in the figure, as repetitions tend to 1, which will correspond to single trial letter identification, the performance suffers greatly as the signal plot is simple too complex to be used as template and the method is completely unable to find good matches
+%\item 
+%\end{itemize}
+%%JG No entiendo nada
+We find that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that could allow the classification to reach a higher level.
+Moreover, as can be seen in the figure, as repetitions tend to 1, which will correspond to single trial letter identification, the performance suffers greatly as the signal plot is  too complex to be used as template and the method is completely unable to find good matches
 
@@
 It is known (REFERENCE) that the point-to-point averaging suffers from inter-trial variability (REFERENCIA A ESE PAPER Con ese nombre) due to latencies shifts.
-We tested a naive approach to verify if the morphological shape of the p300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied.
-We also applied DTW but couldn't found a sustantial improvement.
-Amplitude width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals
+We tested a naive approach to verify if the morphological shape of the P300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied.
+We also applied DTW but couldn't found a substantial improvement.
+Amplitude width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals.
 
-Considering the amplitude variation of the p300 and its components, we proposed a conservative approach by standarizing the signal (which can have both effects, of reducing noise that were not reduced by the averaging procedure but at the same time it can reduce the amplitude of a significant p300 compoent).
-The standarization also provides another benefit which is sometimes neglected as unimportant.
+Considering the amplitude variation of the P300 and its components, we proposed a conservative approach by standardizing the signal (which can have both effects, of reducing noise that were not reduced by the averaging procedure but at the same time it can reduce the amplitude of a significant p300 component).
+The standardization also provides another benefit which is sometimes neglected as unimportant.
 The variance of the averaged signal depends on the number of signals averaged.
-As this method analyze precisely the shape of the signal in the plot, which is very affected by the variance, the method will not be able to discriminate solely based on the underlying cognitive event, it will be confused by the number of signals that were used to calculate the averaged signal.
+As this method analyze the shape of the signal in the plot, which is  affected by the variance, the method is not be able to discriminate solely based on the underlying cognitive event, it is confused by the number of signals that are used to calculate the averaged signal.
 
@@
 
+Figure~\ref{fig:p300templates} shows the P300 templates patches for subjects 8 and 3 from the public dataset.
+It can be observed that in coincidence with the performance results, the P300 signature is more clear and consistent for subject 8 (Fig.~\ref{subject8})  while for subject 3 (Fig.~\ref{subject3}) the characteristic pattern is more difficult to perceive.
+
 \begin{figure}[H]
 \centering
-\includegraphics[width=10cm]{subject8.png}
-\includegraphics[width=10cm]{subject3.png}
-\caption{Ten P300 template patches found for subjects 8 (up) and 3(down).
-In coincidence with the performance results, the P300 signature is more clear and consistent for subject 8 (higher performance) while for subject 3 (lower performance) the characteristic pattern is much more difficult to perceive.}
+\subfigure[Subject 8.]{\includegraphics[width=10cm]{subject8.png}\label{subject8}}
+\subfigure[Subject 3.]{\includegraphics[width=10cm]{subject3.png}\label{subject3}}
+\caption{Ten P300 template patches for subjects 8 and 3. }
 \label{fig:p300templates}
@@
 
+In Table~\ref{tab:resultsowndataset} results obtained for the 8 healthy subjects is shown.
+It shows the results obtained by the present method as described in previous sections.
+We used an additional subject as control as he was instructed to count only the letters and to stare at the screen but  neglecting row/column flashing, which achieved zero performance.
+We also validated the procedure by visualizing clearly the SSVEP generated by the 4 Hz flashing of lights.
 \begin{table}[H]
@@
 
-In Table \ref{tab:resultsowndataset} results obtained for the 8 healthy subjects is shown.
-It shows the results obtained by the present method as described in previous sections.
-We used an additional subject as control as he was instructed to count only the letters and to stare at the screen but  neglecting row/column flashing, which achieved zero performance.
-We also validated the procedure by visualizing clearly the SSVEP generated by the 4 Hz flashing of lights.
-
-The method benefical aspect is that its single channel approach and also the most important part is that the method allows close collaboration because templates can be selected by visually inspecting them.
+The method beneficial aspect is that its single channel approach and also the most important part is that the method allows close collaboration because templates can be selected by visually inspecting them.
 
@@
 
-The adaptive behaviour of the algorithm make it well suited when the shape of the pattern elicited by the P300 response does not conform to the predicted structure.
+The adaptive behavior of the algorithm make it well suited when the shape of the pattern elicited by the P300 response does not conform to the predicted structure.
 This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e.

### 2017-12-18 article.tex
Muchos cambios

@@
 
-Moreover, EEG markers~\citep{Clerc} that can be used to volitionally transmit information are limited, and each one of them has a particular combination of appropriate methods to decode them.
+Moreover, EEG markers~\citep{Clerc} that can be used to  transmit volitional information are limited, and each one of them has a particular combination of appropriate methods to decode them.
 Inevitably, it is necessary to implement  distinct and specialized algorithmic methods, to filter the signal, enhance its Signal to Noise Ratio (SNR), and try to determine some meaning out of it.
@@
 %Patients suffering from ALS their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
-
-%Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
-It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
 
@@
 
-%JG Muy desordenada la intro, escribir un poco mejor
 Several works have explored the idea of exploiting the signal waveform to analyze the EEG signal.
-In~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} an approach based on Slope Chain Code is presented, as well as the seminal work of Bondt-Pompe Permutation Entropy~\citep{Berger2017}.%JG Y también qué?
-
-%In~\citep{Alvarado-Gonzalez2016,Yamaguchi2009} an approach based on the morphological analysis of the shape of the EEG is presented. In this work the authors use slop chain code.
-
+In~\citep{Alvarado-Gonzalez2016} an approach based on Slope Horizontal Chain Code is presented, whereas in ~\citep{Yamaguchi2009} a similar procedure was implemented based on Mathematical Morphological Analysis.
+The seminal work of Bandt-Pompe Permutation Entropy~\citep{Berger2017} also explores succinctly this idea as a basis to establish the time series ordinal patterns.
 In the article~\citep{Ramele2016},  the authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves  and Motor Imagery rolandic central $\mu$ rhythms using the histogram of gradient orientations of signal plots.
 Inspired in that work, we proposed a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.
-The proposed approach is based on the morphological analysis of the shape of the EEG signal, but using histogram descriptors.
+The proposed approach is based on the waveform analysis of the shape of the EEG signal, but using histogram of gradient orientations.
 The method is built by mimicking what traditionally electroencephalographers have been performing for almost a century: visually inspecting raw signal plots, as it is described in~\citep{Hartman2005}.
@@
 This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradient Orientations of the Signal Plot method is explained.
-Section~\ref{Pipeline} describe the processing pipeline.
+Section~\ref{Pipeline} describes the processing pipeline.
 Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.
-Section \ref{Classification}  presents the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008} and the final Section shows results and discussion where we expose our remarks, conclusions and future work.
+Section \ref{Classification}  presents the Speller Matrix Letter Identification procedure including the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008} and the final Section shows results and discussion where we expose our remarks, conclusions and future work.
 %This kind of EEG events are described in~\citep{Hartman2005} and~\citep{Ben-Simon2013}.
+
 \section{Materials and Methods}
@@
 one complete sequence of intensification of each of the $6$ rows plus the $6$ columns, a basic artefact elimination procedure is implemented by removing the entire sequence when any signal deviates above/bellow $ \pm 70 \mu V $.
-\item \textbf{Segmentation}: For each of the $12$ stimuli,  a window of $1$ second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
+\item \textbf{Segmentation}: For each of the $12$ stimuli,  a window of $t_{max} = 1$ second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
 Two of these segments should contain the P300 ERP signature time-locked to the flashing stimulus, one for the row, and one for the column.
@@
 \begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, t \in \{0, 1, \dots, F_s\},c \in \{1,2,\dots,8\}
+\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, t \in \{0, 1, \dots, t_{max}\},c \in \{1,2,\dots,Ch\}
+\label{eq:standarizedaverages}
 \end{equation}
@@
 \noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related with the image scale.
-On the other hand, $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for channel $c$.
-The parameter $F_s$ is the sampling frequency.
+Moreover, $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for channel $c$.
+The parameter $F_s$ is the sampling frequency and $Ch$ is the number of available EEG channels.
 Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
@@
 
-\noindent where $ (z_1,z_2) \in \mathcal{N} \times \mathcal{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $t \in \{0, 1, \dots, F_s\}$ and $c \in \{1,2,\dots,8\}$.
-The function $z(c)$, $c \in \{1,2,\dots,8\}$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
+\noindent where $ (z_1,z_2) \in \mathcal{N} \times \mathcal{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $t \in \{0, 1, \dots, F_s\}$ and $c \in \{1,2,\dots,Ch\}$.
+The function $z(c)$, $c \in \{1,2,\dots,Ch\}$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
 
 \begin{equation}
-z(c) = \left \lfloor{ \frac{\max \tilde{x}(t,c)  - \min \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max \tilde{x}(t,c)  + \min \tilde{x}(t,c)}{ 2} }\right \rfloor
+z(c) = \left \lfloor{ \frac{\max_{0 \leq t\leq t_{max}} \tilde{x}(t,c)  - \min_{0 \leq t\leq t_{max}} \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{0 \leq t\leq t_{max}} \tilde{x}(t,c)  + \min_{0 \leq t\leq t_{max}} \tilde{x}(t,c)}{ 2} }\right \rfloor
 \label{eq:zerolevel}
@@
 A local image patch of size $S_p \times S_p$ pixels is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
-It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, then $S_p = 4.3s $ pixels.
+It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, thus $S_p = 4.3s $ pixels.
 
@@
 \begin{enumerate}
-\item \label{paso1}Highlight randomly the rows and columns from the matrix.
-There is one row and one column that should match the letter selected by the subject.
-For each illuminated row or column, an image is obtained.
-\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t \in \{0, 1, \dots, F_s\}$ and $c \in \{1,2,\dots,8\}$ correspond to time and channel, respectively.
+\item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
+\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t \in \{0, 1, \dots, t_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to time and channel, respectively.
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
@@
 \begin{equation}
-x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c), t \in \{0, 1, \dots, F_s\}, c \in \{1,\dots,8\}
+x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c), t \in \{0, 1, \dots, t_{max}\}, c \in \{1,\dots,Ch\}
 \label{averaging}
@@
 
-\item \label{paso4}Plot the signal $x(t,c)$,  $t \in \{0, 1, \dots, F_s\}$, $c \in \{1,\dots,8\}$,  according Section~\ref{Plot}.
+\item \label{paso4}Plot the signal $x(t,c)$,  $t \in \{0, 1, \dots, t_{max}\}$, $c \in \{1,\dots,Ch\}$,  according Section~\ref{Plot}.
 
@@
 All healthy subjects have normal or corrected-to normal vision and no history of neurological disorders.
-The experiment is performed with 20 subjects, 17 males, 3 females, average age 26.85 years, standard deviation  8.8 years, range 19-50 years.
+The experiment is performed with 20 subjects, 6 males, 2 females, 6 right-handed, 2 left-handed, average age 28.25 years, standard deviation  9.58 years, range 20-50 years.
 
@@
 Participants are seated in a comfortable chair, with their vision aligned to a computer screen located one meter in front of them.
-The handling and processing of the data and stimulations is handled by the OpenVibe platform~\citep{Renard2010}.
+The handling and processing of the data and stimulations is conducted by the OpenVibe platform~\citep{Renard2010}.
 
@@
 
-% Antes estaba esto aca
-
-% The P300 event latency and amplitude vary greatly between subjects,
-
-% lo saque porque confunde, ya que despues de todo, dejamos todo fijo, e ignoramos completamente el hecho de que varie.
-
-%Parameters are selected according to the experimental protocol. 
 The patch size is $S_P = 3s.4 \times 3s.4$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
-The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $.
+The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $~\citep{Rao2013}.
 Hence it is necessary to utilize a size patch $S_P$ that could capture an entire transient event.
 With this purpose in consideration, the $s$ value election is essential.
+
 %necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
@@
 This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ on the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds on the horizontal direction.
-Finally, descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ (see Eq.~\ref{eq:zerolevel}).
+Descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ (see Eq.~\ref{eq:zerolevel}).
 In this way the whole transient event is captured.
@@
 The vertical size consists of $4$ blocks $B_{i,j}$  of size $3 s_y$ pixels which is long enough as to contain the signal $\Delta  \mu V $, the peak-to-peak amplitude of the transient event.
-The horizontal size on the other hand, includes $4$ blocks  of $3 s_x$ and cover the entire duration in seconds of the transient signal event, $ \lambda $.
+The horizontal size on the other hand, includes $4$ blocks  of $3 s_x$ and covers the entire duration in seconds of the transient signal event, $ \lambda $.
 }
@@
 
-%Hacemos algún comentario de k_a?
+Lastly, the number of channels $Ch$ is equal to $8$ for both datasets, and the number of intensification sequences $k_a$ is statically assigned to $10$.
+The parameter $k$ used in the Near Neighbor $NN_T(d^l_u), \;l\in\{row,{col}\}$ is set to $k=7$.
+In addition, the distance function used on  equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol} was the cosine norm, and descriptors were normalized to $ \left[ -1, 1 \right] $.
 
@@
 
+The P300 ERP  consists of two overlapping components the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\citep{Polich2007} then, the classical approach finds the stronger response on the central channel Cz~\citep{Riccio2013}.
+However, Wolpaw et al~\citep{Krusienski2006} show that the response may also arise in occipital channels.
+In our approach, occipital channels PO8 and PO7 show higher performances for some subjects.
+%For example in the articles~\citep{Huggins2016,Jure2016} the authors show that channels PO8 and PO7 have higher performances.
+
+%The authors~\citep{Riccio2013} that analyzed the public dataset concluded that Cz is the best performing channel, more representative of the difference in amplitudes.
+However, in our approach, occipital channels,  PO8 and PO7 show higher performances~\citep{Huggins2016,Jure2016}.
+
 Table~\ref{tab:results} shows the results of applying the algorithm to the subjects of the public dataset of ALS patients.
 The percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
-From the seven words for each subject, the first three were used as training, and the remaining four for testing.
+From the seven words for each subject, the first three are used as training, and the remaining four for testing.
 The best performing channel is informed as well.
 The chance level is $2\%$.
+It can be observed that the best performance of the letter identification method is reached in various channels depending on the subject in study.
 
 \begin{table}[H]
-\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject.  }
+\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject of the public dataset 008-2014~\citep{Riccio2013}.
+Twenty letters, four words of five letters each, were used for testing.
+}
 \centering
@@
 
-%RR: esto corresponde al punto 4.1 que dijo el revisor y tiene que ver con proveer informacion del canal en el que la clasificacion dio mejor.
+In Table~\ref{tab:resultsowndataset} results obtained for the 8 healthy subjects are shown.   It shows the results obtained by the present method as described in previous sections.
 
-The P300 wave which was traditionally considered as one component, consists of two overlapping components the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\citep{Polich2007}.
-Although the traditional approach is to expect to find a stronger response on the central channel Cz~\citep{Riccio2013}, Wolpaw et al.
-shows that the response may also arise in occipital channels as well.
-The authors~\citep{Riccio2013} that analyzed the public dataset concluded that Cz was the best performing channel, more representative of the difference in amplitudes.
-However, in our approach, occipital channels,  PO8 and PO7 showed higher performances~\citep{Huggins2016,Jure2016}.
+%We used an additional subject as control as he was instructed to count only the letters and to stare at the screen but  neglecting row/column flashing, which achieved zero performance.
+We also validated the procedure by visualizing clearly the SSVEP generated by the 4 Hz flashings.
 
-%JG ITR y BTR nunca fueron definidos
+\begin{table}[H]
+\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each healthy subject.
+The method is applied on a channel by channel basis, and the best performing channel is shown.
+}
+\centering
+%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
+\begin{tabular}{cccc}
+\toprule
+\textbf{Participant}	&  \textbf{BPC}	& \textbf{Performance}\\
+\midrule
+1     &     Oz   &     $40\%$  \\
+2     &     P07   &     $30\%$ \\
+3     &     P4   &     $40\%$ \\
+4     &     P4   &     $45\%$ \\
+5     &     P4   &      $60\%$ \\
+6     &     Pz   &      $50\%$ \\
+7     &     PO7   &      $70\%$ \\
+8     &     P4   &      $50\%$ \\
+
+\bottomrule
+\end{tabular}
+\label{tab:resultsowndataset}
+\end{table}
+
 The Information Transfer Rate (ITR), or Bit Transfer Rate (BTR), in the case of reactive BCIs~\citep{WolpawJonathanR2012}  depends on the amount of signal averaging required to transmit a valid and robust selection.
 Fig.~\ref{fig:performance} shows the performance curves.
-It can be observed that the percentage of correctly identified letters depends on the number of repetitions that are used to obtain the averaged signal.
-For both datasets, the experimental protocol uses a very short ISI which has the potential to increase the ITR but at the same time it is inversely proportional to the amplitude of the P300 response (i.e.
-harder to detect).
-As an unexpected consequence, $15$ out of $20$ subject reported high ocular tiredness by using this configuration.
+It can be observed that the percentage of correctly identified letters depends on the number of intensification sequences $k_a$ that are used to obtain the averaged signal.
 
@@
 \includegraphics[width=10cm]{performance.eps}
-\caption{Performance curves for the eight subjects included in the ALS dataset.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
+\caption{Performance curves for the eight subjects included in the dataset of ALS patients.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
 \label{fig:performance}
@@
 %%JG No entiendo nada
-We find that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that could allow the classification to reach a higher level.
-Moreover, as can be seen in the figure, as repetitions tend to 1, which will correspond to single trial letter identification, the performance suffers greatly as the signal plot is  too complex to be used as template and the method is completely unable to find good matches
+%We find that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that allow  to reach a higher level classification.  
+As can be seen in the figure, as the number of intensification sequences tend to 1, which corresponds to single trial letter identification, the performance is reduced.
+As mentioned before, the SNR of the single trial P300 is very low.
+The ensemble average is carried out in order to improve the SNR but other problems appear, for example inter-trial variability due to latencies shifts that may produce an unstable P300 signature.
+To solve this problem, we tested an approach to assess if the morphological shape of the P300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied, but results are still in process.
+We also applied DTW~\citep{Casarotto2005} but we were unable to find a substantial improvement.
+On the other hand amplitude width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals\citep{Ouyang2017}.
+This is another source of instability of the P300 signature component that may need to be considered.
 
-We hypothesize that as subjects may have different latencies, amplitudes and width of the P300 signal \citep{Riccio2013}, it may also be the case that the shape of the generated ERP may vary greatly in an intra-subject manner, thus the pattern could not be well generalized and a higher performance could not be reached (Fig.
-\ref{fig:p300templates}).
-It is known (REFERENCE) that the point-to-point averaging suffers from inter-trial variability (REFERENCIA A ESE PAPER Con ese nombre) due to latencies shifts.
-We tested a naive approach to verify if the morphological shape of the P300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied.
-We also applied DTW but couldn't found a substantial improvement.
-Amplitude width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals.
-
-Considering the amplitude variation of the P300 and its components, we proposed a conservative approach by standardizing the signal (which can have both effects, of reducing noise that were not reduced by the averaging procedure but at the same time it can reduce the amplitude of a significant p300 component).
-The standardization also provides another benefit which is sometimes neglected as unimportant.
-The variance of the averaged signal depends on the number of signals averaged.
-As this method analyze the shape of the signal in the plot, which is  affected by the variance, the method is not be able to discriminate solely based on the underlying cognitive event, it is confused by the number of signals that are used to calculate the averaged signal.
-
-The SIFT method is invariant to the scale, but in order to allow our method to be also invariant to the amplitude of the signal we had to modify the geometry of the descriptor to remove the squared constrain.
-
-Figure~\ref{fig:p300templates} shows the P300 templates patches for subjects 8 and 3 from the public dataset.
+As subjects may have different \textit{latencies}, \textit{amplitudes} and \textit{width} of their P300 components, they may also have distinct \textit{shapes} of the generated ERP.
+Figure~\ref{fig:p300templates} shows the P300 templates patches for patients 8 and 3 from the public dataset.
 It can be observed that in coincidence with the performance results, the P300 signature is more clear and consistent for subject 8 (Fig.~\ref{subject8})  while for subject 3 (Fig.~\ref{subject3}) the characteristic pattern is more difficult to perceive.
@@
 \subfigure[Subject 3.]{\includegraphics[width=10cm]{subject3.png}\label{subject3}}
-\caption{Ten P300 template patches for subjects 8 and 3. }
+\caption{P300 template patches for subjects 8 and 3. As traditional done in neuroscience research, downward is positive. }
 \label{fig:p300templates}
@@
 
-In Table~\ref{tab:resultsowndataset} results obtained for the 8 healthy subjects is shown.
-It shows the results obtained by the present method as described in previous sections.
-We used an additional subject as control as he was instructed to count only the letters and to stare at the screen but  neglecting row/column flashing, which achieved zero performance.
-We also validated the procedure by visualizing clearly the SSVEP generated by the 4 Hz flashing of lights.
-\begin{table}[H]
-\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each healthy subject. Chance level is $2\%$. }
-\centering
-%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{cccc}
-\toprule
-\textbf{Participant}	&  \textbf{BPC}	& \textbf{Performance}\\
-\midrule
-1     &     Fz   &     $40\%$  \\
-2     &     Oz   &     $30\%$ \\
-3     &     Cz   &     $40\%$ \\
-4     &     Cz   &     $45\%$ \\
-5     &     P4   &      $60\%$ \\
-6     &     P3   &      $50\%$ \\
-7     &     PO7   &      $70\%$ \\
-8     &     PO8   &      $50\%$ \\
+%\begin{itemize}
+%\item Lo que pusimos antes de la sincronización de las señales apunta a la latencia.
+%\item Por el lado de la amplitud, lo que hicimos fue ir por un lado conservador de llevar todo a varianza uno.
+%\item Si bien el zscore nos ayuda también genera otro problema adicional que es el que ponemos en el texto (y por el cual con razón Juan me criticaba).
 
-\bottomrule
-\end{tabular}
-\label{tab:resultsowndataset}
-\end{table}
+%\end{itemize}
 
-The method beneficial aspect is that its single channel approach and also the most important part is that the method allows close collaboration because templates can be selected by visually inspecting them.
+%La estandarizacion produce dos efectos contrarios entre si, contrapues.
+
+%Por un lado reduce el ruido y outliers
+%por otro lado reduce aquellos p300 significativos positivos que ayudarian su identificacion.
+
+Another problem is the amplitude variation of the P300.
+We propose an approach by standardizing the signal, shown in Eq.~\ref{standarizedaverages}.
+It first has the effect of normalizing the peak-to-peak amplitude, moderating its variation.
+Additionally,  it has the advantage of reducing noise that were not reduced by the averaging procedure.
+In addition, the signal variance depends on the number of single-trials used to compute it \citep{van2006signal}.
+The standardizing process converts the signal to unit signal variance which makes it independent of the number $k_a$ of signals averaged.
+This is another advantage of this process.
+On the downsize, the standardizing process reduces the amplitude of any significant P300 complex diminishing its automatic interpretation capability.
+
+For both datasets, the experimental protocol uses a very short inter-stimulus interval which has the potential to increase the ITR but at the same time it reduces the amplitude of the P300 response, hence it may be more difficult to detect~\citep{Rao2013}.
+
+%As an unexpected consequence, $15$ out of $20$ subject reported high ocular tiredness by using this configuration.
 
@@
 
-A method to characterize and classify EEG signals where their characterization is transient in time-space, like the P300 ERP, has been presented.
+%QUE HACER
+
+% Escribir conclusiones buenas sobre todo los mismos puntos que se tocan en la discusion
+% Si bien
+
+% El tema del estudio lo mismo que hace el medico pero de manera automatico.
+
+% Letter.  Esto esta aca.
+
+%The SIFT method is invariant to the scale, so in order to allow our method to be also invariant to the amplitude of the signal we had to modify the geometry of the patch to remove the squared constrain.
+
+The method beneficial aspect is that its single channel approach and also the most important part is that the method allows close collaboration because templates can be selected by visually inspecting them.
+
+A method to characterize and classify EEG signals where their characterization is transient in time-space, like the P300 ERP, is presented.
 
@@
 The adaptive behavior of the algorithm make it well suited when the shape of the pattern elicited by the P300 response does not conform to the predicted structure.
+
 This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e.
@@
 It may be worthy of further consideration to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
+
+Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
+It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
 

### 2017-12-19 response.letter.tex
Revisé varios puntos, reescribí partes, agregué más referencias y completé los Resultados y la Discusión.  Queda pendiente las conclusiones finales

@@
 \usepackage{url,booktabs}
+\usepackage[round]{natbib}
 
@@
 ``EEG characterization and classification based on image gradient histograms''\\
-(now ``Oriented Gradient Histogram applied to P300 Detection'')}
+(now ``Histogram of Gradient Orientations of Signal Plots applied to P300 Detection'')}
 \author{}
@@
 
-We are grateful to the reviewers for pointing at relevant issues in our manuscript.
+We are grateful to the reviewer for pointing out relevant issues in our manuscript.
 
@@
 
-\section*{\fbox{Reviewer \#1:}}
+\section*{\fbox{Reviewer \#1 Transcript:}}
 
@@
 You're  right, thank you very much for your comment.
-We published in Ramele et al 2016 the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to decode Motor Imagery.
-We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and to decode SSVEP patterns.
-We added this information on the article (section INTRODUCTION and section CONCLUSION).
-We hope this new information enrich the article and clarifies this point.
+We published in ~\citep{Ramele2016} the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to classify Motor Imagery.
+We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and classification of  SSVEP patterns.
+We added this information on the article (Introduction and Conclusion, section 3).
+We hope this new information enriches the article and clarifies this point.
 
@@
 We worked hard to improve all the mathematical description of the method.
-Section FEATURE was modified as well as section PARAMETERS.
-Section SIGNAL PLOT was revised and its mathematical terminology were corrected.}
+The issue with variable T was fixed and we improved on the font styling to differentiate vectors and scalars.
+Section 1.1(Feature Extraction...) was modified as well as section 1.2.3(Parameters).
+The mathematical terminology were corrected all along the article.}
 \end{quotation}
@@
 {\color{blue}
-We have revised the manuscript and corrected equations (7) and (8) which specify the ration between the signal properties and the SIFT's geometric parameters. 
+We have revised the manuscript and modified equations (7) and (8) (now (8) and (9)) which are located on Section 1.2.3 (Parameters).
+We added more information on the text and specified subscript indexes to avoid the inconsistency.
 We thank you for noticing this.
@@
 Thank you for the observation.
-We modified all equations to emphasize the point that they are calculated in digital space.
-We also added that image gradients are also calculated using finite differences.
+We modified all equations to emphasize the point that they are calculated in digital space, particularly on equation (4).
+We also added that image gradients are also calculated using finite differences, and added information about the support of each equation to clarifies that they are actually calculated in digital space.
 With this, we hope we have addressed your point.
@@
 Thank you for the observation. 
-We fully addressed your point.
  }
@@
 Thank you for your suggestion.
-The original dataset was published at the BNCI-Horizon website, which is very popular in BCI research.
-Our approach emphasize the structure of the signal in terms of their morphological components in their visual description.% THIS IS TOUGH TO JUSTIFY
-In order to address this issue, we included a new dataset that we generated ourselves, on healthy subjects that we hope now improves the number of different subjects and that can help to validate the method.
+The original dataset was published as the 004-2014 dataset at the BNCI-Horizon website, which is very popular in BCI Research.
+As far as we are aware of, experiments on this dataset of ALS patients has not been published by any independent group.
+In order to address this issue, we included a new own dataset with the same experimental conditions that we generated ourselves, on healthy subjects.
+We believe it now improves the number of different subjects and that can help to validate the method.
+We are also publishing this dataset on the kaggle platform for replication.
 }
@@
 {\color{blue}
-THIS IS EVEN TOUGHER.
-
-ACA LE PUEDO PONER QUE HAY POCOS METODOS QUE PERMITAN ENGANCHAR BIEN QUE EL METODO
-
-All these points were modified according to the suggestions.
+We added more detailed information that address this issue.
+First in Section 1.1.4 (Speller Matrix Letter Identification) we included information about how we approach the signal averaging procedure.
+In Section 1.2.3 (Parameters) we added the parameters that were used.
+More importantly, in Section 2(Results and Discussion) we added information and references about the tests that we did to address latency and amplitude issues.
+We included in Section 3(Conclusion) information about future and current works where we are also dealing with these problems.
 We hope this is acceptable.
@@
 {\color{blue}
-The Results section was modified to clarify more precisely the expectation about the EEG channels where a visual P300 detection task would be obtained.
+We apologize for our mistake.
+The Results section was modified to clarify more precisely the expectation about the EEG channels where a visual P300 oddball paradigm generates stronger signals.
 We added references and improved the section to clarify this point more thoughtfully.
-We hope this is acceptable.
+We hope we may have addressed your point with this.
 }
@@
 Thank you a lot for this important comment.
-We re-read all the manuscript and asked to collegues to proofread the article.
+We re-read all the manuscript and asked  colleagues to proofread the article.
 They gave us interesting suggestions that helped us to improve different sections.
@@
 Thank you very much for this information.
-We reviewed completely that equation (it is now equation (6) and equation (7)) and we included a step-by-step procedure that describes the algorithm that we propose and we used.
+We reviewed completely that equation (it is now equation (6) and equation (7)) and we included a step-by-step procedure that describes the algorithm that we propose and we used as well as a schematic chart in Figure 3.
 }
@@
 
+\bibliography{response.letter}
+
 \end{document}

### 2017-12-19 article.tex
Revisé varios puntos, reescribí partes, agregué más referencias y completé los Resultados y la Discusión.  Queda pendiente las conclusiones finales

@@
 BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon~\citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
-New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments~\citep{Huggins2016}.
+New developments have overcome the out-of-the-lab high-bar and they are starting to be used in real world environments~\citep{Guger2017,Huggins2016}.
 However, they still lack the necessary robustness, and its performance is well behind any other method of human computer interaction, including any kind of detection of residual muscular movement~\citep{Clerc}.
 
-Several works have explored the idea of exploiting the signal waveform to analyze the EEG signal.
+A few works have explored the idea of exploiting the signal waveform to analyze the EEG signal.
 In~\citep{Alvarado-Gonzalez2016} an approach based on Slope Horizontal Chain Code is presented, whereas in ~\citep{Yamaguchi2009} a similar procedure was implemented based on Mathematical Morphological Analysis.
@@
 
-Averaged signal segments are scaled and standardized by
+Averaged signal segments are standardized and scaled by
 
@@
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
-\item \label{paso3} Compute  
+\item \label{paso3} Compute the Ensemble Average by
 \begin{equation}
@@
 By computing the aforementioned equations, the letter of the matrix can be determined from the intersection of the row $ \hat{row} $ and column $ \hat{col} $. 
-Figure~\ref{fig:classification} shows a schema of this process. 
+Figure~\ref{fig:classification} shows a scheme of this process. 
 \begin{figure}[htp]
@@
 All healthy subjects have normal or corrected-to normal vision and no history of neurological disorders.
-The experiment is performed with 20 subjects, 6 males, 2 females, 6 right-handed, 2 left-handed, average age 28.25 years, standard deviation  9.58 years, range 20-50 years.
+The experiment is performed with 8 subjects, 6 males, 2 females, 6 right-handed, 2 left-handed, average age 28.25 years, standard deviation  9.58 years, range 20-50 years.
 
@@
 
-\noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sample frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
+\noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sampling frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
 The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.
@@
 The chance level is $2\%$.
-It can be observed that the best performance of the letter identification method is reached in various channels depending on the subject in study.
+It can be observed that the best performance of the letter identification method is reached in various channels depending on the subject on study.
 
@@
 
-In Table~\ref{tab:resultsowndataset} results obtained for the 8 healthy subjects are shown.   It shows the results obtained by the present method as described in previous sections.
+In Table~\ref{tab:resultsowndataset} results obtained for 8 healthy subjects are shown.   These results were obtained by the present method as described in previous sections.
 
@@
 The Information Transfer Rate (ITR), or Bit Transfer Rate (BTR), in the case of reactive BCIs~\citep{WolpawJonathanR2012}  depends on the amount of signal averaging required to transmit a valid and robust selection.
-Fig.~\ref{fig:performance} shows the performance curves.
+Fig.~\ref{fig:performance} shows the performance curves for varying intensification sequences.
 It can be observed that the percentage of correctly identified letters depends on the number of intensification sequences $k_a$ that are used to obtain the averaged signal.
@@
 
-%\begin{itemize}
-%\item PODRIA ACA MOSTRAR ADEMAS PLOTS DE SINGLE TRIAL PARA QUE VEAN QUE SON UN QUILOMBO.%JG no sé, no lo pidió
-%\item UNA DE LAS RAZONES POR LA CUAL PODEMOS LLEGAR A ARGUMENTAR QUE NO FUIMOS MAS A FONDO CON EL TEMA ES QUE NO QUERIAMOS DISTORSIONAR DEMASIADO LA SENAL COMPLEJIZANDO EL ALGORITMO Y QUERIAMOS MANTENERNOS ENTRE LO MAS ESTANDAR POSIBLE (PUEDO CITAR MAS O MENOS QUE NUESTRO PROCESSING PIPELINE ES ESTANDARD).%%JG No, a menos que lo pregunte.
-Tenemos que transmitir que fuimos a fondo en el tema
-%\item ACA EL PUNTO PARA MOSTRAR ES QUE DE ALGUNA MANERA ESTAMOS COMO MOSTRANDO UN "SI SE PUEDE" DE UN TRABAJO QUE PUEDE CONTINUAR Y EXPANDIRSE.
-POR ESO LA IDEA DE MOSTRAR PUNTAS DE COSAS QUE VIMOS E HICIMOS, PERO SIN MOSTRAR RESULTADOS CONCRETOS.
-%\item La pregunta:  ¿ Es posible utilizando solamente la información de los plots, aún utilizando preprocesamiento y una mejora en el SNR basado en el esquema tradicional de generación del patrón de p300, detectarlo con alto grado de éxito ?
-%%JG esto puede ir a las conclusiones.
-
-%\begin{itemize}
-%\item Es posible pero sólo en algunos pacientes.
-%\item Se requiere de mecanismos para estabilizar el waveform de p300.
-
-%\end{itemize}
-
-%\item 
-%\end{itemize}
-%%JG No entiendo nada
-%We find that only by using the visual aspect of the P300 signal plot,  for some subjects it was not possible to find templates that allow  to reach a higher level classification.  
 As can be seen in the figure, as the number of intensification sequences tend to 1, which corresponds to single trial letter identification, the performance is reduced.
 As mentioned before, the SNR of the single trial P300 is very low.
-The ensemble average is carried out in order to improve the SNR but other problems appear, for example inter-trial variability due to latencies shifts that may produce an unstable P300 signature.
+The ensemble average is carried out in order to improve the SNR but other problems arise, for example inter-trial variability due to latencies shifts that may produce an unstable P300 signature.
 To solve this problem, we tested an approach to assess if the morphological shape of the P300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied, but results are still in process.
 We also applied DTW~\citep{Casarotto2005} but we were unable to find a substantial improvement.
-On the other hand amplitude width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals\citep{Ouyang2017}.
-This is another source of instability of the P300 signature component that may need to be considered.
+On the other hand amplitude, width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals\citep{Ouyang2017}.
+This is another source of instability of the P300 signature component that may need to be addressed.
 
@@
 Another problem is the amplitude variation of the P300.
-We propose an approach by standardizing the signal, shown in Eq.~\ref{standarizedaverages}.
-It first has the effect of normalizing the peak-to-peak amplitude, moderating its variation.
-Additionally,  it has the advantage of reducing noise that were not reduced by the averaging procedure.
-In addition, the signal variance depends on the number of single-trials used to compute it \citep{van2006signal}.
+We propose an approach by standardizing the signal, shown in Eq.~\ref{eq:standarizedaverages}.
+First, it has the effect of normalizing the peak-to-peak amplitude, moderating its variation.
+It has also the advantage of reducing noise that were not reduced by the averaging procedure.
+It is important to remark that the signal variance depends on the number of single-trials used to compute it \citep{van2006signal}.
 The standardizing process converts the signal to unit signal variance which makes it independent of the number $k_a$ of signals averaged.
-This is another advantage of this process.
+This is another advantage of this approach.
 On the downsize, the standardizing process reduces the amplitude of any significant P300 complex diminishing its automatic interpretation capability.
 
-For both datasets, the experimental protocol uses a very short inter-stimulus interval which has the potential to increase the ITR but at the same time it reduces the amplitude of the P300 response, hence it may be more difficult to detect~\citep{Rao2013}.
+For both datasets, the experimental protocol uses a very short inter-stimulus interval which has the potential to increase the ITR but at the same time it reduces the amplitude of the P300 response, hence it may be more difficult to detect it~\citep{Rao2013}.
 
@@
 
-%QUE HACER
+A method to detect transient P300 components from EEG signals based on their characterization in time-space is presented.
 
-% Escribir conclusiones buenas sobre todo los mismos puntos que se tocan en la discusion
-% Si bien
+Stability of the P300 in terms of shape: Synchronization averaging, increase inter-stimulus interval, different montages, spatial filters.
+Collaboration
+It can work adaptitevely and also from a set of preconcepts.  This can be used to build a database of descriptors and improve atlases (REFERENCE ATLAS OF EEG).
 
-% El tema del estudio lo mismo que hace el medico pero de manera automatico.
+The SIFT method is invariant to the scale, so in order to allow our method to be also invariant to the amplitude of the signal we had to modify the geometry of the patch to remove the squared constrain.
+This will allow us to not use the zscore anymore.
 
-% Letter.  Esto esta aca.
-
-%The SIFT method is invariant to the scale, so in order to allow our method to be also invariant to the amplitude of the signal we had to modify the geometry of the patch to remove the squared constrain.
+%Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).
 
 The method beneficial aspect is that its single channel approach and also the most important part is that the method allows close collaboration because templates can be selected by visually inspecting them.
-
-A method to characterize and classify EEG signals where their characterization is transient in time-space, like the P300 ERP, is presented.
-
-%Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).
 
@@
 \textit{how actually they looked like}) for the training and calibration step, and they do not require any prior knowledge about the signal.
+
 In contrast, when the shape of the ERP response is not consistent across the same subject, this method would not be able to find the templates and speller performance could be penalized.
 
-At the same time, by analyzing the generated descriptors, which map in a very detailed and synthetic structure the shape information contained within the patch, a metric about the consistency of the shape of the generated P300 response could also be derived.
+At the same time, by analyzing the generated descriptors, which map in a very detailed and synthetic structure the shape information contained within the patch, a metric about the consistency of the \textit{shape} of the generated P300 response could also be derived.
+
 It may be worthy of further consideration to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
@@
 It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
+Althought this, we did not find such an evidence in terms of the shape.
+The best response was obtained by a ALS subject.
 
@@
 We are currently working in unpublished material analyzing KComplex that may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods, frequent in sleep research.
+
+Finally, waveform analysis and the methods that can be used to analyze them is an area where active research can benefit greatly both, physicians who need to automate some of the tools and practices and Brain Computer Interfaces or automatic EEG processing by using the extensive knowledge body available for clinical practices.
+APUNTAR AL PAPER DE LOTTE MUY BUENO DEL JOURNAL DE BCI.
 

### 2017-12-19 article.tex
Casi todo listo.

@@
 % Abstract (Do not use inserted blank lines, i.e. \\) 
-\abstract{%EL OBJETIVO DE ESTE PAPER, EL HILO CONDUCTOR, ES DEMOSTRAR QUE ESTE METODO QUE ESTA BASADO EN LA ESTRUCTURA DE LA SENIAL SE PUEDE UTILIZAR TAMBIEN PARA IDENTIFICAR EVENTOS TRANSIENTES COMO EL P300, Y QUE CON ESO SE PUEDE IMPLEMENTAR UN SPELLER DE P300.
-The analysis of Electroencephalographic (EEG) signals is of ulterior importance to elucidate patterns that could improve the implementation of Brain Computer Interfaces (BCI).
+\abstract{The analysis of Electroencephalographic (EEG) signals is of ulterior importance to elucidate patterns that could improve the implementation of Brain Computer Interfaces (BCI).
 These systems are meant to provide alternative pathways to transmit volitional information which could potentially enhance the quality of life of patients affected by neurodegenerative disorders or improve Human Computer Interaction systems.
@@
 
-%Patients suffering from ALS their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
-
 BCI has gained mainstream public awareness with worldwide challenge competitions like Cybathlon~\citep{Riener2014} and even been broadcasted during the inauguration ceremony of the 2014 Soccer World Cup.
@@
 
-%The underlying idea of this method is to generate templates for the signal \citep{Ramele2016}, based on the image plot.  Hence, the first step is the processing of the EEG signal.
 In this section, the signal preprocessing, the method for generating images from signal plots, the feature extraction procedure and the speller matrix identification are described.
@@
 %necesitamos definir el valor de s en función de los parámetros de la señal, de modo tal que el parche cubra el evento completo.  
-We propose the equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
+We propose the Equations~\ref{eq:mapping2} and~\ref{eq:mapping1} to compute the scale value in horizontal and vertical directions, respectively. 
 \begin{equation}
@@
 The parameter $k$ used in the Near Neighbor $NN_T(d^l_u), \;l\in\{row,{col}\}$ is set to $k=7$.
-In addition, the distance function used on  equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol} was the cosine norm, and descriptors were normalized to $ \left[ -1, 1 \right] $.
+In addition, the norm used on  Equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol} was the cosine norm, and descriptors were normalized to $ \left[ -1, 1 \right] $.
 
@@
 
-The P300 ERP  consists of two overlapping components the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\citep{Polich2007} then, the classical approach finds the stronger response on the central channel Cz~\citep{Riccio2013}.
+The P300 ERP  consists of two overlapping components: the P3a and P3b, the former with frontocentral distribution while the later stronger on centroparietal region~\citep{Polich2007} then, the classical approach finds the stronger response on the central channel Cz~\citep{Riccio2013}.
 However, Wolpaw et al~\citep{Krusienski2006} show that the response may also arise in occipital channels.
@@
 1     &     Oz   &     $40\%$  \\
-2     &     P07   &     $30\%$ \\
+2     &     PO7   &     $30\%$ \\
 3     &     P4   &     $40\%$ \\
@@
 \centering
-\includegraphics[width=10cm]{performance.eps}
+\includegraphics[width=10cm]{performance2.eps}
 \caption{Performance curves for the eight subjects included in the dataset of ALS patients.  Three out of eight subjects achieved the necessary performance to implement a valid P300 speller.}
@@
 To solve this problem, we tested an approach to assess if the morphological shape of the P300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied, but results are still in process.
-We also applied DTW~\citep{Casarotto2005} but we were unable to find a substantial improvement.
+We also applied Dynamic Time Warping (DTW)~\citep{Casarotto2005} but we were unable to find a substantial improvement.
 On the other hand amplitude, width and latency are generally affected by habituation, fatigue or level of attention and may lead to null signals\citep{Ouyang2017}.

### 2017-12-20 article.tex
Version DRAFT completa 1

@@
 In the article~\citep{Ramele2016},  the authors introduce a method for classification of rhythmic EEG events like Visual Occipital Alpha Waves  and Motor Imagery rolandic central $\mu$ rhythms using the histogram of gradient orientations of signal plots.
-Inspired in that work, we proposed a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.
+Inspired in that work, we propose a novel application of the developed method to classify and describe transient events, particularly the P300 Event Related Potential.
 The proposed approach is based on the waveform analysis of the shape of the EEG signal, but using histogram of gradient orientations.
@@
 This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradient Orientations of the Signal Plot method is explained.
-Section~\ref{Pipeline} describes the processing pipeline.
+Section~\ref{Pipeline} describes the preprocessing pipeline.
 Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.
@@
 \item \textbf{Signal Enhancement}: The preprocessing stage consists of the enhancement of the SNR of the P300 pattern above the level of basal EEG.
-The processing pipeline starts by applying a notch filter to the raw signal, a
+The processing pipeline starts by applying a notch filter to the raw digital signal, a
 $4$th degree $10$ Hz lowpass Butterworth filter and finally a decimation with a Finite Impulse Response (FIR) filter of order $30$ from the original sampling frequency down to $16$ Hz \citep{Krusienski2006}.
-\item \textbf{Artefact Removal}: The EEG signal matrix is processed on a channel by channel basis.
+\item \textbf{Artifact Removal}: The EEG signal matrix is processed on a channel by channel basis.
 For every 12 flashing stimuli, i.e.
@@
 \item \textbf{Signal Averaging}:  The P300 ERP is deeply buried under background EEG so the traditional approach to identify it is by point-to-point averaging the time-locked stacked signal segments.
-Hence the values which are not related to, and not time-locked to the onset of the stimulus are canceled out \citep{Liang2008}.
+Hence the values which are not related to, and not time-locked to the onset of the stimulus are canceled out~\citep{Liang2008}.
 \end{itemize}
@@
 \noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related with the image scale.
-Moreover, $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for channel $c$.
-The parameter $F_s$ is the sampling frequency and $Ch$ is the number of available EEG channels.
-Lastly, $\bar{x}(c)$ and $ \sigma (c) $ are the mean and standard deviation of $x$.
+In addition, $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for channel $c$.
+The parameter  $Ch$ is the number of available EEG channels.
+Lastly, $$\bar{x}(c) =\frac{1}{t_{max}}\sum_{t=1}^{t_{max}}x(t,c)$$ and $$ \sigma(c) = \frac{1}{t_{max}}\sum_{t=1}^{t_{max}}(x(t,c)-\bar{x}(c))^2 $$ are the mean and standard deviation of $x(t,c), t \in 1,\dots, t_{max}$, for each channel $c$.
 
@@
 
-\noindent where $ (z_1,z_2) \in \mathcal{N} \times \mathcal{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $t \in \{0, 1, \dots, F_s\}$ and $c \in \{1,2,\dots,Ch\}$.
+\noindent where $ (z_1,z_2) \in \mathcal{N} \times \mathcal{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $t \in \{0, 1, \dots, t_{max}\}$ and $c \in \{1,2,\dots,Ch\}$.
 The function $z(c)$, $c \in \{1,2,\dots,Ch\}$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
@@
 
-Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same locations Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.
+Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same locations Fz, Cz, Pz, P3,P4, Oz, PO7 and PO8.
 Reference is set to the right ear lobe and ground is preset as the AFz position.
@@
 
-%EEG data were recorded from the same 8 electrodes positions (C3, Cz, C4, CPz, P3, Pz, P4, POz) with all the systems.
-The reference positions were placed at the right ear.
-The ground positions was placed at the left ear, except with the g.Nautilus system (placed at AFz, because this could not be changed).
-Sampling rate was set at 250 Hz, the closest possible to predefined g.Mobilab sampling rate (256 Hz).
-Data were acquired notch filtered at 50 Hz and passband filtered between 2 Hz and 30 Hz.
-The filtering was performed via the proprietary Simulink HighSpeed Online Processing block modules for g.MOBIlab and g.Nautilus systems, and via MATLAB code for the Xpress system.
-Table~\ref{tab:results} summarizes the acquisition settings for each system.
+%Fz, Cz, P3, Pz, P4, PO7, PO8 and Oz.
 
@@
 This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ on the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds on the horizontal direction.
-Descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ (see Eq.~\ref{eq:zerolevel}).
+Descriptors  $\mathbf{kp}$  are located at $ (x_{kp}, y_{kp} )= ( 0.55 Fs.\gamma, z(c) )= (35,  z(c)) $ for the corresponding channel $c$ (see Eq.~\ref{eq:zerolevel}).
 In this way the whole transient event is captured.
@@
 The size of the patch is $S_p \times S_p$ pixels.
-The vertical size consists of $4$ blocks $B_{i,j}$  of size $3 s_y$ pixels which is long enough as to contain the signal $\Delta  \mu V $, the peak-to-peak amplitude of the transient event.
+The vertical size consists of $4$ blocks of size $3 s_y$ pixels which is long enough as to contain the signal $\Delta  \mu V $, the peak-to-peak amplitude of the transient event.
 The horizontal size on the other hand, includes $4$ blocks  of $3 s_x$ and covers the entire duration in seconds of the transient signal event, $ \lambda $.
@@
 
-Table~\ref{tab:results} shows the results of applying the algorithm to the subjects of the public dataset of ALS patients.
+Table~\ref{tab:results} shows the results of applying the algorithm to the subjects of the public dataset of ALS patients~\citep{Riccio2013}.
 The percentage of correctly spelled letters is calculated while performing an offline BCI Simulation.
@@
 \begin{table}[H]
-\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject of the public dataset 008-2014~\citep{Riccio2013}.
+\caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject of the public dataset 008-2014.
 Twenty letters, four words of five letters each, were used for testing.
@@
 
-A method to detect transient P300 components from EEG signals based on their characterization in time-space is presented.
+%In this paper, a new unsupervised method to enhance evoked response by target stimuli in an oddball paradigm was presented.
+Only given the time indexes of rows/columns intensifications, the proposed algorithm estimates the main components of the P300 subspace by providing the best SNR.
+It was shown to efficiently improve the quality of the evoked responses by taking into account the signal and the noise, as opposed to principal component analysis, which only considers the signal.
+Using this method to enhance P300 subspace before the BCI classification task speeds up the BCI since less words are required to train the spatial filters and the linear classifier, given a certain percentage of good symbol prediction.
+Moreover, using this spatial enhancement significantly reduces the dimension of the feature vector used to predict words.
 
-Stability of the P300 in terms of shape: Synchronization averaging, increase inter-stimulus interval, different montages, spatial filters.
-Collaboration
-It can work adaptitevely and also from a set of preconcepts.  This can be used to build a database of descriptors and improve atlases (REFERENCE ATLAS OF EEG).
+%RECORDAR CALIBRATION STEP
 
-The SIFT method is invariant to the scale, so in order to allow our method to be also invariant to the amplitude of the signal we had to modify the geometry of the patch to remove the squared constrain.
-This will allow us to not use the zscore anymore.
+A method to detect transient P300 components from EEG signals based on their waveform characterization in time-space was presented.
+Additionally its validity was verified against a public dataset of ALS patients and an own dataset of healthy subjects.
 
-%Databases of descriptors.  Flexibility of the descriptor in terms of configuration and size (reference variants of SIFT on lower dimensiones).
-
-The method beneficial aspect is that its single channel approach and also the most important part is that the method allows close collaboration because templates can be selected by visually inspecting them.
-
+This method can be both, based on predefined templates or adaptive by including a training or calibration step to identify the templates.
 The adaptive behavior of the algorithm make it well suited when the shape of the pattern elicited by the P300 response does not conform to the predicted structure.
 
-This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e.
-\textit{how actually they looked like}) for the training and calibration step, and they do not require any prior knowledge about the signal.
+The method works on a channel by channel basis.
+Should a calibration step be included, it can be used to identify the best performing channel, and may drastically reduce the number of required EEG electrodes, leading to the use of more ergonomic capturing device.
 
-In contrast, when the shape of the ERP response is not consistent across the same subject, this method would not be able to find the templates and speller performance could be penalized.
-
+We verified that the shape of the wave is more stable in occipital channels, where the performance at identifying letters  is higher.
+Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
+It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
+It may be worthy of further consideration to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
+Althought this, we did not find such an evidence in terms of the shape.
+The best response was obtained by an ALS subject.
+Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc}, i.e.
+to detect BCI illiteracy or to predict the achievable performance of a given method.
+In the original work~\citep{Riccio2013} where the dataset was published, a link between a task prediction test and the performance of the p300 was found.
 At the same time, by analyzing the generated descriptors, which map in a very detailed and synthetic structure the shape information contained within the patch, a metric about the consistency of the \textit{shape} of the generated P300 response could also be derived.
 
-It may be worthy of further consideration to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
+We also found that the stability of the P300 in terms of shape is crucial: Synchronization averaging, increase inter-stimulus interval, different montages, spatial filters can all of them affect the generated shape, and it is advisable to study which methods or combinations are best in order to stabilize the shape of the p300.
 
-Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
-It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
-Althought this, we did not find such an evidence in terms of the shape.
-The best response was obtained by a ALS subject.
-
+Clearly, the best characteristic of this method is that closer collaboration with physicians can be fostered.
 We believe that the expanding and the understanding of this tool in order to automatically classify those patterns in EEG that are specifically identified by their shapes (e.g.
@@
 We are currently working in unpublished material analyzing KComplex that may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods, frequent in sleep research.
+Additionally, it can be used as a tool for artifact removal (which is performed on many occasions by visually inspecting the signal).
+This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e.
+\textit{how actually they looked like}).
+In line with these applications,  it can be used to build a database~\citep{Chavarriaga2017} of descriptors and improve atlases~\citep{Hartman2005}.
 
-Finally, waveform analysis and the methods that can be used to analyze them is an area where active research can benefit greatly both, physicians who need to automate some of the tools and practices and Brain Computer Interfaces or automatic EEG processing by using the extensive knowledge body available for clinical practices.
-APUNTAR AL PAPER DE LOTTE MUY BUENO DEL JOURNAL DE BCI.
+In terms of the tool itself, the SIFT method is invariant to the scale, so in order to allow our method to be also invariant to the amplitude of the signal we had to modify the geometry of the patch to remove the squared constrain.
 
-Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc}, i.e.
-to detect BCI illiteracy or to predict the achievable performance of a given method, or even as a tool for artefact removal (which is performed on many occasions by visually inspecting the signal).
+Finally, waveform analysis and the methods that can be used to analyze them is an area where active research can benefit greatly both, physicians who need to automate some of the tools and practices and Brain Computer Interfaces or automatic EEG processing by using the extensive knowledge body available for clinical practices~\citep{Cole2017}
+
+%To further improve the performance of the P300-speller BCI, additional work should be considered.
+For instance, to better estimate the response evoked by target stimuli, a multi- stimuli model should be deemed by assuming that all the non- target stimuli – as well as the target stimuli – evoked specific responses.
+Finally, sensor selection should also be considered, in order to drastically reduce the number of required EEG electrodes, leading thus to a more ergonomic BCI.
 

### 2017-12-20 article.tex
Los cambios de hoy

@@
 
-This article unfolds as follows: in Section \ref{Feature} the Feature Extraction based on Histogram of Gradient Orientations of the Signal Plot method is explained.
-Section~\ref{Pipeline} describes the preprocessing pipeline.
-Section \ref{Plot}  clarifies how the image of the signal plot is constructed whereas Section \ref{SIFT}  describes in detail the feature extraction procedure.
-Section \ref{Classification}  presents the Speller Matrix Letter Identification procedure including the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008} and the final Section shows results and discussion where we expose our remarks, conclusions and future work.
+This article unfolds as follows: Section~\ref{Feature} is dedicated to explain the Feature Extraction method based on Histogram of Gradient Orientations of the Signal Plot: Section~\ref{Pipeline} shows the preprocessing pipeline,  Section~\ref{Plot}  describes the image generation of the signal plot, Section~\ref{SIFT}  presents the feature extraction procedure while  Section~\ref{Classification}  introduces the whole  Speller Matrix Letter Identification procedure including the classification algorithm based on Naive Bayes Nearest Neighbor (NBNN)~\citep{Boiman2008}.
+In Section~\ref{Protocol}, the experimental protocol is expounded.
+Section~\ref{Results} shows the results of applying the proposal technique and a discussion is presented.
+In the final Section~\ref{conclusion}  we expose our remarks, conclusions and future work.
 %This kind of EEG events are described in~\citep{Hartman2005} and~\citep{Ben-Simon2013}.
@@
 \begin{equation}
-\tilde{x}(t,c) = \left \lfloor{ \gamma \cdot \frac{( x(t,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, t \in \{0, 1, \dots, t_{max}\},c \in \{1,2,\dots,Ch\}
+\tilde{x}(n,c) = \left \lfloor{ \gamma \cdot \frac{( x(n,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, n \in \{ 1, \dots, n_{max}\},c \in \{1,2,\dots,Ch\}
 \label{eq:standarizedaverages}
@@
 \noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related with the image scale.
-In addition, $ x(t,c) $ is the point-to-point averaged EEG matrix for time $t$ and for channel $c$.
+In addition, $ x(n,c) $ is the point-to-point averaged EEG matrix for the sample point $n$ and for channel $c$, $n_{max}=F_s.t_{max}$ and $F_s$ is the sampling frequency.
 The parameter  $Ch$ is the number of available EEG channels.
-Lastly, $$\bar{x}(c) =\frac{1}{t_{max}}\sum_{t=1}^{t_{max}}x(t,c)$$ and $$ \sigma(c) = \frac{1}{t_{max}}\sum_{t=1}^{t_{max}}(x(t,c)-\bar{x}(c))^2 $$ are the mean and standard deviation of $x(t,c), t \in 1,\dots, t_{max}$, for each channel $c$.
+Lastly, $$\bar{x}(c) =\frac{1}{n_{max}}\sum_{n=1}^{n_{max}}x(n,c)$$ and $$ \sigma(c) = (\frac{1}{n_{max}-1}\sum_{n=1}^{n_{max}}(x(n,c)-\bar{x}(c))^2 )^{\frac{1}{2}}$$ are the mean and standard deviation of $x(n,c), n \in 1,\dots,n_{max}$, for each channel $c$.
 
@@
 
-\noindent where $ (z_1,z_2) \in \mathcal{N} \times \mathcal{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $t \in \{0, 1, \dots, t_{max}\}$ and $c \in \{1,2,\dots,Ch\}$.
+\noindent where $ (z_1,z_2) \in \mathbb{N} \times \mathbb{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$.
 The function $z(c)$, $c \in \{1,2,\dots,Ch\}$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
@@
 \begin{equation}
-z(c) = \left \lfloor{ \frac{\max_{0 \leq t\leq t_{max}} \tilde{x}(t,c)  - \min_{0 \leq t\leq t_{max}} \tilde{x}(t,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{0 \leq t\leq t_{max}} \tilde{x}(t,c)  + \min_{0 \leq t\leq t_{max}} \tilde{x}(t,c)}{ 2} }\right \rfloor
+z(c) = \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(n,c)  - \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(t,c)  + \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c)}{ 2} }\right \rfloor
 \label{eq:zerolevel}
@@
 \item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
-\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(t,c),\dots,S_{k_a}(t,c)$, where the variables $t \in \{0, 1, \dots, t_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to time and channel, respectively.
+\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(n,c),\dots,S_{k_a}(n,c)$, where the variables $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to time and channel, respectively.
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
@@
 \begin{equation}
-x(t,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(t,c), t \in \{0, 1, \dots, t_{max}\}, c \in \{1,\dots,Ch\}
+x(n,c)= \frac{1}{k_a}\sum_{i=1}^{k_a}S_i(n,c),n \in \{1, \dots, n_{max}\}, c \in \{1,\dots,Ch\}
 \label{averaging}
@@
 
-\item \label{paso4}Plot the signal $x(t,c)$,  $t \in \{0, 1, \dots, t_{max}\}$, $c \in \{1,\dots,Ch\}$,  according Section~\ref{Plot}.
+\item \label{paso4}Plot the signal $x(n,c)$,  $n \in \{1, \dots, n_{max}\}$, $c \in \{1,\dots,Ch\}$,  according Section~\ref{Plot}.
 
@@
 
-The patch size is $S_P = 3s.4 \times 3s.4$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+The patch size is $S_P = 4.3s \times 4.3s$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
 The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $~\citep{Rao2013}.
@@
 Lastly, the number of channels $Ch$ is equal to $8$ for both datasets, and the number of intensification sequences $k_a$ is statically assigned to $10$.
-The parameter $k$ used in the Near Neighbor $NN_T(d^l_u), \;l\in\{row,{col}\}$ is set to $k=7$.
+The parameter $k$ used in the Near Neighbor $NN_T(d^l_u), \;l\in\{row,{col}\}$ is set to $k=7$, following the suggestion of the article~\citep{Boiman2008}.
 In addition, the norm used on  Equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol} was the cosine norm, and descriptors were normalized to $ \left[ -1, 1 \right] $.
@@
 \section{Conclusion}
+\label{conclusion}
 
@@
 
-A method to detect transient P300 components from EEG signals based on their waveform characterization in time-space was presented.
-Additionally its validity was verified against a public dataset of ALS patients and an own dataset of healthy subjects.
+Among other applications of BCI analysis, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
 
-This method can be both, based on predefined templates or adaptive by including a training or calibration step to identify the templates.
-The adaptive behavior of the algorithm make it well suited when the shape of the pattern elicited by the P300 response does not conform to the predicted structure.
+In this work, a method to detect transient P300 components from EEG signals based on their waveform characterization in digital time-space, is presented.
+Additionally, its validity is evaluated using a public dataset of ALS patients and an own dataset of healthy subjects.
+The objective of this article is to explore new techniques of P300 automatic interpretation.
+We know that we have to keep working but the results are promising.
 
-The method works on a channel by channel basis.
-Should a calibration step be included, it can be used to identify the best performing channel, and may drastically reduce the number of required EEG electrodes, leading to the use of more ergonomic capturing device.
+The method works on a channel by channel basis; in this way the best performing channel can be identified and used it to reduce the number of required EEG electrodes, leading to the development of more ergonomic capturing device.
 
-We verified that the shape of the wave is more stable in occipital channels, where the performance at identifying letters  is higher.
-Besides novel applications of BCI, the goal of the entire discipline is to provide communication assistance to people affected by neuro-degenerative diseases.
-It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
-It may be worthy of further consideration to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
-Althought this, we did not find such an evidence in terms of the shape.
-The best response was obtained by an ALS subject.
-Moreover, this method can be used as an alternate \textit{BCI predictor} \citep{Clerc}, i.e.
-to detect BCI illiteracy or to predict the achievable performance of a given method.
-In the original work~\citep{Riccio2013} where the dataset was published, a link between a task prediction test and the performance of the p300 was found.
-At the same time, by analyzing the generated descriptors, which map in a very detailed and synthetic structure the shape information contained within the patch, a metric about the consistency of the \textit{shape} of the generated P300 response could also be derived.
+We observed that the shape of the wave is more stable in occipital channels, where the performance for identifying letters is higher.
+% It is understood that for people suffering from ALS, even though this pathology do not harm the CNS, their brain signals do suffer slow and deteriorating changes that may lead to different EEG patterns alterations \citep{Nijboer2009,Riener2014}.
+It is important to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
+Although,  we did not find any evidence in terms of the shape digital signal and the best response was obtained for an ALS subject.
+Moreover, this method can be used as an alternate BCI predictor.
 
-We also found that the stability of the P300 in terms of shape is crucial: Synchronization averaging, increase inter-stimulus interval, different montages, spatial filters can all of them affect the generated shape, and it is advisable to study which methods or combinations are best in order to stabilize the shape of the p300.
+By analyzing the generated descriptors, a metric about the shape of the P300 response can also be derived.
 
-Clearly, the best characteristic of this method is that closer collaboration with physicians can be fostered.
-We believe that the expanding and the understanding of this tool in order to automatically classify those patterns in EEG that are specifically identified by their shapes (e.g.
-K-Complex, Vertex Waves, Positive Occipital Sharp Transient \citep{Hartman2005}) is a prospect future work to be considered.
-We are currently working in unpublished material analyzing KComplex that may also provide  assistance to physician or electroencephalographers to help them locate these EEG patterns particularly in long recording periods, frequent in sleep research.
-Additionally, it can be used as a tool for artifact removal (which is performed on many occasions by visually inspecting the signal).
-This is due to the fact that the descriptors are directly based on how the signals behave in shape domain (i.e.
-\textit{how actually they looked like}).
+By means of the empirical experiments, we concluded that the stability of the P300 in terms of shape is crucial: synchronization averaging,  inter-stimulus interval,  montages, the signal to noise ratio and spatial filters can all of them affect the stability of the shape of the P300 ERP.
+
+In our opinion, the best application of this approach is that a closer collaboration with physicians can be fostered, because our method intent to imitate human visual observation.
+Automatically classify those patterns in EEG that are specifically identified by their shapes (e.g.
+K-Complex, Vertex Waves, Positive Occipital Sharp Transient~\citep{Hartman2005}) is a prospect future work to be considered.
+We are currently working in unpublished material analyzing KComplex that also provide  assistance to physician to locate these EEG patterns, specially in long recording periods, frequent in sleep research.
+Additionally, it can be used for artifact removal which is performed on many occasions by visually inspecting the signal.
+This is due to the fact that the descriptors are directly based on the signals behavior in shape domain.
 In line with these applications,  it can be used to build a database~\citep{Chavarriaga2017} of descriptors and improve atlases~\citep{Hartman2005}.
 
-In terms of the tool itself, the SIFT method is invariant to the scale, so in order to allow our method to be also invariant to the amplitude of the signal we had to modify the geometry of the patch to remove the squared constrain.
+We also want to solve the problem of signal stabilization by applying different shifts on the averaging procedure and another methods for obtaining a correct synchronization.
 
-Finally, waveform analysis and the methods that can be used to analyze them is an area where active research can benefit greatly both, physicians who need to automate some of the tools and practices and Brain Computer Interfaces or automatic EEG processing by using the extensive knowledge body available for clinical practices~\citep{Cole2017}
+%Finally, waveform analysis and the methods that can be used to analyze them is an area where active research can benefit greatly both, physicians who need to automate some of the tools and practices and Brain Computer Interfaces or automatic EEG processing by using the extensive knowledge body available for clinical practices~\citep{Cole2017}
 

### 2017-12-20 response.letter.tex
Casi todo listo.
Agregamos todos los cambios de las conclusiones, correcciones generales sobre todo el texto, el cambio para hacer digital las señales, otros cambios en los gráficos, cambios en los textos y otros.

@@
 {\color{blue}
-Thank you for your feedback. We have modified several points according to your recommendations, we hope the article is now better suited for publication. 
+Thank you for your comments and observations. We have modified several points according to your recommendations, we hope the article is now better suited for publication. 
 }
@@
 {\color{blue}
-You're  right, thank you very much for your comment.
+Thank you very much for your comment.
 We published in ~\citep{Ramele2016} the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to classify Motor Imagery.
-We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and classification of  SSVEP patterns.
+We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and classification of  SSVEP patterns but the results are in progress.
 We added this information on the article (Introduction and Conclusion, section 3).
@@
 The issue with variable T was fixed and we improved on the font styling to differentiate vectors and scalars.
-Section 1.1(Feature Extraction...) was modified as well as section 1.2.3(Parameters).
-The mathematical terminology were corrected all along the article.}
+Section 1.1 was modified as well as section 1.2.3.
+The mathematical terminology were corrected all along the article.
+We hope we have full addressed this point.
+}
 \end{quotation}
@@
 %We based that equation on the description of how the histogram of oriented gradients is calculated published in Vedaldi and Szeliski, Computer Vision and Algorithms.
-We reviewed the equation, clarified their terms and we hope it is now more clear.
+We reviewed the equation, corrected their terms and we hope it is now more clear.
 }
@@
 We modified all equations to emphasize the point that they are calculated in digital space, particularly on equation (4).
-We also added that image gradients are also calculated using finite differences, and added information about the support of each equation to clarifies that they are actually calculated in digital space.
-With this, we hope we have addressed your point.
+We also added that image gradients are also calculated using finite differences, and added information about the support of each equation to explain that they are actually calculated in digital space.
+With this, we hope we have full addressed your point.
  }
@@
 
-4.1) Another issue that I find worrisome is the design of experiments.\\
+4.1) Another issue that I find worrisome is the design of experiments. It seems that the number of subjects is rather low, it would be interesting to know why the authors decided to use such a subset.
 
@@
 {\color{blue}
-Thank you for the observation. 
- }
-\end{quotation}
-%
-\vskip+1ex
-\noindent \dotfill
-\vskip+1ex
-
-4.2) it seems that the number of subjects is rather low, it would be interesting to know why the authors decided to use such a subset.
-
-\vskip+1ex
-\noindent \dotfill
-\vskip+1ex
-%
-\begin{quotation}
-{\color{blue}
-Thank you for your suggestion.
-The original dataset was published as the 004-2014 dataset at the BNCI-Horizon website, which is very popular in BCI Research.
+Thank you for your comment.
+The original dataset was published as the 008-2014 dataset at the BNCI-Horizon website, which is very important in BCI Research.
 As far as we are aware of, experiments on this dataset of ALS patients has not been published by any independent group.
 In order to address this issue, we included a new own dataset with the same experimental conditions that we generated ourselves, on healthy subjects.
-We believe it now improves the number of different subjects and that can help to validate the method.
-We are also publishing this dataset on the kaggle platform for replication.
+We published this dataset on the kaggle platform for replication.
 }
@@
 {\color{blue}
+Thank you for your comments.
 We added more detailed information that address this issue.
-First in Section 1.1.4 (Speller Matrix Letter Identification) we included information about how we approach the signal averaging procedure.
-In Section 1.2.3 (Parameters) we added the parameters that were used.
-More importantly, in Section 2(Results and Discussion) we added information and references about the tests that we did to address latency and amplitude issues.
-We included in Section 3(Conclusion) information about future and current works where we are also dealing with these problems.
-We hope this is acceptable.
+First in Section 1.1.4  we included information about the signal averaging procedure.
+In Section 1.2.3 we added the parameters that were used.
+More importantly, in Section 2 we added information and references about the tests that we did to address latency and amplitude issues.
+We included in Section 3 information about future and current works where we are also dealing with these problems.
+We hope we full addressed your remarks.
 }
@@
 {\color{blue}
-We apologize for our mistake.
-The Results section was modified to clarify more precisely the expectation about the EEG channels where a visual P300 oddball paradigm generates stronger signals.
-We added references and improved the section to clarify this point more thoughtfully.
-We hope we may have addressed your point with this.
+Thank you for your suggestion.
+The Results section was modified to clarify the expectation about the EEG channels where a visual P300 oddball paradigm generates stronger signals.
+We also added references.
+We hope we  have addressed your point with this.
 }
@@
 Thank you very much for this information.
-We reviewed completely that equation (it is now equation (6) and equation (7)) and we included a step-by-step procedure that describes the algorithm that we propose and we used as well as a schematic chart in Figure 3.
+We reviewed completely that equation (it is now equation (6) and equation (7)).
+We included a step-by-step procedure that describes the proposed algorithm and we added a schematic chart in Figure 3.
 }

### 2017-12-20 article.tex
Agregamos todos los cambios de las conclusiones, correcciones generales sobre todo el texto, el cambio para hacer digital las señales, otros cambios en los gráficos, cambios en los textos y otros.

@@
 \usepackage{subfigure} 
+\usepackage{siunitx}
 
@@
 
-This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of gradient orientations extracted from the image of signal plots, and (2) describe the way in which this classification can be used to implement an offline P300-based BCI Speller application using two public datasets.
+This paper reports a method to, (1) classify P300 signals based on the identification of their structure in the shape domain using histograms of gradient orientations extracted from the image of signal plots, and (2) describe the way in which this classification can be used to implement an offline P300-based BCI Speller application using two datasets.
 Its validity is verified by processing offline data for ALS patients and for healthy subjects.
@@
 In Section~\ref{Protocol}, the experimental protocol is expounded.
-Section~\ref{Results} shows the results of applying the proposal technique and a discussion is presented.
+Section~\ref{Results} shows the results of applying the proposal technique and a discussion is presented as well.
 In the final Section~\ref{conclusion}  we expose our remarks, conclusions and future work.
@@
 
-\begin{figure}[H]
+\begin{figure}[htb]
 \centering
@@
 
-In this section, the signal preprocessing, the method for generating images from signal plots, the feature extraction procedure and the speller matrix identification are described.
+In this section, the signal preprocessing, the method for generating images from signal plots, the feature extraction procedure and the Speller Matrix identification are described.
 
@@
 \item \textbf{Artifact Removal}: The EEG signal matrix is processed on a channel by channel basis.
-For every 12 flashing stimuli, i.e.
-one complete sequence of intensification of each of the $6$ rows plus the $6$ columns, a basic artefact elimination procedure is implemented by removing the entire sequence when any signal deviates above/bellow $ \pm 70 \mu V $.
+For every complete sequence of intensification of $6$ rows and $6$ columns, a basic artifact elimination procedure is implemented by removing the entire sequence when any signal deviates above/bellow $ \pm 70 \mu V $.
 \item \textbf{Segmentation}: For each of the $12$ stimuli,  a window of $t_{max} = 1$ second of the multichannel signal is extracted, starting from the stimulus onset, corresponding to each row/column intensification.
@@
 \begin{equation}
-\tilde{x}(n,c) = \left \lfloor{ \gamma \cdot \frac{( x(n,c) - \bar{x}(c)  )}{ \sigma(c) } }\right \rfloor, n \in \{ 1, \dots, n_{max}\},c \in \{1,2,\dots,Ch\}
+\tilde{x}(n,c) = \left \lfloor{ \gamma \cdot \frac{( x(n,c) - \bar{x}(c)  )}{ \hat{\sigma}(c) } }\right \rfloor, \; n \in \{ 1, \dots, n_{max}\},\; c \in \{1,2,\dots,Ch\}
 \label{eq:standarizedaverages}
@@
 
-\noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related with the image scale.
+\noindent where $\gamma > 0$ is an input parameter of the algorithm and  it is related to the image scale.
 In addition, $ x(n,c) $ is the point-to-point averaged EEG matrix for the sample point $n$ and for channel $c$, $n_{max}=F_s.t_{max}$ and $F_s$ is the sampling frequency.
 The parameter  $Ch$ is the number of available EEG channels.
-Lastly, $$\bar{x}(c) =\frac{1}{n_{max}}\sum_{n=1}^{n_{max}}x(n,c)$$ and $$ \sigma(c) = (\frac{1}{n_{max}-1}\sum_{n=1}^{n_{max}}(x(n,c)-\bar{x}(c))^2 )^{\frac{1}{2}}$$ are the mean and standard deviation of $x(n,c), n \in 1,\dots,n_{max}$, for each channel $c$.
+Lastly, $$\bar{x}(c) =\frac{1}{n_{max}}\sum_{n=1}^{n_{max}}x(n,c)$$ and $$ \hat{\sigma}(c) = (\frac{1}{n_{max}-1}\sum_{n=1}^{n_{max}}(x(n,c)-\bar{x}(c))^2 )^{\frac{1}{2}}$$ are the mean and estimated standard deviation of $x(n,c), n \in 1,\dots,n_{max}$, for each channel $c$.
 
@@
 
+%$\max\limits_{1\leq j\leq n}$
+
 \noindent where $ (z_1,z_2) \in \mathbb{N} \times \mathbb{N}$ iterate over the width (based on the length of the signal segment) and height (based on the peak-to-peak amplitude) of the newly created image,  $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$.
-The function $z(c)$, $c \in \{1,2,\dots,Ch\}$ is the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image:
+The values $z(c)$, $c \in \{1,2,\dots,Ch\}$ are the location on the image where the signal's zero value has to be located in order to fit the entire signal within the image for each c:
 
@@
 \begin{equation}
- h(\theta,i,j) = 3s \sum_{\mathbf{x} \in B_{i,j}} w_\mathrm{ang}(\angle J(\mathbf{x}) - \theta)\, w_{ij}\left(\frac{\mathbf{x} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{x})|
+ h(\theta,i,j) = 3s \sum_{\mathbf{p} \in B_{i,j}} w_\mathrm{ang}(\angle J(\mathbf{p}) - \theta)\, w_{ij}\left(\frac{\mathbf{p} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{p})|
 \label{eq:histogram}
@@
 
-\noindent  where $\mathbf{x}$ is a pixel from  the $i,j$-block $B_{i,j}$, $ i,j \in \{1,2,3,4\} $, $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{x})| $ is the norm of the gradient vector in the pixel $\mathbf{x}$ and it is computed using finite differences, $\angle J(\mathbf{x}) $ is the angle of the gradient vector,  $\theta$ is the angle bin,   and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
+\noindent  where $\mathbf{p}$ is a pixel from  the $i,j$-block $B_{i,j}$, $ i,j \in \{1,2,3,4\} $,  $\theta$ is the angle bin with $ \theta \in \{0, 45, 90, 135, 180, 225, 270, 315\} $,  $ |J(\mathbf{p})| $ is the norm of the gradient vector in the pixel $\mathbf{p}$ and it is computed using finite differences, $\angle J(\mathbf{p}) $ is the angle of the gradient vector, and $ w_\mathrm{ang}(\cdot) $  and $ w_{ij}(\cdot) $ are linear interpolation functions used by Lowe and Vedaldi et al.
 in~\citep{Lowe2004,Vedaldi2010}.
@@
 As the patch has  $16$ blocks and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained.
-It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{x})| $, so the method considers both, the magnitude and the orientation of the gradient vector.
+It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{p})| $, so the method considers both, the magnitude and the orientation of the gradient vector.
 
@@
 The aim is to identify the selected letter from the matrix.
-Previously, during training phase, two descriptors are extracted from averaged signal segments which correspond to the letter where the user was supposed to be focusing onto.
+Previously, during the training phase, two descriptors are extracted from averaged signal segments which correspond to the letter where the user was supposed to be focusing onto.
 These descriptors are the P300 templates which are grouped in a template set called $ T $.
+This set is constructed using the steps 1-6 of the following algorithm.
 Segments corresponding to rows are labeled 1-6, whereas those corresponding to columns are labeled 7-12.
-The process has the following steps:
+The whole process has the following steps:
 
@@
 \item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
-\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(n,c),\dots,S_{k_a}(n,c)$, where the variables $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to time and channel, respectively.
+\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(n,c),\dots,S_{k_a}(n,c)$, where the variables $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to sample points and channel, respectively.
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
@@
 where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u, \;l\in\{row,{col}\}$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l\in\{row,{col}\}$.
+This set is obtained according to the k-NNBN  algorithm~\citep{Boiman2008}.
 
@@
 Figure~\ref{fig:classification} shows a scheme of this process.
+
 \begin{figure}[htp]
@@
 All healthy subjects have normal or corrected-to normal vision and no history of neurological disorders.
-The experiment is performed with 8 subjects, 6 males, 2 females, 6 right-handed, 2 left-handed, average age 28.25 years, standard deviation  9.58 years, range 20-50 years.
+The experiment is performed with 8 subjects, 6 males, 2 females, 6 right-handed, 2 left-handed, average age 29.00 years, standard deviation  11.56 years, range 20-56 years.
 
@@
 The vertical size consists of $4$ blocks of size $3 s_y$ pixels which is long enough as to contain the signal $\Delta  \mu V $, the peak-to-peak amplitude of the transient event.
-The horizontal size on the other hand, includes $4$ blocks  of $3 s_x$ and covers the entire duration in seconds of the transient signal event, $ \lambda $.
+The horizontal size includes $4$ blocks  of $3 s_x$ and covers the entire duration in seconds of the transient signal event, $ \lambda $.
 }
@@
 The parameter $k$ used in the Near Neighbor $NN_T(d^l_u), \;l\in\{row,{col}\}$ is set to $k=7$, following the suggestion of the article~\citep{Boiman2008}.
-In addition, the norm used on  Equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol} was the cosine norm, and descriptors were normalized to $ \left[ -1, 1 \right] $.
+In addition, the norm used on  Equations \ref{eq:multiclassificationrow} and \ref{eq:multiclassificationcol} is the cosine norm, and descriptors are normalized to $ \left[ -1, 1 \right] $.
 
@@
 
-\begin{table}[H]
+\begin{table}[htb]
 \caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each subject of the public dataset 008-2014.
-Twenty letters, four words of five letters each, were used for testing.
+The spelled words are \textit{GATTO}, \textit{MENTE}, \textit{VIOLA} and \textit{REBUS}.
 }
@@
 
-In Table~\ref{tab:resultsowndataset} results obtained for 8 healthy subjects are shown.   These results were obtained by the present method as described in previous sections.
+In Table~\ref{tab:resultsowndataset} results obtained for 8 healthy subjects are shown.
 
@@
 
-\begin{table}[H]
+\begin{table}[htb]
 \caption{Percentage of correctly predicted letters while performing an offline BCI Simulation for the best performing channel for each healthy subject.
-The method is applied on a channel by channel basis, and the best performing channel is shown.
+The spelled words are \textit{MANSO},\textit{CINCO},\textit{JUEGO} and \textit{QUESO}.
 }
@@
 The ensemble average is carried out in order to improve the SNR but other problems arise, for example inter-trial variability due to latencies shifts that may produce an unstable P300 signature.
-To solve this problem, we tested an approach to assess if the morphological shape of the P300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct resynchronization is applied, but results are still in process.
+To solve this problem, we tested an approach to assess if the morphological shape of the P300 can be stabilized by applying different shifts and we verified that there is a better performance when a correct single-trial alignment is applied, but results are still in process.
 We also applied Dynamic Time Warping (DTW)~\citep{Casarotto2005} but we were unable to find a substantial improvement.
@@
 \end{figure}
-
-%\begin{itemize}
-%\item Lo que pusimos antes de la sincronización de las señales apunta a la latencia.
-%\item Por el lado de la amplitud, lo que hicimos fue ir por un lado conservador de llevar todo a varianza uno.
-%\item Si bien el zscore nos ayuda también genera otro problema adicional que es el que ponemos en el texto (y por el cual con razón Juan me criticaba).
-
-%\end{itemize}
-
-%La estandarizacion produce dos efectos contrarios entre si, contrapues.
-
-%Por un lado reduce el ruido y outliers
-%por otro lado reduce aquellos p300 significativos positivos que ayudarian su identificacion.
 
@@
 This is another advantage of this approach.
-On the downsize, the standardizing process reduces the amplitude of any significant P300 complex diminishing its automatic interpretation capability.
+On the hand, the standardizing process reduces the amplitude of any significant P300 complex diminishing its automatic interpretation capability.
 
@@
 It is important to evaluate if there is any correlation between the lower performance obtained for some subjects and the characterization of their ALS stage.
-Although,  we did not find any evidence in terms of the shape digital signal and the best response was obtained for an ALS subject.
+Although,  we did not find any evidence in terms of the digital signal shape and the best response was obtained for an ALS subject.
 Moreover, this method can be used as an alternate BCI predictor.
@@
 In our opinion, the best application of this approach is that a closer collaboration with physicians can be fostered, because our method intent to imitate human visual observation.
-Automatically classify those patterns in EEG that are specifically identified by their shapes (e.g.
+Automatic classification of patterns in EEG that are specifically identified by their shapes (e.g.
 K-Complex, Vertex Waves, Positive Occipital Sharp Transient~\citep{Hartman2005}) is a prospect future work to be considered.

### 2017-12-21 article.tex
Casi listo

@@
 % Keywords
-\keyword{EEG; BCI; P300; ALS; NBNN; HGOSP; SIFT}
+\keyword{EEG; BCI; SIFT; P300; ALS; NBNN; HGOSP}
 
@@
 \begin{equation}
-z(c) = \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(n,c)  - \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(t,c)  + \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c)}{ 2} }\right \rfloor
+z(c) = \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(n,c)  - \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c) }{2} }\right \rfloor -   \left \lfloor{ \frac{\max_{1 \leq n\leq n_{max}} \tilde{x}(n,c)  + \min_{1 \leq n\leq n_{max}} \tilde{x}(n,c)}{ 2} }\right \rfloor
 \label{eq:zerolevel}
@@
 A local image patch of size $S_p \times S_p$ pixels is constructed by dividing the window in $16$ blocks of size $3s$ each one,  where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
-It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, thus $S_p = 4.3s $ pixels.
+It is arranged in a $4 \times 4$ grid and the pixel $ \mathbf{kp}$ is the patch center, thus $S_p = 4.3.s $ pixels.
 
@@
 As the patch has  $16$ blocks and  $8$ bin angles are considered, a descriptor of $128$ dimension is obtained.
-It can be observed that in each step, the histogram is computed by multiplying by $ |J(\mathbf{p})| $, so the method considers both, the magnitude and the orientation of the gradient vector.
+It can be observed that the histogram is computed by multiplying by $ |J(\mathbf{p})| $, so the method considers both, the magnitude and the orientation of the gradient vector.
+
+%\begin{figure}[htp]
+%\centering
+%\includegraphics[width=8cm]{histogramchart.eps}
+%\includegraphics[width=6cm]{samplegradients.png}
+%\caption{Single trial segments $S_i$ are averaged for the 6 rows and 6 columns.
+From the averaged signal, the image of the signal plot is generated and each descriptor is computed.
+By comparing each descriptor against the set of templates, the P300 ERP can be detected, and finally the desired letter from the matrix can be inferred.}
+%\label{fig:fdsfdsfsd}
+%\end{figure}
+%
+%\begin{figure}[htb]
+%\centering
+%\subfigure[Plot of the signal, a keypoint and the corresponding patch.]{\includegraphics[width=.49\linewidth]{histogramchart.eps}\label{histogram}}
+%\subfigure[Orientations on two blocks of the patch.]{\includegraphics[width=.49\linewidth]{samplegradients.png}\label{samplegradients}}
+%\caption{ Example of a patch and a scheme of the orientation's histogram computation.}
+%\label{fig:histogramchart}
+%\end{figure}
 
@@
 \item \label{paso1}Highlight randomly the rows and columns from the matrix.  There is one row and one column that should match the letter selected by the subject.
-\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(n,c),\dots,S_{k_a}(n,c)$, where the variables $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to sample points and channel, respectively.
+\item  \label{paso2} Repeat step~\ref{paso1} $k_a$ times, obtaining the single trial segments $S_1(n,c),\dots,S_{k_a}(n,c)$, of the EEG signal where the variables $n \in \{1, \dots, n_{max}\}$ and $c \in \{1,2,\dots,Ch\}$ correspond to sample points and channel, respectively.
 The parameter $k_a$ is the number of repetitions and it is an input parameter of the algorithm.
@@
 
-\item \label{paso4}Plot the signal $x(n,c)$,  $n \in \{1, \dots, n_{max}\}$, $c \in \{1,\dots,Ch\}$,  according Section~\ref{Plot}.
+\item \label{paso4}Plot the signals $x(n,c)$,  $n \in \{1, \dots, n_{max}\}$, $c \in \{1,\dots,Ch\}$,  according Section~\ref{Plot}.
 
@@
 
-where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u, \;l\in\{row,{col}\}$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u), \;l\in\{row,{col}\}$.
-This set is obtained according to the k-NNBN  algorithm~\citep{Boiman2008}.
+where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u)$.
+This set is obtained according to the k-NBNN  algorithm~\citep{Boiman2008}.
 
@@
 
-Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same locations Fz, Cz, Pz, P3,P4, Oz, PO7 and PO8.
+Gel-based active electrodes (g.LADYbird, g.Tec, Austria) are used on the same locations Fz, Cz, Pz, Oz, P3,P4, PO7 and PO8.
 Reference is set to the right ear lobe and ground is preset as the AFz position.
@@
 
-The patch size is $S_P = 4.3s \times 4.3s$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
+The patch size is $S_P = 4.3.s \times 4.3.s$ pixels, where $s$ is the scale of the local patch and it is an input parameter of the algorithm.
 The P300 event can have a span of $400$ ms and its amplitude can reach $ 10 \mu V $~\citep{Rao2013}.
@@
 \noindent where $ \lambda $ is the length in seconds covered by the patch, $ Fs $ is the sampling frequency of the EEG signal (downsampled to 16 Hz) and  $\Delta  \mu V $ corresponds to the amplitude in microvolts that can be covered by the height of the patch.
-The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.
+The geometric structure of the patch forces a squared configuration, then we discerned that by using $ s =s_x =s_y = 3 $ and $ \gamma = 4 $,  the local patch and the descriptor can identify events of 9 $ \mu V $ of amplitude, with a span of $ \lambda = 0.56$ seconds.
 This also provides that $ 1 $ pixel represents $ \frac{1}{\gamma}= \frac{1}{4} \mu V $ on the vertical direction and $\frac{1}{F_s.\gamma}=\frac{1}{64}$ seconds on the horizontal direction.
@@
 
+El uso de descriptores basados en histogramas, presentado en este trabajo, también puede ser usado para derivar una métrica acerca de la forma de la señal de P300.
+
 By means of the empirical experiments, we concluded that the stability of the P300 in terms of shape is crucial: synchronization averaging,  inter-stimulus interval,  montages, the signal to noise ratio and spatial filters can all of them affect the stability of the shape of the P300 ERP.

### 2017-12-21 response.letter.tex
Casi listo
Ultimos cambios finales. ya mandé el paper.

@@
 Thank you very much for your comment.
-We published in ~\citep{Ramele2016} the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to classify Motor Imagery.
+We published in~\citep{Ramele2016} the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to classify Motor Imagery.
 We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and classification of  SSVEP patterns but the results are in progress.
@@
 In order to address this issue, we included a new own dataset with the same experimental conditions that we generated ourselves, on healthy subjects.
-We published this dataset on the kaggle platform for replication.
+We published this dataset on the Kaggle platform for replication.
 }
@@
 \vskip+1ex
-
-\bibliography{response.letter}
+\bibliographystyle{mdpi}
+\bibliography{article}
 

### 2017-12-21 article.tex
Ultimos cambios finales. ya mandé el paper.

@@
 EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process.
-Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.
+Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring adaptive and co-adaptive calibration and learning procedures~\citep{Clerc}.
 Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.
@@
 \begin{equation}
- h(\theta,i,j) = 3s \sum_{\mathbf{p} \in B_{i,j}} w_\mathrm{ang}(\angle J(\mathbf{p}) - \theta)\, w_{ij}\left(\frac{\mathbf{p} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{p})|
+ h(\theta,i,j) = 3.s \sum_{\mathbf{p} \in B_{i,j}} w_\mathrm{ang}(\angle J(\mathbf{p}) - \theta)\, w_{ij}\left(\frac{\mathbf{p} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{p})|
 \label{eq:histogram}
@@
 
-where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u)$.
+where $NN_T(d^l_u)$  is the set of the $k$ nearest neighbors to $d^l_u$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u),\;l\in\{row,{col}\}$.
 This set is obtained according to the k-NBNN  algorithm~\citep{Boiman2008}.
@@
 Participants are seated in a comfortable chair, with their vision aligned to a computer screen located one meter in front of them.
-The handling and processing of the data and stimulations is conducted by the OpenVibe platform~\citep{Renard2010}.
+The handling and processing of the data and stimuli is conducted by the OpenVibe platform~\citep{Renard2010}.
 
@@
 
-By analyzing the generated descriptors, a metric about the shape of the P300 response can also be derived.
-
-El uso de descriptores basados en histogramas, presentado en este trabajo, también puede ser usado para derivar una métrica acerca de la forma de la señal de P300.
+%By analyzing the generated descriptors, a metric about the shape of the P300 response can also be derived.  
+The use of descriptors based on  histogram of gradient orientation, presented in this work, also can be utilized for deriving a shape metric in the space of the P300 signals.
+%El uso de descriptores basados en histogramas, presentado en este trabajo, también puede ser usado para derivar una métrica acerca de la forma de la señal de P300.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.
-The authors would like to thank Dr.
+\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.} %The authors would like to thank Dr.
 Juliana Gambini for their insights and help with the detailed description of the histogram of gradients.}
